---
title: "Collaboration by Time of Day with Viva Insights in R"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: "lumen"
---

```{r setup, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(tidyverse)
library(vivainsights)
library(lubridate)
```

# Collaboration by Time of Day

## Introduction

A common question is: *"what time does the typical person start and end their
working day, and how does that differ by team, region, or role?"* Viva Insights
exposes hourly **collaboration by time of day** metrics, which count the emails
and chats sent and the fractions of the hour spent in meetings and unscheduled
calls for each hour of the day, and these let us answer it.

This document shows how to turn those hourly buckets into a typical **start** and
**end** of day, using the same definition of an "active hour" as the Microsoft
product: an hour is active if **any** of chats sent, emails sent, meetings, or
unscheduled calls is greater than zero.

The package's sample datasets do not include the hourly "by time of day" columns,
so we **simulate** a person-by-day dataset with those columns, named as an
`import_query()` of a Person or Daily query would name them. To run on your own
data, select the *Emails sent*, *Chats sent*, *Meetings*, and *Unscheduled calls*
time-of-day metrics in a Person or Daily query, import it, and replace the
`simulate_time_of_day()` call below, and the downstream code is unchanged.

## Simulate a person-by-day dataset

Each person has a latent start and end of day (managers end later); activity is
generated per hour, denser inside working hours and sparse outside them.

```{r}
simulate_time_of_day <- function(n_person = 200, n_days = 20, seed = 7) {
  set.seed(seed)
  persons    <- sprintf("P%03d", seq_len(n_person) - 1)
  is_manager <- runif(n_person) < 0.25
  start_true <- pmin(pmax(rnorm(n_person, 8.5, 0.7), 6), 10)
  end_true   <- pmin(pmax(rnorm(n_person, 18.0, 1.0) + is_manager * 1.0, 15), 22)

  # One row per person-day (weekdays only)
  all_dates <- seq(as.Date("2024-01-01"), by = "day", length.out = 40)
  dates     <- all_dates[!(lubridate::wday(all_dates) %in% c(1, 7))][seq_len(n_days)]

  grid <- expand.grid(PersonId = persons, MetricDate = dates,
                      stringsAsFactors = FALSE)
  m   <- nrow(grid)
  pos <- match(grid$PersonId, persons)

  # m x 24 hour grid; comparison recycles the length-m person vectors down columns
  hours   <- matrix(0:23, nrow = m, ncol = 24, byrow = TRUE)
  in_work <- (hours >= start_true[pos]) & (hours < end_true[pos])
  active  <- matrix(runif(m * 24), nrow = m) < ifelse(in_work, 0.85, 0.02)

  out <- grid
  for (h in 0:23) {
    a   <- active[, h + 1]
    col <- sprintf("%02d_%02d", h, h + 1)
    out[[paste0("Chats_sent_", col)]]        <- a * rpois(m, 1.0)
    out[[paste0("Emails_sent_", col)]]       <- a * rpois(m, 0.6)
    out[[paste0("Meetings_", col)]]          <- a * runif(m) * 0.5   # fraction of hour
    out[[paste0("Unscheduled_calls_", col)]] <- a * (runif(m) < 0.1) * runif(m)
  }
  out$IsManager <- is_manager[pos]
  tibble::as_tibble(out)
}

# On your own data, replace this with vivainsights::import_query("your_query.csv").
df <- simulate_time_of_day()
dim(df)
```

## Build the activity matrix and find first / last active hour

An hour is active if any of the four metrics is greater than zero in that bucket.
We use a fast vectorised approach: build a logical matrix of active hours, then use
`max.col()` to locate the first and last active hour in each row.

```{r}
METRICS <- c("Chats_sent", "Emails_sent", "Meetings", "Unscheduled_calls")
hour_columns <- function(prefix) sprintf("%s_%02d_%02d", prefix, 0:23, 1:24)

active_matrix <- matrix(FALSE, nrow = nrow(df), ncol = 24)
for (metric in METRICS) {
  active_matrix <- active_matrix | (as.matrix(df[hour_columns(metric)]) > 0)
}

has_any  <- rowSums(active_matrix) > 0
first_hr <- max.col(active_matrix, ties.method = "first") - 1L
last_hr  <- 24L - max.col(active_matrix[, 24:1, drop = FALSE], ties.method = "first")

# Start = bucket start hour; End = bucket end hour (+1). Drop days with no activity.
day_level <- df %>%
  transmute(PersonId, MetricDate, IsManager,
            start = ifelse(has_any, first_hr, NA_real_),
            end   = ifelse(has_any, last_hr + 1, NA_real_)) %>%
  drop_na(start, end)

head(day_level)
```

## A typical start and end of day

We aggregate in **two stages**: each person's median across their days, then the
mean across people. This picks up sub-hour shifts a single pooled median would
miss, while staying robust to a few unusual days per person. We format decimal
hours as `HH:MM`.

```{r}
hhmm <- function(x) {
  ifelse(is.na(x), NA_character_,
         sprintf("%02d:%02d", floor(x), round((x - floor(x)) * 60)))
}

person_med <- day_level %>%
  group_by(PersonId) %>%
  summarise(start = median(start), end = median(end), .groups = "drop")

typical <- person_med %>% summarise(start = mean(start), end = mean(end))
pooled  <- day_level  %>% summarise(start = median(start), end = median(end))

tibble(
  measure = c("Typical (two-stage)", "Pooled median (check)"),
  start   = c(hhmm(typical$start), hhmm(pooled$start)),
  end     = c(hhmm(typical$end),   hhmm(pooled$end))
)
```

## Cuts by day of week and by role

```{r}
# Derive weekday numerically then map to labels (label = TRUE is unreliable on
# some platforms).
day_level <- day_level %>%
  mutate(weekday = factor(lubridate::wday(MetricDate, week_start = 1),
                          levels = 1:7,
                          labels = c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun")))

two_stage_by <- function(data, group_cols) {
  data %>%
    group_by(across(all_of(c("PersonId", group_cols)))) %>%
    summarise(start = median(start), end = median(end), .groups = "drop") %>%
    group_by(across(all_of(group_cols))) %>%
    summarise(start = mean(start), end = mean(end), .groups = "drop") %>%
    mutate(start = hhmm(start), end = hhmm(end))
}

two_stage_by(day_level, "weekday")
```

```{r}
two_stage_by(day_level, "IsManager") %>%
  mutate(role = ifelse(IsManager, "Manager", "Non-manager")) %>%
  select(role, start, end)
```

## Notes

- The two-stage (person-median then mean-across-people) metric picks up sub-hour
  shifts that a single pooled median would miss, while staying robust to a few
  unusual days per person.
- Days with no recorded activity are excluded rather than counted as a `00:00`
  start, which would badly bias the result.
- On real data, be careful about the timezone basis of the hourly buckets for a
  globally distributed population, and compare within a region where possible.
