---
title: "Copilot Consumption and Ways of Working"
subtitle: "The art of the possible with Consumption, GitHub credits and Viva Insights data (fully synthetic)"
date: "Synthetic window: 4 Jan-28 Jun 2026"
output:
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: scroll
    theme: lumen
    self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.width = 8.5, fig.height = 4.5)
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
library(knitr)
library(ggrepel)

MIN_GROUP_N <- 10L
PAL <- c("#1F4E79", "#2E75B6", "#7FA8D0", "#BDD3E8",
         "#C9C9C9", "#B03A2E", "#5B8C5A", "#7D5BA6")

theme_t <- function() {
  theme_minimal(base_size = 11) +
    theme(
      panel.grid.minor = element_blank(),
      legend.position = "bottom",
      plot.title.position = "plot",
      plot.title = element_text(face = "bold", colour = PAL[1]),
      plot.subtitle = element_text(colour = "grey30",
                                   margin = margin(t = 2, b = 9)),
      plot.caption = element_text(colour = "grey45", size = 8, hjust = 0,
                                  margin = margin(t = 7)),
      plot.margin = margin(t = 10, r = 16, b = 4, l = 4)
    )
}

read_demo <- function(...) {
  read.csv(file.path("_data", ...),
           check.names = FALSE, stringsAsFactors = FALSE)
}

week_start_sunday <- function(x) {
  x <- as.Date(x)
  x - as.integer(format(x, "%w"))
}

clean_label <- function(x, fallback = "Not disclosed") {
  x <- as.character(x)
  if_else(is.na(x) | !nzchar(trimws(x)), fallback, x)
}

safe_divide <- function(num, den) {
  den <- rep_len(den, length(num))
  out <- rep(NA_real_, length(num))
  valid <- !is.na(den) & den > 0
  out[valid] <- num[valid] / den[valid]
  out
}

people_meta <- read_demo("consumption-query", "PeopleMetaData.csv")
m365_daily_raw <- read_demo("consumption-query", "PersonM365CreditsMetrics.csv") |>
  mutate(
    MetricDate = as.Date(MetricDate),
    WeekStart = week_start_sunday(MetricDate)
  )
github_daily_raw <- read_demo("consumption-query", "PersonGitHubCreditsMetrics.csv") |>
  mutate(
    MetricDate = as.Date(MetricDate),
    WeekStart = week_start_sunday(MetricDate)
  )
pq <- read_demo("person-query", "PersonQuery.csv") |>
  mutate(MetricDate = as.Date(MetricDate))

stopifnot(
  max(abs(pq$Meeting_hours + pq$Scheduled_call_hours +
            pq$Available_to_focus_hours - 40), na.rm = TRUE) <= 0.01,
  max(abs(pq$Uninterrupted_hours + pq$Interrupted_hours -
            pq$Available_to_focus_hours), na.rm = TRUE) <= 0.01,
  all(pq$Uninterrupted_hours <= pq$Available_to_focus_hours + 0.01,
      na.rm = TRUE),
  all(pq$Open_1_hour_block <= pq$Available_to_focus_hours + 0.01,
      na.rm = TRUE)
)

person_keys <- bind_rows(
  select(m365_daily_raw, PersonId, PeopleHistoricalId),
  select(github_daily_raw, PersonId, PeopleHistoricalId)
) |>
  filter(!is.na(PersonId), nzchar(PersonId),
         !is.na(PeopleHistoricalId), nzchar(PeopleHistoricalId)) |>
  distinct()

stopifnot(
  nrow(count(person_keys, PersonId) |> filter(n > 1)) == 0,
  nrow(count(person_keys, PeopleHistoricalId) |> filter(n > 1)) == 0
)

credit_key_rows <- bind_rows(
  select(m365_daily_raw, PeopleHistoricalId),
  select(github_daily_raw, PeopleHistoricalId)
)

unmatched_key_counts <- tibble(
  Check = c(
    "PeopleMetaData people with no activity-file crosswalk",
    "Credit activity rows with PeopleHistoricalId absent from PeopleMetaData"
  ),
  Count = c(
    anti_join(people_meta, person_keys, by = "PeopleHistoricalId") |> nrow(),
    anti_join(credit_key_rows, people_meta, by = "PeopleHistoricalId") |> nrow()
  )
)

m365_daily_person <- m365_daily_raw |>
  group_by(PeopleHistoricalId, PersonId, MetricDate, WeekStart) |>
  summarise(
    m365_sessions = sum(`Session count`, na.rm = TRUE),
    m365_credits = sum(`Total Copilot Credits used`, na.rm = TRUE),
    .groups = "drop"
  )

m365_weekly <- m365_daily_person |>
  group_by(PersonId, MetricDate = WeekStart) |>
  summarise(
    m365_sessions = sum(m365_sessions, na.rm = TRUE),
    m365_credits = sum(m365_credits, na.rm = TRUE),
    .groups = "drop"
  )

m365_service_weekly <- m365_daily_raw |>
  group_by(PersonId, MetricDate = WeekStart, ServiceName) |>
  summarise(
    m365_sessions = sum(`Session count`, na.rm = TRUE),
    m365_credits = sum(`Total Copilot Credits used`, na.rm = TRUE),
    .groups = "drop"
  )

github_weekly <- github_daily_raw |>
  group_by(PersonId, MetricDate = WeekStart) |>
  summarise(github_credits = sum(`Total GitHub AI Credits used`, na.rm = TRUE),
            .groups = "drop")

PRODUCT_WINDOW_START <- min(c(m365_weekly$MetricDate, github_weekly$MetricDate),
                            na.rm = TRUE)
PRODUCT_WINDOW_END <- max(c(m365_weekly$MetricDate, github_weekly$MetricDate),
                          na.rm = TRUE)

product_weeks <- sort(unique(c(m365_weekly$MetricDate, github_weekly$MetricDate)))

person_week <- pq |>
  filter(MetricDate %in% product_weeks) |>
  select(PersonId, MetricDate, Collaboration_hours,
         Meeting_hours, Email_hours, Chat_hours, Scheduled_call_hours,
         Unscheduled_call_hours, Channel_message_hours,
         Active_connected_hours, External_collaboration_hours,
         Available_to_focus_hours, Uninterrupted_hours, Interrupted_hours,
         Open_1_hour_block,
         Recurring_meeting_hours, Conflicting_meeting_hours,
         Meeting_hours_with_six_or_fewer_hours_of_advanced_notice,
         Emails_sent, Chats_sent, Meetings, Calls,
         After_hours_collaboration_hours, After_hours_meeting_hours,
         After_hours_email_hours, After_hours_chat_hours,
         Weekend_collaboration_hours, Collaboration_span,
         Internal_network_size, External_network_size, Strong_ties,
         Diverse_ties, Network_outside_organization,
         Total_Copilot_enabled_days) |>
  left_join(select(m365_weekly, PersonId, MetricDate,
                   m365_sessions, m365_credits),
            by = c("PersonId", "MetricDate")) |>
  left_join(select(github_weekly, PersonId, MetricDate, github_credits),
            by = c("PersonId", "MetricDate")) |>
  mutate(
    m365_sessions = replace_na(m365_sessions, 0),
    m365_credits = replace_na(m365_credits, 0),
    github_credits = replace_na(github_credits, 0),
    total_credits = m365_credits + github_credits,
    credits_per_m365_session = safe_divide(m365_credits, m365_sessions)
  )

WOW_METRICS <- c(
  "Collaboration_hours", "Active_connected_hours",
  "Email_hours", "Chat_hours", "Meeting_hours",
  "Unscheduled_call_hours", "After_hours_collaboration_hours",
  "Collaboration_span", "Weekend_collaboration_hours",
  "Available_to_focus_hours", "Uninterrupted_hours", "Interrupted_hours",
  "Internal_network_size", "External_network_size"
)

metric_labels <- c(
  Collaboration_hours = "Collaboration hours / person / week",
  Active_connected_hours = "Active connected hours / person / week",
  Email_hours = "Email hours / person / week",
  Chat_hours = "Chat hours / person / week",
  Meeting_hours = "Meeting hours / person / week",
  Unscheduled_call_hours = "Unscheduled call hours / person / week",
  After_hours_collaboration_hours =
    "After-hours collaboration hours / person / week",
  Collaboration_span = "Collaboration span (hours)",
  Weekend_collaboration_hours =
    "Weekend collaboration hours / person / week",
  Available_to_focus_hours = "Available-to-focus hours / person / week",
  Uninterrupted_hours = "Uninterrupted hours / person / week",
  Interrupted_hours = "Interrupted hours / person / week",
  Internal_network_size = "Internal network size (people)",
  External_network_size = "External network size (people)"
)

person_consumption <- person_week |>
  group_by(PersonId) |>
  summarise(
    weekly_total_credits = mean(total_credits),
    weekly_m365_credits = mean(m365_credits),
    weekly_github_credits = mean(github_credits),
    weekly_m365_sessions = mean(m365_sessions),
    active_weeks = sum(total_credits > 0),
    total_credits = sum(total_credits),
    total_m365_credits = sum(m365_credits),
    total_github_credits = sum(github_credits),
    total_m365_sessions = sum(m365_sessions),
    product_weeks = n(),
    credits_per_m365_session =
      safe_divide(sum(m365_credits), sum(m365_sessions)),
    .groups = "drop"
  ) |>
  mutate(
    positive_consumer = total_credits > 0,
    active_week_share = active_weeks / product_weeks,
    positive_quartile = if_else(
      positive_consumer,
      ntile(if_else(positive_consumer, weekly_total_credits, NA_real_), 4L),
      NA_integer_
    ),
    CreditQuartile = factor(
      case_when(
        !positive_consumer ~ "No consumption",
        positive_quartile == 1L ~ "Q1 lowest",
        positive_quartile == 2L ~ "Q2",
        positive_quartile == 3L ~ "Q3",
        TRUE ~ "Q4 highest"
      ),
      levels = c("No consumption", "Q1 lowest", "Q2", "Q3", "Q4 highest")
    )
  )

high_credit_cut <- quantile(
  person_consumption$weekly_total_credits[
    person_consumption$positive_consumer
  ],
  0.75, na.rm = TRUE
)

high_cost_cut <- quantile(
  person_consumption$credits_per_m365_session[
    person_consumption$total_m365_sessions >= MIN_GROUP_N
  ],
  0.75, na.rm = TRUE
)

person_attributes <- pq |>
  arrange(PersonId, MetricDate) |>
  group_by(PersonId) |>
  summarise(
    Organization = first(clean_label(Organization)),
    FunctionType = first(clean_label(FunctionType)),
    LevelDesignation = first(clean_label(LevelDesignation)),
    IsManager = first(clean_label(IsManager)),
    Team = first(clean_label(Team)),
    Role = first(clean_label(Role)),
    Seniority = first(clean_label(Seniority)),
    Tenure = first(clean_label(Tenure)),
    IsDeveloper = first(clean_label(IsDeveloper)),
    IsCopilotLicensed = any(Total_Copilot_enabled_days > 0, na.rm = TRUE),
    .groups = "drop"
  )

person_base <- person_attributes |>
  left_join(person_keys, by = "PersonId") |>
  left_join(person_consumption, by = "PersonId") |>
  mutate(
    across(c(weekly_total_credits, weekly_m365_credits,
             weekly_github_credits, weekly_m365_sessions, total_credits,
             total_m365_credits, total_github_credits,
             total_m365_sessions, active_weeks, active_week_share),
           \(x) replace_na(x, 0)),
    positive_consumer = replace_na(positive_consumer, FALSE),
    CreditQuartile = replace_na(as.character(CreditQuartile),
                                "No consumption"),
    CreditQuartile = factor(
      CreditQuartile,
      levels = c("No consumption", "Q1 lowest", "Q2", "Q3", "Q4 highest")
    ),
    CreditBand = case_when(
      weekly_total_credits >= high_credit_cut & positive_consumer ~
        "High credit",
      positive_consumer ~ "Lower credit",
      TRUE ~ "No consumption"
    ),
    CostPerSessionProfile = case_when(
      !positive_consumer ~ "No consumption",
      total_m365_sessions < MIN_GROUP_N ~ "Limited M365 sessions",
      credits_per_m365_session >= high_cost_cut ~ "High cost per session",
      TRUE ~ "Standard cost per session"
    ),
    ConsumptionProfile = case_when(
      CreditBand == "High credit" &
        CostPerSessionProfile == "High cost per session" ~
          "High credit, high cost/session",
      CreditBand == "High credit" ~ "High credit, standard cost/session",
      CreditBand == "Lower credit" &
        CostPerSessionProfile == "High cost per session" ~
          "Lower credit, high cost/session",
      CreditBand == "Lower credit" ~ "Lower credit, standard cost/session",
      TRUE ~ "No consumption"
    ),
    CreditActivitySegment = case_when(
      !positive_consumer ~ "Non-user",
      active_weeks >= 10 & weekly_total_credits >= high_credit_cut ~
        "Power User",
      active_weeks >= 8 ~ "Habitual User",
      active_weeks >= 3 ~ "Novice User",
      TRUE ~ "Low User"
    ),
    CreditActivitySegment = factor(
      CreditActivitySegment,
      levels = c("Power User", "Habitual User", "Novice User",
                 "Low User", "Non-user"),
      ordered = TRUE
    ),
    IsCopilotLicensed = if_else(IsCopilotLicensed, "true", "false")
  )

person_wow <- person_week |>
  group_by(PersonId) |>
  summarise(across(all_of(WOW_METRICS), \(x) mean(x, na.rm = TRUE)),
            .groups = "drop") |>
  left_join(person_base, by = "PersonId")

function_summary <- person_wow |>
  group_by(FunctionType) |>
  summarise(
    People = n(),
    `Positive users` = sum(positive_consumer),
    `Weekly credits` = mean(weekly_total_credits),
    `M365 credits` = mean(weekly_m365_credits),
    `GitHub credits` = mean(weekly_github_credits),
    `M365 sessions` = mean(weekly_m365_sessions),
    `M365 credits / session` =
      safe_divide(sum(total_m365_credits), sum(total_m365_sessions)),
    `Collaboration hours` = mean(Collaboration_hours),
    `After-hours hours` = mean(After_hours_collaboration_hours),
    `Collaboration span` = mean(Collaboration_span),
    `Available-to-focus hours` = mean(Available_to_focus_hours),
    `Uninterrupted hours` = mean(Uninterrupted_hours),
    `Interrupted hours` = mean(Interrupted_hours),
    `Internal network size` = mean(Internal_network_size),
    .groups = "drop"
  ) |>
  filter(People >= MIN_GROUP_N)

FUNCTION_ORDER <- function_summary |>
  arrange(`Weekly credits`) |>
  pull(FunctionType)

overall <- person_base |>
  summarise(
    People = n(),
    Developers = sum(IsDeveloper == "Yes"),
    Licensed = sum(IsCopilotLicensed == "true"),
    Consumers = sum(positive_consumer),
    `Weekly credits` = mean(weekly_total_credits),
    `M365 credits / session` =
      safe_divide(sum(total_m365_credits), sum(total_m365_sessions)),
    `M365 share` =
      safe_divide(sum(total_m365_credits), sum(total_credits)),
    `GitHub share` =
      safe_divide(sum(total_github_credits), sum(total_credits))
  )

population_counts <- tibble(
  Metric = c("People", "Developers", "Organisations", "Functions"),
  Value = c(
    nrow(person_base),
    sum(person_base$IsDeveloper == "Yes"),
    n_distinct(person_base$Organization),
    n_distinct(person_base$FunctionType)
  )
)

CHART_CAPTION <- sprintf(
  paste0(
    "Synthetic data | n = %s people (%s developers) | ",
    "product window %s to %s"
  ),
  comma(overall$People),
  comma(overall$Developers),
  format(PRODUCT_WINDOW_START, "%d %b %Y"),
  format(PRODUCT_WINDOW_END, "%d %b %Y")
)

std_diff <- function(x, grp) {
  keep <- !is.na(x) & !is.na(grp)
  x <- x[keep]
  grp <- grp[keep]
  a <- x[grp]
  b <- x[!grp]
  if (length(a) < MIN_GROUP_N || length(b) < MIN_GROUP_N) return(NA_real_)
  pooled <- sqrt(((length(a) - 1) * var(a) + (length(b) - 1) * var(b)) /
                   (length(a) + length(b) - 2))
  if (is.na(pooled) || pooled == 0) return(NA_real_)
  (mean(a) - mean(b)) / pooled
}
```

<style>
body {
  font-family: "Segoe UI Variable", "Segoe UI", system-ui, sans-serif;
  color: #374151;
  background: #F4F4F4;
}
.navbar {
  min-height: 52px;
}
.navbar-brand {
  font-weight: 600;
}
.chart-wrapper {
  border: 1px solid #E5E7EB;
  border-radius: 8px;
  box-shadow: none;
}
.chart-title {
  color: #374151;
  font-weight: 500;
}
.value-box {
  min-height: 108px !important;
  height: 108px !important;
  color: #111827 !important;
  background: #FFFFFF !important;
  border: 1px solid #E5E7EB;
  border-top: 4px solid #3B82F6;
  border-radius: 8px;
  box-shadow: none;
}
.value-box-info { border-top-color: #14B8A6; }
.value-box-success { border-top-color: #8B5CF6; }
.value-box-warning { border-top-color: #F59E0B; }
.value-box > .inner {
  padding: 13px 18px !important;
}
.value-box .value {
  color: #111827 !important;
  font-size: 30px !important;
  line-height: 1.05;
}
.value-box .caption {
  color: #374151 !important;
  font-size: 13px !important;
  line-height: 1.25;
}
.value-box .icon i {
  top: 15px !important;
  right: 16px !important;
  font-size: 52px !important;
  color: rgba(17, 24, 39, 0.09) !important;
}
.insight-banner {
  margin: 8px 8px 0 0;
  padding: 12px 18px;
  border-left: 4px solid #3B82F6;
  border-radius: 8px;
  background: #EFF6FF;
  color: #1E3A5F;
  font-size: 15px;
}
.synthetic-pill {
  display: inline-block;
  margin-right: 8px;
  padding: 2px 8px;
  border: 1px solid #F59E0B;
  border-radius: 999px;
  background: #FFF7E6;
  color: #92400E;
  font-size: 12px;
  font-weight: 600;
}
table {
  font-variant-numeric: tabular-nums;
}
th {
  background: #F9FAFB;
}
.chart-stage img {
  max-width: 100% !important;
  height: auto !important;
}
@media (max-width: 767px) {
  body { padding-left: 0 !important; }
  html, body, #dashboard-container, .tab-content,
  .dashboard-page-wrapper, .dashboard-column-orientation,
  .dashboard-column, .chart-wrapper, .value-box {
    min-width: 0 !important;
    max-width: 100% !important;
    width: 100% !important;
    box-sizing: border-box !important;
  }
  .dashboard-row {
    min-width: 0 !important;
    flex-direction: column !important;
  }
  .chart-wrapper {
    min-width: 0 !important;
    width: calc(100% - 8px) !important;
  }
  .value-box {
    min-height: 92px !important;
    height: 92px !important;
  }
  .value-box .value { font-size: 27px !important; }
  .chart-stage { overflow-x: auto !important; }
  table { font-size: 12px; }
}
</style>

Overview {data-icon="fa-dashboard"}
=====================================

<div class="insight-banner">
<span class="synthetic-pill">Synthetic demonstration</span>
The rebuilt files now use real query-shaped exports: weekly Ways of Working,
daily Microsoft 365 credit consumption by service, daily GitHub AI credits, and
Consumption metadata joined through the activity-file crosswalk.
</div>

Row {data-height=120}
-------------------------------------

### People represented

```{r}
flexdashboard::valueBox(
  comma(overall$People), "People in the analysis",
  icon = "fa-users", color = "primary"
)
```

### Licensed population

```{r}
flexdashboard::valueBox(
  percent(overall$Licensed / overall$People, accuracy = 0.1),
  paste0(comma(overall$Licensed), " licensed people"),
  icon = "fa-id-card", color = "info"
)
```

### Credits per person per week

```{r}
flexdashboard::valueBox(
  format(round(overall$`Weekly credits`, 1), big.mark = ",", nsmall = 1),
  "M365 + GitHub credits / person / week",
  icon = "fa-calculator", color = "success"
)
```

### M365 credits per session

```{r}
flexdashboard::valueBox(
  format(round(overall$`M365 credits / session`, 2), nsmall = 2),
  "Credits / observed M365 session",
  icon = "fa-tachometer", color = "warning"
)
```

Row {data-height=510}
-------------------------------------

### Consumption baselines by function

```{r function-baseline, fig.height=4.4, fig.alt="Horizontal bars comparing average weekly Copilot credits and Microsoft 365 sessions per person across eight functions."}
baseline_long <- function_summary |>
  select(FunctionType, People, `Weekly credits`, `M365 sessions`) |>
  pivot_longer(c(`Weekly credits`, `M365 sessions`),
               names_to = "Measure", values_to = "Value") |>
  mutate(
    Measure = recode(
      Measure,
      `Weekly credits` = "Credits / person / week",
      `M365 sessions` = "M365 sessions / person / week"
    ),
    Label = if_else(
      Measure == "Credits / person / week",
      format(round(Value, 1), nsmall = 1),
      format(round(Value, 1), nsmall = 1)
    ),
    FunctionType = factor(FunctionType, levels = FUNCTION_ORDER)
  )

ggplot(baseline_long,
       aes(FunctionType, Value, fill = Measure)) +
  geom_col(position = position_dodge(width = 0.72), width = 0.66) +
  geom_text(aes(label = Label),
            position = position_dodge(width = 0.72),
            hjust = -0.10, size = 3) +
  coord_flip() +
  facet_wrap(~Measure, scales = "free_x") +
  scale_fill_manual(values = PAL[c(1, 3)], guide = "none") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18))) +
  theme_t() +
  labs(
    subtitle = paste0(
      "Person-level averages across the product window; groups below n = ",
      MIN_GROUP_N, " are suppressed"
    ),
    x = NULL, y = NULL,
    caption = CHART_CAPTION
  )
```

### Baseline table

```{r}
function_summary |>
  arrange(desc(`Weekly credits`)) |>
  transmute(
    Function = FunctionType,
    People,
    `Positive users` = `Positive users`,
    `Credits / person / week` = round(`Weekly credits`, 1),
    `M365 credits / person / week` = round(`M365 credits`, 1),
    `GitHub credits / person / week` = round(`GitHub credits`, 1),
    `M365 sessions / person / week` = round(`M365 sessions`, 1),
    `M365 credits / session` = round(`M365 credits / session`, 2)
  ) |>
  kable(
    caption = "Function baseline. Credits per session is calculated from observed Microsoft 365 sessions, not inferred from tokens."
  )
```

Row {data-height=335}
-------------------------------------

### Population-level highlights

**Consumption volume and cost intensity are separate.** Credit volume is total
resource use. M365 credits per session is a cost-intensity measure calculated
from real consumption fields.

**Service mix is now observable.** The Consumption query exposes ServiceName, so
the report can show whether credits are concentrated in Analyst, Cowork,
Researcher or WorkIQ rather than relying on a fabricated task-type dimension.

**Ways of Working comparisons use the aligned product window.** The person query
has 26 weeks of collaboration history, while the credit feeds have 13 weeks of
weekday daily records. Relationships in this report use only the product weeks.
Available-to-focus hours are calendar availability within configured working
hours, not a measure of coding flow or psychological focus.

### How to read cost intensity

**M365 credits per session** is the number of Microsoft 365 Copilot credits used
per observed Consumption-query session. It is a cost-intensity measure, not a
productivity or value measure. It does not establish that a group is more or
less efficient, because the report does not observe output quality, task
difficulty or business value.

Credit distribution {data-icon="fa-chart-area"}
=====================================

Row {data-height=470}
-------------------------------------

### Concentration of credit consumption across users

```{r credit-concentration, fig.height=4.4, fig.alt="Concentration curve showing the cumulative share of Copilot credits accounted for by the highest-credit share of users."}
credit_rank <- person_consumption |>
  filter(positive_consumer) |>
  arrange(desc(weekly_total_credits)) |>
  mutate(
    n = n(),
    `Share of users` = row_number() / n,
    `Share of credits` =
      cumsum(weekly_total_credits) / sum(weekly_total_credits)
  )

marker_x <- c(0.10, 0.25)
marker_y <- vapply(marker_x, function(p) {
  credit_rank$`Share of credits`[
    which.min(abs(credit_rank$`Share of users` - p))
  ]
}, numeric(1))
markers <- tibble(x = marker_x, y = marker_y,
                  lbl = percent(marker_y, accuracy = 1))

ggplot(credit_rank, aes(`Share of users`, `Share of credits`)) +
  geom_area(fill = PAL[3], alpha = 0.35) +
  geom_line(colour = PAL[1], linewidth = 1) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed",
              colour = "grey55") +
  geom_point(data = markers, aes(x, y), colour = PAL[6], size = 2.6,
             inherit.aes = FALSE) +
  geom_text(data = markers, aes(x, y, label = lbl), inherit.aes = FALSE,
            vjust = -1.0, hjust = -0.15, size = 3.3, colour = PAL[6]) +
  scale_x_continuous(labels = percent_format(accuracy = 1),
                     expand = expansion(mult = c(0.01, 0.06))) +
  scale_y_continuous(labels = percent_format(accuracy = 1),
                     expand = expansion(mult = c(0.01, 0.10))) +
  theme_t() +
  labs(
    title = "Cumulative share of credits by user rank",
    subtitle = paste0(
      "Users ranked from highest to lowest weekly credit volume; the dashed ",
      "line marks equal consumption"
    ),
    x = "Share of positive users (ranked highest to lowest)",
    y = "Cumulative share of credits",
    caption = CHART_CAPTION
  )
```

### Concentration bands

```{r}
band_levels <- c("Top 1%", "Next 4% (top 1-5%)", "Next 5% (top 5-10%)",
                 "Next 15% (top 10-25%)", "Next 25% (top 25-50%)",
                 "Bottom 50%")

band_table <- credit_rank |>
  mutate(
    Band = case_when(
      `Share of users` <= 0.01 ~ band_levels[1],
      `Share of users` <= 0.05 ~ band_levels[2],
      `Share of users` <= 0.10 ~ band_levels[3],
      `Share of users` <= 0.25 ~ band_levels[4],
      `Share of users` <= 0.50 ~ band_levels[5],
      TRUE ~ band_levels[6]
    ),
    Band = factor(Band, levels = band_levels)
  ) |>
  group_by(Band) |>
  summarise(
    People = n(),
    `Weekly credits (band total)` = sum(weekly_total_credits),
    .groups = "drop"
  ) |>
  arrange(Band) |>
  mutate(
    `Share of users` = percent(People / sum(People), accuracy = 1),
    `Share of credits` = percent(
      `Weekly credits (band total)` / sum(`Weekly credits (band total)`),
      accuracy = 1
    )
  ) |>
  select(Band, People, `Share of users`, `Share of credits`)

band_table |>
  kable(caption = paste0(
    "Users ranked by average weekly credits, positive users only (n = ",
    comma(sum(band_table$People)), ")."
  ))
```

Row {data-height=440}
-------------------------------------

### Shape of the credit-volume distribution

```{r credit-histogram, fig.height=4.1, fig.alt="Histogram of average weekly Copilot credits per positive user on a logarithmic scale, with median and mean reference lines."}
credit_stats <- person_consumption |>
  filter(positive_consumer, weekly_total_credits > 0)

med_credits <- median(credit_stats$weekly_total_credits)
mean_credits <- mean(credit_stats$weekly_total_credits)
ref_lines <- tibble(
  stat = c("Median", "Mean"),
  value = c(med_credits, mean_credits),
  colour = c(PAL[1], PAL[6])
)

ggplot(credit_stats, aes(weekly_total_credits)) +
  geom_histogram(bins = 40, fill = PAL[2], colour = "white",
                 linewidth = 0.2) +
  geom_vline(data = ref_lines, aes(xintercept = value, colour = stat),
             linetype = "dashed", linewidth = 0.7, show.legend = TRUE) +
  scale_colour_manual(values = c(Median = PAL[1], Mean = PAL[6])) +
  scale_x_log10(labels = label_number()) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.10))) +
  theme_t() +
  labs(
    title = "Distribution of weekly credits per user",
    subtitle = paste0(
      "Logarithmic x-axis; positive users only. The right tail shows whether ",
      "a small group accounts for a disproportionate share of credits"
    ),
    x = "Credits / user / week (log scale)", y = "Number of users",
    colour = NULL,
    caption = CHART_CAPTION
  )
```

### M365 and GitHub credit split

```{r credit-source-split, fig.height=4.1, fig.alt="Stacked bars showing the share of total credits from Microsoft 365 and GitHub by credit-volume group."}
source_split <- person_base |>
  group_by(CreditQuartile) |>
  summarise(
    People = n(),
    `Microsoft 365` = sum(total_m365_credits),
    GitHub = sum(total_github_credits),
    .groups = "drop"
  ) |>
  pivot_longer(c(`Microsoft 365`, GitHub),
               names_to = "Source", values_to = "Credits") |>
  group_by(CreditQuartile) |>
  mutate(Share = safe_divide(Credits, sum(Credits))) |>
  ungroup() |>
  filter(!is.na(Share))

ggplot(source_split, aes(CreditQuartile, Share, fill = Source)) +
  geom_col(width = 0.68) +
  geom_text(
    aes(label = ifelse(Share >= 0.06, percent(Share, accuracy = 1), "")),
    position = position_stack(vjust = 0.5),
    colour = "white", size = 3
  ) +
  coord_flip() +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  scale_fill_manual(values = c("Microsoft 365" = PAL[1], GitHub = PAL[7])) +
  theme_t() +
  labs(
    title = "Credit source mix by credit-volume group",
    subtitle = "GitHub credits are observed separately from Microsoft 365 Consumption-query credits",
    x = NULL, y = "Share of credits", fill = "Credit source"
  )
```

Credit relationships
=====================================

Row {data-height=470}
-------------------------------------

### Ways of Working metrics by credit-consumption group

```{r credit-heatmap, fig.height=4.4, fig.alt="Heatmap of standardised Ways of Working metric differences for five credit-consumption groups."}
credit_levels <- c("No consumption", "Q1 lowest", "Q2", "Q3", "Q4 highest")

scan <- expand_grid(
  Metric = WOW_METRICS,
  CreditQuartile = credit_levels
) |>
  rowwise() |>
  mutate(d = std_diff(
    person_wow[[Metric]],
    person_wow$CreditQuartile == CreditQuartile
  )) |>
  ungroup() |>
  filter(!is.na(d)) |>
  mutate(
    Metric = unname(metric_labels[Metric]),
    CreditQuartile = factor(CreditQuartile, levels = rev(credit_levels))
  )

ggplot(scan, aes(CreditQuartile, reorder(Metric, d), fill = d)) +
  geom_tile(colour = "white", linewidth = 0.6) +
  geom_text(aes(label = sprintf("%+.2f", d)), size = 3) +
  scale_fill_gradient2(
    low = PAL[6], mid = "#F5F2EA", high = PAL[1],
    midpoint = 0, limits = c(-1.3, 1.3), oob = squish
  ) +
  theme_t() +
  theme(legend.position = "right") +
  labs(
    title = "Standardised difference against everyone outside the group",
    subtitle = "Positive values indicate the group scores above the rest of the population on that measure",
    x = NULL, y = NULL, fill = "Std. difference"
  )
```

### Credit groups

```{r}
person_base |>
  group_by(CreditQuartile) |>
  summarise(
    People = n(),
    `Credits / person / week` = mean(weekly_total_credits),
    `M365 credits / session` =
      safe_divide(sum(total_m365_credits), sum(total_m365_sessions)),
    .groups = "drop"
  ) |>
  mutate(
    `Share of population` = percent(People / sum(People), accuracy = 0.1),
    `Credits / person / week` = round(`Credits / person / week`, 1),
    `M365 credits / session` = if_else(
      is.na(`M365 credits / session`),
      "N/A",
      format(round(`M365 credits / session`, 2), nsmall = 2)
    )
  ) |>
  kable(caption = "Credit-volume groups. Quartiles are calculated among positive users; non-users are retained separately.")
```

Row {data-height=455}
-------------------------------------

### Collaboration mode mix by credit-consumption group

```{r mode-mix, fig.height=4.1, fig.alt="One hundred percent stacked bars comparing email, chat, meeting and unscheduled-call shares across credit groups."}
modes <- c("Email_hours", "Chat_hours", "Meeting_hours",
           "Unscheduled_call_hours")

mode_share <- person_wow |>
  group_by(CreditQuartile) |>
  summarise(People = n(), across(all_of(modes), mean), .groups = "drop") |>
  pivot_longer(all_of(modes), names_to = "Mode", values_to = "Hours") |>
  group_by(CreditQuartile) |>
  mutate(Share = Hours / sum(Hours)) |>
  ungroup() |>
  mutate(Mode = recode(
    Mode,
    Email_hours = "Email",
    Chat_hours = "Chat",
    Meeting_hours = "Meetings",
    Unscheduled_call_hours = "Unscheduled calls"
  ))

ggplot(mode_share, aes(CreditQuartile, Share, fill = Mode)) +
  geom_col(width = 0.68) +
  geom_text(
    aes(label = ifelse(Share >= 0.06, percent(Share, accuracy = 1), "")),
    position = position_stack(vjust = 0.5),
    colour = ifelse(mode_share$Mode %in% c("Meetings", "Unscheduled calls"),
                    "#1F2937", "white"),
    size = 3
  ) +
  coord_flip() +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  scale_fill_manual(values = PAL[1:4]) +
  theme_t() +
  labs(
    title = "Share of collaboration hours by mode",
    subtitle = "Mode hours are shown before deduplication, so shares may not sum exactly across modes",
    x = NULL, y = NULL, fill = "Collaboration mode"
  )
```

### Workload, focus and network

```{r credit-outcomes, fig.height=4.1, fig.alt="Small bar charts comparing after-hours collaboration, available-to-focus hours, uninterrupted hours, interrupted hours and network metrics across credit groups."}
person_wow |>
  group_by(CreditQuartile) |>
  summarise(
    People = n(),
    `After-hours collaboration hours / person / week` =
      mean(After_hours_collaboration_hours),
    `Available-to-focus hours / person / week` =
      mean(Available_to_focus_hours),
    `Uninterrupted hours / person / week` = mean(Uninterrupted_hours),
    `Interrupted hours / person / week` = mean(Interrupted_hours),
    `Internal network size (people)` = mean(Internal_network_size),
    `External network size (people)` = mean(External_network_size),
    .groups = "drop"
  ) |>
  pivot_longer(-c(CreditQuartile, People),
               names_to = "Metric", values_to = "Value") |>
  ggplot(aes(CreditQuartile, Value)) +
  geom_col(fill = PAL[2], width = 0.66) +
  geom_text(aes(label = round(Value, 1)), vjust = -0.35, size = 2.8) +
  facet_wrap(~Metric, scales = "free_y", ncol = 2) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.16))) +
  theme_t() +
  theme(axis.text.x = element_text(angle = 25, hjust = 1)) +
  labs(
    title = "Workload, focus and network by credit-volume group",
    subtitle = "Network measures come directly from the weekly person query and are averaged, not summed",
    x = NULL, y = NULL
  )
```

Function drill-down {data-icon="fa-sitemap"}
=====================================

Row {data-height=505}
-------------------------------------

### Consumption and collaboration volume by function

```{r bubble-collaboration, fig.height=4.5, fig.alt="Bubble plot of weekly credits and collaboration hours by function. Bubble size represents positive users and colour represents Microsoft 365 credits per session."}
ggplot(
  function_summary,
  aes(`Weekly credits`, `Collaboration hours`,
      size = `Positive users`, colour = `M365 credits / session`,
      label = FunctionType)
) +
  geom_point(alpha = 0.85) +
  geom_text_repel(size = 3.2, colour = "grey20", seed = 2026,
                  min.segment.length = 0, max.overlaps = Inf,
                  box.padding = 0.6, point.padding = 0.3) +
  scale_size_area(max_size = 14) +
  scale_colour_gradient(low = PAL[3], high = PAL[6], na.value = PAL[5]) +
  scale_x_continuous(expand = expansion(mult = 0.18)) +
  scale_y_continuous(expand = expansion(mult = 0.18)) +
  theme_t() +
  labs(
    title = "Credit consumption against collaboration hours, by function",
    subtitle = "Bubble size shows positive users; colour shows Microsoft 365 credits per observed session",
    x = "Credits / user / week",
    y = "Collaboration hours / user / week",
    colour = "M365 credits / session",
    size = "Positive users",
    caption = CHART_CAPTION
  )
```

### Reading the chart

The upper-right functions combine high credit volume with high collaboration
volume. That does not imply that one causes the other. Their role mix,
seniority and working practices may raise both.

The colour adds a third dimension: cost per observed Microsoft 365 session.
That can indicate more credit-intensive sessions, but it should be interpreted
with service mix, entitlement and task context rather than treated as a
performance score.

Row {data-height=490}
-------------------------------------

### Workload and credit intensity by function

```{r bubble-workload, fig.height=4.4, fig.alt="Bubble plot of credits and after-hours collaboration by function. Bubble size represents positive users and colour represents Microsoft 365 credits per session."}
ggplot(
  function_summary,
  aes(`Weekly credits`, `After-hours hours`,
      size = `Positive users`, colour = `M365 credits / session`,
      label = FunctionType)
) +
  geom_point(alpha = 0.85) +
  geom_text_repel(size = 3.2, colour = "grey20", seed = 2026,
                  min.segment.length = 0, max.overlaps = Inf,
                  box.padding = 0.6, point.padding = 0.3) +
  scale_size_area(max_size = 14) +
  scale_colour_gradient(low = PAL[3], high = PAL[6], na.value = PAL[5]) +
  scale_x_continuous(expand = expansion(mult = 0.18)) +
  scale_y_continuous(expand = expansion(mult = 0.18)) +
  theme_t() +
  labs(
    title = "Credit consumption against after-hours workload, by function",
    subtitle = "Same bubble-size and colour encoding as the chart above, for comparison across functions",
    x = "Credits / user / week",
    y = "After-hours hours / user / week",
    colour = "M365 credits / session",
    size = "Positive users",
    caption = CHART_CAPTION
  )
```

### Function comparison matrix

```{r function-heatmap, fig.height=4.4, fig.alt="Indexed heatmap comparing consumption and Ways of Working measures across eight functions."}
heat_metrics <- c(
  "Weekly credits", "M365 credits / session", "Collaboration hours",
  "After-hours hours", "Available-to-focus hours", "Uninterrupted hours",
  "Interrupted hours", "Internal network size"
)

function_summary |>
  mutate(FunctionType = factor(FunctionType, levels = FUNCTION_ORDER)) |>
  select(FunctionType, all_of(heat_metrics)) |>
  pivot_longer(-FunctionType, names_to = "Metric", values_to = "Value") |>
  filter(!is.na(Value)) |>
  mutate(Metric = recode(
    Metric,
    `Weekly credits` = "Credits / user / week",
    `M365 credits / session` = "M365 credits / session",
    `Collaboration hours` = "Collaboration hours / person / week",
    `After-hours hours` = "After-hours hours / person / week",
    `Available-to-focus hours` =
      "Available-to-focus hours / person / week",
    `Uninterrupted hours` = "Uninterrupted hours / person / week",
    `Interrupted hours` = "Interrupted hours / person / week",
    `Internal network size` = "Internal network size (people)"
  )) |>
  group_by(Metric) |>
  mutate(Index = 100 * Value / mean(Value)) |>
  ungroup() |>
  ggplot(aes(FunctionType, Metric, fill = Index)) +
  geom_tile(colour = "white", linewidth = 0.7) +
  geom_text(aes(label = round(Index)), size = 3) +
  scale_fill_gradient2(
    low = "#E8EEF5", mid = "#F5F2EA", high = PAL[1], midpoint = 100
  ) +
  theme_t() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1),
        legend.position = "right") +
  labs(
    title = "Each metric indexed to the function average (100)",
    subtitle = "A compact view of consumption, focus and network measures across functions",
    x = NULL, y = NULL, fill = "Index"
  )
```

Consumption mix {data-icon="fa-adjust"}
=====================================

Row {data-height=480}
-------------------------------------

### Credit volume by cost-per-session profile

```{r cost-profile, fig.height=4.3, fig.alt="Grouped bars comparing weekly credit volume for high cost per session and standard cost per session profiles within each credit band."}
profile_summary <- person_wow |>
  filter(ConsumptionProfile != "No consumption") |>
  group_by(ConsumptionProfile) |>
  summarise(
    People = n(),
    `Credits / person / week` = mean(weekly_total_credits),
    `M365 credits / session` =
      safe_divide(sum(total_m365_credits), sum(total_m365_sessions)),
    `After-hours hours` = mean(After_hours_collaboration_hours),
    `Available-to-focus hours` = mean(Available_to_focus_hours),
    `Uninterrupted hours` = mean(Uninterrupted_hours),
    .groups = "drop"
  ) |>
  filter(People >= MIN_GROUP_N)

profile_grid <- profile_summary |>
  mutate(
    CreditBand = if_else(grepl("^High credit", ConsumptionProfile),
                         "High credit", "Lower credit"),
    `Cost-per-session profile` = if_else(
      grepl("high cost/session", ConsumptionProfile),
      "High cost per session", "Standard cost per session"
    )
  )

ggplot(
  profile_grid,
  aes(CreditBand, `Credits / person / week`,
      fill = `Cost-per-session profile`)
) +
  geom_col(position = position_dodge(width = 0.7), width = 0.6) +
  geom_text(
    aes(label = round(`Credits / person / week`, 1)),
    position = position_dodge(width = 0.7), vjust = -0.4, size = 3.2
  ) +
  scale_fill_manual(values = c(
    "High cost per session" = PAL[6],
    "Standard cost per session" = PAL[3]
  )) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18))) +
  theme_t() +
  labs(
    title = "Credit volume and M365 cost intensity",
    subtitle = "Cost intensity is based on observed Consumption-query sessions, not token mix",
    x = "Credit-volume band", y = "Credits / person / week",
    fill = "Cost-per-session profile"
  )
```

### Profile definitions

| Profile component | Definition | Interpretation |
|---|---|---|
| High credit | Upper quartile of average weekly total credits among positive users | Higher total credit volume |
| Lower credit | Other positive users | Lower total credit volume |
| High cost per session | Upper quartile of M365 credits per session among people with at least 10 observed M365 sessions | More credits used per observed M365 session |
| Standard cost per session | Below the high-cost threshold, with enough observed sessions | Lower observed M365 cost intensity |

The label **high cost per session** describes a measurable consumption pattern.
It is not a wellbeing, productivity or value judgement.

Row {data-height=455}
-------------------------------------

### Workload by cost-per-session profile within credit bands

```{r cost-outcomes, fig.height=4.1, fig.alt="Bars comparing after-hours collaboration, available-to-focus hours and uninterrupted hours across standard-cost and high-cost-per-session profiles."}
profile_summary |>
  select(ConsumptionProfile, People,
         `After-hours hours`, `Available-to-focus hours`,
         `Uninterrupted hours`) |>
  pivot_longer(-c(ConsumptionProfile, People),
               names_to = "Metric", values_to = "Value") |>
  ggplot(
    aes(reorder(ConsumptionProfile, Value), Value,
        fill = grepl("high cost/session", ConsumptionProfile))
  ) +
  geom_col(width = 0.65) +
  geom_text(aes(label = round(Value, 2)), hjust = -0.12, size = 3) +
  coord_flip() +
  facet_wrap(~Metric, scales = "free_x") +
  scale_fill_manual(
    values = c(`TRUE` = PAL[6], `FALSE` = PAL[2]),
    labels = c(`TRUE` = "High cost per session",
               `FALSE` = "Standard cost per session")
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18))) +
  theme_t() +
  labs(
    title = "Workload by M365 cost intensity, within credit bands",
    subtitle = "Compared within the same broad credit-volume band",
    x = NULL, y = NULL, fill = "Cost-per-session profile"
  )
```

### Service mix

```{r service-mix, fig.height=4.1, fig.alt="Stacked bars comparing the share of Microsoft 365 credits used by service for high-credit and lower-credit users."}
m365_service_weekly |>
  left_join(select(person_base, PersonId, CreditBand),
            by = "PersonId") |>
  filter(CreditBand != "No consumption") |>
  group_by(CreditBand, ServiceName) |>
  summarise(Credits = sum(m365_credits), People = n_distinct(PersonId),
            .groups = "drop") |>
  filter(People >= MIN_GROUP_N) |>
  group_by(CreditBand) |>
  mutate(Share = Credits / sum(Credits)) |>
  ungroup() |>
  ggplot(aes(CreditBand, Share, fill = reorder(ServiceName, Share))) +
  geom_col(width = 0.68) +
  geom_text(
    aes(label = ifelse(Share >= 0.06, percent(Share, accuracy = 1), "")),
    position = position_stack(vjust = 0.5),
    colour = "grey15", size = 2.9
  ) +
  coord_flip() +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  scale_fill_brewer(palette = "Blues", direction = -1) +
  theme_t() +
  labs(
    title = "Microsoft 365 credit mix by service",
    subtitle = "ServiceName is a real Consumption-query dimension",
    x = NULL, y = "Share of M365 credits", fill = "Service"
  )
```

Usage segments {data-icon="fa-users"}
=====================================

Row {data-height=480}
-------------------------------------

### Credit-volume distribution within activity segments

```{r segment-credit, fig.height=4.3, fig.alt="Stacked bars showing credit-consumption quartiles within Power, Habitual, Novice, Low and Non-user credit activity segments."}
segment_credit <- person_base |>
  count(CreditActivitySegment, CreditQuartile, name = "People") |>
  group_by(CreditActivitySegment) |>
  mutate(Share = People / sum(People)) |>
  ungroup()

ggplot(segment_credit, aes(CreditActivitySegment, Share,
                           fill = CreditQuartile)) +
  geom_col(width = 0.68) +
  coord_flip() +
  scale_x_discrete(limits = rev(levels(person_base$CreditActivitySegment))) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  scale_fill_manual(values = PAL[c(5, 4, 3, 2, 1)]) +
  theme_t() +
  labs(
    title = "Credit-volume distribution within credit activity segments",
    subtitle = paste0(
      "Segments are based on active product weeks and observed total credits ",
      "from ", format(PRODUCT_WINDOW_START, "%d %b %Y"), " to ",
      format(PRODUCT_WINDOW_END, "%d %b %Y")
    ),
    x = NULL, y = NULL, fill = "Credit-volume group"
  )
```

### Why replace standard usage segments?

The `vivainsights::identify_usage_segments()` helper expects a sustained
activity metric such as Copilot actions. This export set does not contain a
Copilot actions column in either the Consumption query or the selected person
query. The report therefore uses a transparent credit activity segmentation:
active product weeks plus credit volume. It does not fabricate an actions field.

Row {data-height=400}
-------------------------------------

### Segment baseline

```{r}
person_base |>
  group_by(CreditActivitySegment) |>
  summarise(
    People = n(),
    `Active weeks` = mean(active_weeks),
    `Credits / person / week` = mean(weekly_total_credits),
    `M365 credits / person / week` = mean(weekly_m365_credits),
    `GitHub credits / person / week` = mean(weekly_github_credits),
    `M365 credits / session` =
      safe_divide(sum(total_m365_credits), sum(total_m365_sessions)),
    .groups = "drop"
  ) |>
  arrange(CreditActivitySegment) |>
  mutate(
    across(c(`Active weeks`, `Credits / person / week`,
             `M365 credits / person / week`,
             `GitHub credits / person / week`),
           \(x) round(x, 2)),
    `M365 credits / session` = if_else(
      is.na(`M365 credits / session`),
      "N/A",
      format(round(`M365 credits / session`, 2), nsmall = 2)
    )
  ) |>
  kable(caption = paste0(
    "Segment baseline across ", length(product_weeks),
    " product weeks. Non-users are people with no observed M365 or GitHub credits."
  ))
```

### Interpretation guardrail

Neither activity segment nor credit group is a performance category. They
describe observed use in the product window. Results should be reported for
privacy-safe cohorts and interpreted alongside role, entitlement, service mix
and organisational context.

Methods and appendix {data-icon="fa-book"}
=====================================

Row {data-height=410}
-------------------------------------

### Join and aggregation workflow

```{r workflow, fig.height=3.4}
steps <- tibble(
  x = c(1, 3, 5, 7, 9),
  label = c(
    "Consumption query\nPerson x service x day",
    "Aggregate services\nbefore person-week rates",
    "Person Query\nPerson x week",
    "Join credits by\nPersonId and week",
    "Population,\nservice and function views"
  )
)

ggplot(steps, aes(x, 1)) +
  geom_segment(
    aes(xend = x + 1.35, yend = 1),
    arrow = grid::arrow(length = grid::unit(0.14, "in")),
    colour = PAL[2], linewidth = 0.8
  ) +
  geom_label(
    aes(label = label), fill = "#EAF2F8", colour = PAL[1],
    label.size = 0.25, size = 3.1, lineheight = 0.95
  ) +
  coord_cartesian(xlim = c(0, 10.6), ylim = c(0.6, 1.4), clip = "off") +
  theme_void() +
  labs(title = "Time-aligned person-week panel")
```

### Core rules

| Question | Rule |
|---|---|
| Identity | Take the `PersonId` / `PeopleHistoricalId` crosswalk from Consumption activity files; never derive opaque identifiers by string manipulation |
| Metadata | Join Consumption activity to `PeopleMetaData` with `PeopleHistoricalId`; use person-query attributes for whole-population grouping because a person query alone has no `PeopleHistoricalId` |
| Weekly relationships | Aggregate daily product records to person-week, then join to person-query weeks by `PersonId` and `MetricDate` |
| Service grain | Aggregate across `ServiceName` before computing person-week M365 credits or credits per session |
| Credit quartiles | Calculate among positive users and retain non-users separately |
| Cost intensity | Use M365 credits per observed M365 session; do not infer token mix |
| Focus identity | Check that meetings, scheduled calls and available-to-focus hours reconcile to the configured working-hours basis, and that uninterrupted plus interrupted hours reconcile to available-to-focus hours |
| Network measures | Use weekly person-query network metrics and average them across weeks; never sum them |
| Privacy | Suppress organisational groups below the approved minimum population |

### Key crosswalk check

```{r}
unmatched_key_counts |>
  kable(caption = paste0(
    "Identity checks report counts only. The crosswalk has ",
    comma(nrow(person_keys)),
    " distinct PersonId / PeopleHistoricalId pairs observed in credit files."
  ))
```

Row {data-height=390}
-------------------------------------

### Simulation design

This dashboard is a demonstration of analytical possibilities. Every person,
credit, session and Ways of Working value is synthetic.

The current source set represents `r comma(population_counts$Value[population_counts$Metric == "People"])`
people, including `r comma(population_counts$Value[population_counts$Metric == "Developers"])`
developers, across `r population_counts$Value[population_counts$Metric == "Organisations"]`
organisations and `r population_counts$Value[population_counts$Metric == "Functions"]`
functions. The person query spans 26 Sunday-start weeks. The M365 and GitHub
credit feeds cover the final `r length(product_weeks)` weeks at weekday daily
grain, which is why this report filters Ways of Working measures to the product
window before relating them to credits.

Available-to-focus hours are simulated on a 40-hour weekly basis here, but that
basis is configurable per tenant through metric rules. The metric describes
calendar time left after meetings and scheduled Teams calls, and should not be
read as coding time or psychological flow.

The following effects are deliberately modest:

- Higher consumption is associated with more chat and less email as a share of
  collaboration modes.
- Total collaboration volume is raised mainly by organisational composition.
- Internal network size has a small residual association after composition is
  considered.
- Cost per M365 session has a weak association with after-hours collaboration
  in some credit bands.
- External network size is a null.

### Synthetic source assets

The uncompressed source-shaped CSVs sit under:

```text
examples/utility-r/_data/
  person-query/PersonQuery.csv
  consumption-query/PeopleMetaData.csv
  consumption-query/PersonM365CreditsMetrics.csv
  consumption-query/PersonGitHubCreditsMetrics.csv
```

They are intended for demonstrations, workshops and adaptation of the template.
They contain no customer data or real identifiers. The GitHub-query files in the
same folder are not required for this consumption report.

Row {data-height=430}
-------------------------------------

### Definitions and limitations

**Credits per session** means Microsoft 365 Copilot credits consumed divided by
observed Microsoft 365 Consumption-query sessions. It is preferred over any
token-normalised measure because token fields are not present in this export.

**Available-to-focus hours** means working hours remaining after excluding
meetings and scheduled Teams calls. `Uninterrupted_hours` and
`Interrupted_hours` partition available-to-focus time in this source. The
working-hours basis is tenant-configurable; the 40-hour identity checked here is
specific to this synthetic dataset.

**Credit concentration** ranks people by average weekly total credits and shows
what share of credits is accounted for by the highest-consuming users.

**Credit activity segments** are based on active product weeks and observed
credit volume. They are not the standard `vivainsights` action-based segments,
because this source set does not include a Copilot actions metric.

**PeopleHistoricalId** is the documented join key between the Consumption
activity files and `PeopleMetaData`. It must be taken from the Consumption
export rather than derived from `PersonId`. A person-query export alone cannot
be joined to `PeopleMetaData`; the bridge is an activity file that carries both
`PersonId` and `PeopleHistoricalId`.

The dashboard shows associations. It does not establish that AI consumption
caused a working pattern, and it does not observe output quality, task
difficulty or business value. It also makes no wellbeing claims: uninterrupted,
interrupted, available-to-focus, after-hours and weekend measures describe
observed calendar and collaboration patterns only.

### Replacing the synthetic data

1. Replace the four CSVs listed above with schema-aligned governed exports.
2. Validate the identity crosswalk and disclose the match rate.
3. Confirm whether missing consumption rows mean zero activity, suppression or
   ineligibility before filling absence with zero.
4. Preserve the daily service grain until after service-level totals are needed;
   aggregate services before computing person-week rates.
5. Keep person-level averaging so each person carries equal analytical weight.
6. Retain the privacy threshold and the organisational composition checks.
