---
title: "Copilot Consumption and Ways of Working"
subtitle: "The art of the possible with Consumption, GitHub credits and Viva Insights data (fully synthetic)"
date: "Synthetic window: 4 Jan-28 Jun 2026"
output:
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: scroll
    theme: lumen
    self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.width = 8.5, fig.height = 4.5)
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
library(knitr)
library(ggrepel)
super_panel_helper <- "copilot-super-panel-helpers.R"
if (!file.exists(super_panel_helper) && exists("utility")) {
  super_panel_helper <- file.path(utility, "copilot-super-panel-helpers.R")
}
source(super_panel_helper)

MIN_GROUP_N <- 10L
MIN_M365_SESSIONS <- 10L
PAL <- c("#1F4E79", "#2E75B6", "#7FA8D0", "#BDD3E8",
         "#C9C9C9", "#B03A2E", "#5B8C5A", "#7D5BA6")

theme_t <- function() {
  theme_minimal(base_size = 11) +
    theme(
      panel.grid.minor = element_blank(),
      legend.position = "bottom",
      plot.title.position = "plot",
      plot.title = element_text(face = "bold", colour = PAL[1]),
      plot.subtitle = element_text(colour = "grey30",
                                   margin = margin(t = 2, b = 9)),
      plot.caption = element_text(colour = "grey45", size = 8, hjust = 0,
                                  margin = margin(t = 7)),
      plot.margin = margin(t = 10, r = 16, b = 4, l = 4)
    )
}

# Release complete partitions only: dropping one small cell still exposes it
# through the population total, another panel, or a percentage denominator.
publication_partition <- function(data, groups, contributors = character(),
                                  min_n = MIN_GROUP_N) {
  stopifnot(length(min_n) == 1L, is.finite(min_n), min_n >= 10L)
  cells <- data |>
    group_by(across(all_of(groups))) |>
    summarise(People = n_distinct(PersonId), .groups = "drop")
  safe <- nrow(cells) > 0L && all(cells$People >= min_n) &&
    !anyNA(data$PersonId) && all(nzchar(as.character(data$PersonId)))
  for (metric in contributors) {
    support <- data |>
      group_by(across(all_of(groups))) |>
      summarise(
        People = n_distinct(PersonId),
        Positive = n_distinct(PersonId[!is.na(.data[[metric]]) &
                                         .data[[metric]] > 0]),
        .groups = "drop"
      )
    safe <- safe && all(is.finite(data[[metric]])) &&
      all(support$Positive == 0L | support$Positive >= min_n) &&
      all(support$People - support$Positive == 0L |
            support$People - support$Positive >= min_n)
  }
  if (safe) data else data[0, , drop = FALSE]
}

publication_count <- function(n, total, min_n = MIN_GROUP_N) {
  stopifnot(length(min_n) == 1L, is.finite(min_n), min_n >= 10L)
  ifelse(total >= min_n & (n == 0 | n >= min_n) &
           (total - n == 0 | total - n >= min_n), n, NA_real_)
}

consumption_profile <- function(credit_band, sessions, cost_profile) {
  case_when(
    is.na(sessions) | sessions < MIN_M365_SESSIONS ~
      "Limited M365 sessions",
    credit_band == "High credit" &
      cost_profile == "High cost per session" ~
        "High credit, high cost/session",
    credit_band == "High credit" ~ "High credit, standard cost/session",
    credit_band == "Lower credit" &
      cost_profile == "High cost per session" ~
        "Lower credit, high cost/session",
    credit_band == "Lower credit" ~ "Lower credit, standard cost/session",
    TRUE ~ "No observed M365 credits"
  )
}

super <- build_copilot_super_panel("_data")
pq <- super$panel
people_meta <- super$people_meta
m365_daily_raw <- super$m365_daily_raw
m365_weekly <- super$m365_weekly
m365_service_weekly <- super$m365_service_weekly
github_daily_raw <- super$github_credits_daily
github_weekly <- super$github_credits_weekly
github_activity_weekly <- super$github_activity_weekly

stopifnot(
  max(abs(pq$Meeting_hours + pq$Scheduled_call_hours +
            pq$Available_to_focus_hours - 40), na.rm = TRUE) <= 0.01,
  max(abs(pq$Uninterrupted_hours + pq$Interrupted_hours -
            pq$Available_to_focus_hours), na.rm = TRUE) <= 0.01,
  all(pq$Uninterrupted_hours <= pq$Available_to_focus_hours + 0.01,
      na.rm = TRUE),
  all(pq$Open_1_hour_block <= pq$Available_to_focus_hours + 0.01,
      na.rm = TRUE)
)

person_keys <- super$person_keys
credit_key_rows <- bind_rows(
  select(m365_daily_raw, PeopleHistoricalId),
  select(github_daily_raw, PeopleHistoricalId)
)

unmatched_key_counts <- tibble(
  Check = c(
    "PeopleMetaData people with no activity-file crosswalk",
    "Credit activity rows with PeopleHistoricalId absent from PeopleMetaData"
  ),
  Count = c(
    anti_join(people_meta, person_keys, by = "PeopleHistoricalId") |> nrow(),
    anti_join(credit_key_rows, people_meta, by = "PeopleHistoricalId") |> nrow()
  )
)

PRODUCT_WINDOW_START <- min(c(m365_weekly$MetricDate, github_weekly$MetricDate,
                              github_activity_weekly$MetricDate), na.rm = TRUE)
PRODUCT_WINDOW_END <- max(c(m365_weekly$MetricDate, github_weekly$MetricDate,
                            github_activity_weekly$MetricDate), na.rm = TRUE)

product_weeks <- sort(unique(c(m365_weekly$MetricDate, github_weekly$MetricDate,
                               github_activity_weekly$MetricDate)))

person_week <- pq |>
  filter(MetricDate %in% product_weeks) |>
  select(PersonId, MetricDate, Collaboration_hours,
         Meeting_hours, Email_hours, Chat_hours, Scheduled_call_hours,
         Unscheduled_call_hours, Channel_message_hours,
         Active_connected_hours, External_collaboration_hours,
         Available_to_focus_hours, Uninterrupted_hours, Interrupted_hours,
         Open_1_hour_block,
         Recurring_meeting_hours, Conflicting_meeting_hours,
         Meeting_hours_with_six_or_fewer_hours_of_advanced_notice,
         Emails_sent, Chats_sent, Meetings, Calls,
         After_hours_collaboration_hours, After_hours_meeting_hours,
         After_hours_email_hours, After_hours_chat_hours,
         Weekend_collaboration_hours, Collaboration_span,
         Internal_network_size, External_network_size, Strong_ties,
         Diverse_ties, Network_outside_organization,
         Total_Copilot_enabled_days, m365_sessions, m365_credits,
         m365_observed_days, m365_active_days, m365_service_count,
         github_credits, github_credit_days, github_observed_days,
         github_active_days, github_code_suggestions, github_code_acceptances,
         github_chat_requests, github_agent_adoption_days,
         github_feature_usage_count, github_distinct_features,
         github_distinct_languages_feature, github_distinct_models_language,
         m365_observed_use, github_observed_use, any_observed_product_use,
         github_activity_observed, m365_credit_row_observed,
         credits_per_m365_session)

WOW_METRICS <- c(
  "Collaboration_hours", "Active_connected_hours",
  "Email_hours", "Chat_hours", "Meeting_hours",
  "Unscheduled_call_hours", "After_hours_collaboration_hours",
  "After_hours_meeting_hours", "After_hours_email_hours",
  "After_hours_chat_hours",
  "Collaboration_span", "Weekend_collaboration_hours",
  "Available_to_focus_hours", "Uninterrupted_hours", "Interrupted_hours",
  "Internal_network_size", "External_network_size", "Strong_ties",
  "Diverse_ties", "Network_outside_organization"
)

metric_labels <- c(
  Collaboration_hours = "Collaboration hours / person / week",
  Active_connected_hours = "Active connected hours / person / week",
  Email_hours = "Email hours / person / week",
  Chat_hours = "Chat hours / person / week",
  Meeting_hours = "Meeting hours / person / week",
  Unscheduled_call_hours = "Unscheduled call hours / person / week",
  After_hours_collaboration_hours =
    "After-hours collaboration hours / person / week",
  After_hours_meeting_hours = "After-hours meeting hours / person / week",
  After_hours_email_hours = "After-hours email hours / person / week",
  After_hours_chat_hours = "After-hours chat hours / person / week",
  Collaboration_span = "Collaboration span (hours)",
  Weekend_collaboration_hours =
    "Weekend collaboration hours / person / week",
  Available_to_focus_hours = "Available-to-focus hours / person / week",
  Uninterrupted_hours = "Uninterrupted hours / person / week",
  Interrupted_hours = "Interrupted hours / person / week",
  Internal_network_size = "Internal network size (people)",
  External_network_size = "External network size (people)",
  Strong_ties = "Strong ties",
  Diverse_ties = "Diverse ties",
  Network_outside_organization = "Network outside organization"
)

person_consumption <- person_week |>
  group_by(PersonId) |>
  summarise(
    # Product-specific weekly rates use product-specific observed denominators,
    # so a GitHub-only observed week never dilutes an M365 average.
    m365_observed_weeks = sum(!is.na(m365_observed_days)),
    github_observed_weeks = sum(!is.na(github_observed_days) |
                                  !is.na(github_credit_days)),
    weekly_m365_credits = if_else(
      m365_observed_weeks > 0,
      sum(m365_credits, na.rm = TRUE) / m365_observed_weeks,
      0
    ),
    weekly_github_credits = if_else(
      github_observed_weeks > 0,
      sum(github_credits, na.rm = TRUE) / github_observed_weeks,
      0
    ),
    weekly_m365_sessions = if_else(
      m365_observed_weeks > 0,
      sum(m365_sessions, na.rm = TRUE) / m365_observed_weeks,
      0
    ),
    active_weeks = sum(m365_credits > 0 | m365_sessions > 0, na.rm = TRUE),
    total_m365_credits = sum(m365_credits, na.rm = TRUE),
    total_github_credits = sum(github_credits, na.rm = TRUE),
    total_m365_sessions = sum(m365_sessions, na.rm = TRUE),
    total_m365_active_days = sum(m365_active_days, na.rm = TRUE),
    total_github_active_days = sum(github_active_days, na.rm = TRUE),
    total_github_code_suggestions = sum(github_code_suggestions, na.rm = TRUE),
    total_github_code_acceptances = sum(github_code_acceptances, na.rm = TRUE),
    total_github_chat_requests = sum(github_chat_requests, na.rm = TRUE),
    total_github_agent_adoption_days =
      sum(github_agent_adoption_days, na.rm = TRUE),
    product_weeks = n(),
    credits_per_m365_session =
      safe_divide(sum(m365_credits, na.rm = TRUE), sum(m365_sessions, na.rm = TRUE)),
    .groups = "drop"
  ) |>
  mutate(
    positive_consumer = total_m365_credits > 0 | total_m365_sessions > 0,
    github_positive_consumer = total_github_credits > 0 |
      total_github_active_days > 0 | total_github_code_acceptances > 0 |
      total_github_chat_requests > 0,
    active_week_share = active_weeks / product_weeks,
    positive_quartile = if_else(
      positive_consumer,
      ntile(if_else(positive_consumer, weekly_m365_credits, NA_real_), 4L),
      NA_integer_
    ),
    CreditQuartile = factor(
      case_when(
        !positive_consumer ~ "No observed M365 credits",
        positive_quartile == 1L ~ "Q1 lowest",
        positive_quartile == 2L ~ "Q2",
        positive_quartile == 3L ~ "Q3",
        TRUE ~ "Q4 highest"
      ),
      levels = c("No observed M365 credits", "Q1 lowest", "Q2", "Q3", "Q4 highest")
    )
  )

high_cost_cut <- quantile(
  person_consumption$credits_per_m365_session[
    person_consumption$total_m365_sessions >= MIN_M365_SESSIONS
  ],
  0.75, na.rm = TRUE
)

person_attributes <- super$person_attributes

person_base <- person_attributes |>
  left_join(person_keys, by = "PersonId") |>
  left_join(person_consumption, by = "PersonId") |>
  mutate(
    across(c(weekly_m365_credits,
             weekly_github_credits, weekly_m365_sessions,
             total_m365_credits, total_github_credits,
             total_m365_sessions, total_m365_active_days,
             total_github_active_days, total_github_code_suggestions,
             total_github_code_acceptances, total_github_chat_requests,
             total_github_agent_adoption_days, active_weeks, active_week_share,
             m365_observed_weeks, github_observed_weeks),
           \(x) replace_na(x, 0)),
    positive_consumer = replace_na(positive_consumer, FALSE),
    github_positive_consumer = replace_na(github_positive_consumer, FALSE),
    ProductUsageMix = factor(
      case_when(
        positive_consumer & github_positive_consumer ~ "Observed both",
        positive_consumer ~ "Observed M365 only",
        github_positive_consumer ~ "Observed GitHub only",
        TRUE ~ "No observed product use"
      ),
      levels = c("Observed both", "Observed M365 only",
                 "Observed GitHub only", "No observed product use")
    ),
    CreditQuartile = replace_na(as.character(CreditQuartile),
                                "No observed M365 credits"),
    CreditQuartile = factor(
      CreditQuartile,
      levels = c("No observed M365 credits", "Q1 lowest", "Q2", "Q3", "Q4 highest")
    ),
    CreditBand = case_when(
      CreditQuartile == "Q4 highest" ~ "High credit",
      positive_consumer ~ "Lower credit",
      TRUE ~ "No observed M365 credits"
    ),
    CostPerSessionProfile = case_when(
      total_m365_sessions < MIN_M365_SESSIONS ~ "Limited M365 sessions",
      !positive_consumer ~ "No observed M365 credits",
      credits_per_m365_session >= high_cost_cut ~ "High cost per session",
      TRUE ~ "Standard cost per session"
    ),
    ConsumptionProfile = consumption_profile(
      CreditBand, total_m365_sessions, CostPerSessionProfile
    ),
    # The same condition that labels CostPerSessionProfile. Absent credit rows
    # are unresolved coverage, not established non-use, so the segment records
    # what was observed rather than asserting that the person used nothing.
    CreditActivitySegment = case_when(
      !positive_consumer ~ "No observed M365 credits",
      active_weeks >= 10 & CreditBand == "High credit" ~
        "Power User",
      active_weeks >= 8 ~ "Habitual User",
      active_weeks >= 3 ~ "Novice User",
      TRUE ~ "Low User"
    ),
    CreditActivitySegment = factor(
      CreditActivitySegment,
      levels = c("Power User", "Habitual User", "Novice User",
                 "Low User", "No observed M365 credits"),
      ordered = TRUE
    ),
    M365EnabledUser = replace_na(M365EnabledUser, FALSE),
    IsCopilotLicensed = if_else(M365EnabledUser, "true", "false")
  )

person_wow <- person_week |>
  group_by(PersonId) |>
  summarise(across(all_of(WOW_METRICS), \(x) mean(x, na.rm = TRUE)),
            .groups = "drop") |>
  left_join(person_base, by = "PersonId")

if (n_distinct(person_base$PersonId) < MIN_GROUP_N) {
  stop("Publication withheld: population does not meet the minimum group size.")
}

m365_contributors <- c("total_m365_credits", "total_m365_sessions")
pooled_functions <- person_wow |>
  group_by(FunctionType) |>
  summarise(
    People = n_distinct(PersonId),
    CreditPeople = n_distinct(PersonId[total_m365_credits > 0]),
    SessionPeople = n_distinct(PersonId[total_m365_sessions > 0]),
    .groups = "drop"
  ) |>
  filter(People < MIN_GROUP_N |
           is.na(publication_count(CreditPeople, People)) |
           is.na(publication_count(SessionPeople, People))) |>
  pull(FunctionType)
function_people <- person_wow |>
  mutate(FunctionType = if_else(FunctionType %in% pooled_functions,
                                "Other functions (pooled)", FunctionType))
function_public <- publication_partition(
  function_people, "FunctionType", m365_contributors
)
# Pool unsafe function cells first, then recheck the complete partition;
# a still-small pooled remainder withholds every function output.
credit_public <- publication_partition(
  person_wow, "CreditQuartile", m365_contributors
)
# Both segment outputs use the same joint gate; a baseline must not restore
# the denominator of a suppressed segment-by-quartile composition.
segment_public <- publication_partition(
  person_wow, c("CreditActivitySegment", "CreditQuartile"),
  m365_contributors
)
profile_public <- publication_partition(
  person_wow, c("ConsumptionProfile", "CreditQuartile"), m365_contributors
)
product_mix_public <- publication_partition(
  person_wow, "ProductUsageMix",
  c("total_m365_credits", "total_github_credits")
)
github_public <- publication_partition(
  person_base, character(), "total_github_credits"
)
overall_public <- publication_partition(
  person_base, character(), m365_contributors
)

function_summary <- function_public |>
  group_by(FunctionType) |>
  summarise(
    People = n(),
    `Weekly credits` = mean(weekly_m365_credits),
    `M365 credits` = mean(weekly_m365_credits),
    `M365 sessions` = mean(weekly_m365_sessions),
    `M365 credits / session` =
      safe_divide(sum(total_m365_credits), sum(total_m365_sessions)),
    `Collaboration hours` = mean(Collaboration_hours),
    `After-hours hours` = mean(After_hours_collaboration_hours),
    `Collaboration span` = mean(Collaboration_span),
    `Available-to-focus hours` = mean(Available_to_focus_hours),
    `Uninterrupted hours` = mean(Uninterrupted_hours),
    `Interrupted hours` = mean(Interrupted_hours),
    `Internal network size` = mean(Internal_network_size),
    .groups = "drop"
  )

FUNCTION_ORDER <- function_summary |>
  arrange(`Weekly credits`) |>
  pull(FunctionType)

overall <- person_base |>
  summarise(
    People = n(),
    Developers = publication_count(sum(IsDeveloper == "Yes"), n()),
    Licensed = publication_count(sum(M365EnabledUser), n()),
    Consumers = publication_count(sum(positive_consumer), n()),
    GitHubConsumers = publication_count(sum(github_positive_consumer), n()),
    `Weekly credits` = if (nrow(overall_public)) mean(weekly_m365_credits)
      else NA_real_,
    `M365 credits / session` = if (nrow(overall_public))
      safe_divide(sum(total_m365_credits), sum(total_m365_sessions))
      else NA_real_
  )

usage_mix_summary <- product_mix_public |>
  group_by(ProductUsageMix) |>
  summarise(
    People = n(),
    `M365 enabled users` = sum(M365EnabledUser),
    `M365 credits / person / week` = mean(weekly_m365_credits),
    `GitHub credits / person / week` = mean(weekly_github_credits),
    `M365 sessions / person / week` = mean(weekly_m365_sessions),
    `GitHub active days / person` = mean(total_github_active_days),
    .groups = "drop"
  ) |>
  mutate(`Share of population` = People / sum(People))

profile_summary <- profile_public |>
  group_by(ConsumptionProfile) |>
  summarise(
    People = n(),
    `Credits / person / week` = mean(weekly_m365_credits),
    `M365 credits / session` =
      if (isTRUE(first(ConsumptionProfile) == "Limited M365 sessions")) NA_real_
      else safe_divide(sum(total_m365_credits), sum(total_m365_sessions)),
    `After-hours hours` = mean(After_hours_collaboration_hours),
    `Available-to-focus hours` = mean(Available_to_focus_hours),
    `Uninterrupted hours` = mean(Uninterrupted_hours),
    .groups = "drop"
  )

profile_grid <- profile_summary |>
  mutate(
    CreditBand = case_when(
      ConsumptionProfile == "Limited M365 sessions" ~ "Limited M365 sessions",
      grepl("^High credit", ConsumptionProfile) ~ "High credit",
      grepl("^Lower credit", ConsumptionProfile) ~ "Lower credit",
      TRUE ~ "No observed M365 credits"
    ),
    `Cost-per-session profile` = case_when(
      ConsumptionProfile == "Limited M365 sessions" ~ "Limited M365 sessions",
      grepl("high cost/session", ConsumptionProfile) ~ "High cost per session",
      grepl("standard cost/session", ConsumptionProfile) ~ "Standard cost per session",
      TRUE ~ "No observed M365 credits"
    )
  )

CHART_CAPTION <- sprintf(
  paste0(
    "Synthetic data | n = %s people | ",
    "product window %s to %s"
  ),
  comma(overall$People),
  format(PRODUCT_WINDOW_START, "%d %b %Y"),
  format(PRODUCT_WINDOW_END, "%d %b %Y")
)

std_diff <- function(x, grp) {
  keep <- !is.na(x) & !is.na(grp)
  x <- x[keep]
  grp <- grp[keep]
  a <- x[grp]
  b <- x[!grp]
  if (length(a) < MIN_GROUP_N || length(b) < MIN_GROUP_N) return(NA_real_)
  pooled <- sqrt(((length(a) - 1) * var(a) + (length(b) - 1) * var(b)) /
                   (length(a) + length(b) - 2))
  if (is.na(pooled) || pooled == 0) return(NA_real_)
  (mean(a) - mean(b)) / pooled
}
```

<style>
body {
  font-family: "Segoe UI Variable", "Segoe UI", system-ui, sans-serif;
  color: #374151;
  background: #F4F4F4;
}
.navbar {
  min-height: 52px;
}
.navbar-brand {
  font-weight: 600;
}
.chart-wrapper {
  border: 1px solid #E5E7EB;
  border-radius: 8px;
  box-shadow: none;
}
.chart-title {
  color: #374151;
  font-weight: 500;
}
.value-box {
  min-height: 108px !important;
  height: 108px !important;
  color: #111827 !important;
  background: #FFFFFF !important;
  border: 1px solid #E5E7EB;
  border-top: 4px solid #3B82F6;
  border-radius: 8px;
  box-shadow: none;
}
.value-box-info { border-top-color: #14B8A6; }
.value-box-success { border-top-color: #8B5CF6; }
.value-box-warning { border-top-color: #F59E0B; }
.value-box > .inner {
  padding: 13px 18px !important;
}
.value-box .value {
  color: #111827 !important;
  font-size: 30px !important;
  line-height: 1.05;
}
.value-box .caption {
  color: #374151 !important;
  font-size: 13px !important;
  line-height: 1.25;
}
.value-box .icon i {
  top: 15px !important;
  right: 16px !important;
  font-size: 52px !important;
  color: rgba(17, 24, 39, 0.09) !important;
}
.insight-banner {
  margin: 8px 8px 0 0;
  padding: 12px 18px;
  border-left: 4px solid #3B82F6;
  border-radius: 8px;
  background: #EFF6FF;
  color: #1E3A5F;
  font-size: 15px;
}
.synthetic-pill {
  display: inline-block;
  margin-right: 8px;
  padding: 2px 8px;
  border: 1px solid #F59E0B;
  border-radius: 999px;
  background: #FFF7E6;
  color: #92400E;
  font-size: 12px;
  font-weight: 600;
}
table {
  font-variant-numeric: tabular-nums;
}
th {
  background: #F9FAFB;
}
.chart-stage img {
  max-width: 100% !important;
  height: auto !important;
}
@media (max-width: 767px) {
  body { padding-left: 0 !important; }
  html, body, #dashboard-container, .tab-content,
  .dashboard-page-wrapper, .dashboard-column-orientation,
  .dashboard-column, .chart-wrapper, .value-box {
    min-width: 0 !important;
    max-width: 100% !important;
    width: 100% !important;
    box-sizing: border-box !important;
  }
  .dashboard-row {
    min-width: 0 !important;
    flex-direction: column !important;
  }
  .chart-wrapper {
    min-width: 0 !important;
    width: calc(100% - 8px) !important;
  }
  .value-box {
    min-height: 92px !important;
    height: 92px !important;
  }
  .value-box .value { font-size: 27px !important; }
  .chart-stage { overflow-x: auto !important; }
  table { font-size: 12px; }
}
</style>

Overview {data-icon="fa-dashboard"}
=====================================

<div class="insight-banner">
<span class="synthetic-pill">Synthetic demonstration</span>
Simulated values in real export schemas: a weekly Person query, daily Microsoft 365
credit consumption by service, daily GitHub AI credits, GitHub Copilot activity
metrics, GitHub feature/model/language breakdowns, and Consumption metadata. The
analysis first aggregates product feeds to a single Person × week panel, then
uses that joined panel to compare observed product mix with Ways of Working
outcomes. Empty panels and NA values denote withholding or undefined rates, not
zero.
</div>

Row {data-height=120}
-------------------------------------

### People represented

```{r}
flexdashboard::valueBox(
  comma(overall$People), "People in the analysis",
  icon = "fa-users", color = "primary"
)
```

### Observed M365 enabled population

```{r}
flexdashboard::valueBox(
  percent(overall$Licensed / overall$People, accuracy = 0.1),
  "M365 enabled users; based on people, not enabled-day share",
  icon = "fa-id-card", color = "info"
)
```

### M365 credits per person per week

```{r}
flexdashboard::valueBox(
  format(round(overall$`Weekly credits`, 1), big.mark = ",", nsmall = 1),
  "M365 Copilot credits / person / week",
  icon = "fa-calculator", color = "success"
)
```

### Observed GitHub population

```{r}
flexdashboard::valueBox(
  percent(overall$GitHubConsumers / overall$People, accuracy = 0.1),
  "People with observed GitHub Copilot activity or credits",
  icon = "fa-github", color = "warning"
)
```

Row {data-height=510}
-------------------------------------

### Consumption baselines by function

```{r function-baseline, fig.height=4.4, fig.alt="Horizontal bars comparing average weekly M365 credits and sessions per person across disclosure-safe function cohorts."}
if (!nrow(function_summary)) {
  cat("Function partition withheld: one or more cells or complements are below the publication floor.")
} else {
baseline_long <- function_summary |>
  select(FunctionType, People, `Weekly credits`, `M365 sessions`) |>
  pivot_longer(c(`Weekly credits`, `M365 sessions`),
               names_to = "Measure", values_to = "Value") |>
  mutate(
    Measure = recode(
      Measure,
      `Weekly credits` = "Credits / person / week",
      `M365 sessions` = "M365 sessions / person / week"
    ),
    Label = if_else(
      Measure == "Credits / person / week",
      format(round(Value, 1), nsmall = 1),
      format(round(Value, 1), nsmall = 1)
    ),
    FunctionType = factor(FunctionType, levels = FUNCTION_ORDER)
  )

ggplot(baseline_long,
       aes(FunctionType, Value, fill = Measure)) +
  geom_col(position = position_dodge(width = 0.72), width = 0.66) +
  geom_text(aes(label = Label),
            position = position_dodge(width = 0.72),
            hjust = -0.10, size = 3) +
  coord_flip() +
  facet_wrap(~Measure, scales = "free_x") +
  scale_fill_manual(values = PAL[c(1, 3)], guide = "none") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18))) +
  theme_t() +
  labs(
    subtitle = paste0(
      "Small cells/complements are pooled; the complete partition requires n >= ",
      MIN_GROUP_N
    ),
    x = NULL, y = NULL,
    caption = CHART_CAPTION
  )
}
```

### Baseline table

```{r}
function_summary |>
  arrange(desc(`Weekly credits`)) |>
  transmute(
    Function = FunctionType,
    People,
    `M365 credits / person / week` = round(`M365 credits`, 1),
    `M365 sessions / person / week` = round(`M365 sessions`, 1),
    `M365 credits / session` = round(`M365 credits / session`, 2)
  ) |>
  kable(
    caption = "Function baseline; unsafe contributor/complement cells are pooled before release. Credits per session uses observed Microsoft 365 sessions, not tokens."
  )
```

Row {data-height=335}
-------------------------------------

### Population-level highlights

**Product mix is the new baseline.** The report now separates observed M365
Copilot and GitHub Copilot activity before relating either product to working
patterns. M365 credits, GitHub AI credits and GitHub activity remain separate
measures; no common unit or conversion has been confirmed.

**Enablement is a people measure.** M365 enablement is reported as the share of
people with any observed enabled day, not as a percentage of possible enabled
days.

**Ways of Working comparisons use the joined product window.** The Person query
has 26 weeks of collaboration history, while product feeds cover 13 weeks of
weekday daily records. Relationships in this report use the product weeks after
aggregating all product files to a single Person × week panel.

### How to read cost intensity

**M365 credits per session** is retained as an appendix-style intensity measure.
It is not a productivity or value measure, and it does not establish that a
group is more or less efficient.

Product usage mix {data-icon="fa-adjust"}
=====================================

Row {data-height=470}
-------------------------------------

### Observed product mix

```{r product-mix-share, fig.height=4.2, fig.alt="Horizontal bar chart showing population share with observed M365 only, GitHub only, both products, or no observed product use."}
if (!nrow(usage_mix_summary)) {
  cat("Product-mix partition withheld: one or more cells or complements are below the publication floor.")
} else {
ggplot(usage_mix_summary,
       aes(ProductUsageMix, `Share of population`, fill = ProductUsageMix)) +
  geom_col(width = 0.68) +
  geom_text(aes(label = percent(`Share of population`, accuracy = 1)),
            hjust = -0.12, size = 3.2) +
  coord_flip() +
  scale_y_continuous(labels = percent_format(accuracy = 1),
                     expand = expansion(mult = c(0, 0.18))) +
  scale_fill_manual(values = PAL[c(1, 2, 6, 5)], guide = "none") +
  theme_t() +
  labs(
    title = "Observed product mix across the population",
    subtitle = "Presence-based segments avoid comparing M365 and GitHub credit units directly",
    x = NULL, y = "Share of people",
    caption = CHART_CAPTION
  )
}
```

### Product-mix baseline

```{r}
usage_mix_summary |>
  transmute(
    Segment = ProductUsageMix,
    People,
    `Share of population` = percent(`Share of population`, accuracy = 1),
    `M365 enabled users` = percent(`M365 enabled users` / People, accuracy = 1),
    `M365 credits / person / week` = round(`M365 credits / person / week`, 1),
    `GitHub credits / person / week` = round(`GitHub credits / person / week`, 1),
    `GitHub active days / person` = round(`GitHub active days / person`, 1)
  ) |>
  kable(caption = "Observed product mix from the joined person-week panel. No observed product use is not proof of non-use.")
```

Row {data-height=440}
-------------------------------------

### M365 credit volume within observed product mix

```{r product-mix-m365-volume, fig.height=4.1, fig.alt="Bars comparing M365 credits per person per week across observed product-mix segments."}
if (!nrow(usage_mix_summary)) {
  cat("Product-mix partition withheld: one or more cells or complements are below the publication floor.")
} else {
ggplot(usage_mix_summary,
       aes(ProductUsageMix, `M365 credits / person / week`,
           fill = ProductUsageMix)) +
  geom_col(width = 0.68) +
  geom_text(aes(label = round(`M365 credits / person / week`, 1)),
            vjust = -0.35, size = 3.1) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.16))) +
  scale_fill_manual(values = PAL[c(1, 2, 6, 5)], guide = "none") +
  theme_t() +
  theme(axis.text.x = element_text(angle = 25, hjust = 1)) +
  labs(
    title = "M365 credits remain a product-specific measure",
    subtitle = "GitHub AI credits are shown in the table but are not added to M365 credits",
    x = NULL, y = "M365 credits / person / week",
    caption = CHART_CAPTION
  )
}
```

### Join strategy

The dashboard first aggregates every daily or dimensional product export to
`PersonId × MetricDate`, where `MetricDate` is the Sunday week start. The joined
panel then has one row per Person Query row, so downstream charts cannot
accidentally multiply people by service, feature, language or model dimensions.

Collaboration load {data-icon="fa-comments"}
=====================================

Row {data-height=470}
-------------------------------------

### Collaboration volume by observed product mix

```{r collaboration-load, fig.height=4.4, fig.alt="Small bar charts compare collaboration, meeting, chat and email hours across observed product-mix segments."}
collaboration_summary <- product_mix_public |>
  group_by(ProductUsageMix) |>
  summarise(
    People = n(),
    `Collaboration hours` = mean(Collaboration_hours),
    `Meeting hours` = mean(Meeting_hours),
    `Chat hours` = mean(Chat_hours),
    `Email hours` = mean(Email_hours),
    .groups = "drop"
  ) |>
  pivot_longer(-c(ProductUsageMix, People),
               names_to = "Metric", values_to = "Value")

if (!nrow(collaboration_summary)) {
  cat("Collaboration-load partition withheld: one or more cells or complements are below the publication floor.")
} else {
ggplot(collaboration_summary, aes(ProductUsageMix, Value, fill = ProductUsageMix)) +
  geom_col(width = 0.65) +
  geom_text(aes(label = round(Value, 1)), vjust = -0.35, size = 2.9) +
  facet_wrap(~Metric, scales = "free_y", ncol = 2) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.17))) +
  scale_fill_manual(values = PAL[c(1, 2, 6, 5)], guide = "none") +
  theme_t() +
  theme(axis.text.x = element_text(angle = 25, hjust = 1)) +
  labs(
    title = "Collaboration load by observed product mix",
    subtitle = "Bars show weekly Person Query averages in the product window",
    x = NULL, y = "Hours / person / week",
    caption = CHART_CAPTION
  )
}
```

### Collaboration mode mix

```{r mode-mix, fig.height=4.1, fig.alt="One hundred percent stacked bars comparing email, chat, meeting and unscheduled-call shares across observed product-mix segments."}
modes <- c("Email_hours", "Chat_hours", "Meeting_hours",
           "Unscheduled_call_hours")

mode_share <- product_mix_public |>
  group_by(ProductUsageMix) |>
  summarise(People = n(), across(all_of(modes), mean), .groups = "drop") |>
  pivot_longer(all_of(modes), names_to = "Mode", values_to = "Hours") |>
  group_by(ProductUsageMix) |>
  mutate(Share = Hours / sum(Hours)) |>
  ungroup() |>
  mutate(Mode = recode(
    Mode,
    Email_hours = "Email",
    Chat_hours = "Chat",
    Meeting_hours = "Meetings",
    Unscheduled_call_hours = "Unscheduled calls"
  ))

if (!nrow(mode_share)) {
  cat("Mode-mix partition withheld: one or more cells or complements are below the publication floor.")
} else {
ggplot(mode_share, aes(ProductUsageMix, Share, fill = Mode)) +
  geom_col(width = 0.68) +
  geom_text(
    aes(label = ifelse(Share >= 0.06, percent(Share, accuracy = 1), "")),
    position = position_stack(vjust = 0.5),
    colour = ifelse(mode_share$Mode %in% c("Meetings", "Unscheduled calls"),
                    "#1F2937", "white"),
    size = 3
  ) +
  coord_flip() +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  scale_fill_manual(values = PAL[1:4]) +
  theme_t() +
  labs(
    title = "Share of collaboration hours by mode",
    subtitle = "Mode hours are shown before deduplication, so shares may not sum exactly across modes",
    x = NULL, y = NULL, fill = "Collaboration mode"
  )
}
```

Network breadth {data-icon="fa-sitemap"}
=====================================

Row {data-height=470}
-------------------------------------

### Network size by observed product mix

```{r network-breadth, fig.height=4.4, fig.alt="Small bar charts compare internal network size, external network size, strong ties and diverse ties across observed product-mix segments."}
network_summary <- product_mix_public |>
  group_by(ProductUsageMix) |>
  summarise(
    People = n(),
    `Internal network size` = mean(Internal_network_size),
    `External network size` = mean(External_network_size),
    `Strong ties` = mean(Strong_ties),
    `Diverse ties` = mean(Diverse_ties),
    .groups = "drop"
  ) |>
  pivot_longer(-c(ProductUsageMix, People),
               names_to = "Metric", values_to = "Value")

if (!nrow(network_summary)) {
  cat("Network partition withheld: one or more cells or complements are below the publication floor.")
} else {
ggplot(network_summary, aes(ProductUsageMix, Value, fill = ProductUsageMix)) +
  geom_col(width = 0.65) +
  geom_text(aes(label = round(Value, 1)), vjust = -0.35, size = 2.9) +
  facet_wrap(~Metric, scales = "free_y", ncol = 2) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.17))) +
  scale_fill_manual(values = PAL[c(1, 2, 6, 5)], guide = "none") +
  theme_t() +
  theme(axis.text.x = element_text(angle = 25, hjust = 1)) +
  labs(
    title = "Network breadth by observed product mix",
    subtitle = "Weekly Person Query network metrics are averaged, never summed",
    x = NULL, y = "People / ties",
    caption = CHART_CAPTION
  )
}
```

### Reading the network measures

Network metrics describe collaboration breadth in the observed calendar and
communication graph. They do not measure relationship quality, influence,
delivery quality or business impact. The useful question is whether product-use
segments have visibly different collaboration reach that deserves deeper
analysis.

Row {data-height=425}
-------------------------------------

### M365 credit volume and network

```{r credit-network, fig.height=4.0, fig.alt="Bars compare internal and external network size across M365 credit-volume groups."}
if (!nrow(credit_public)) {
  cat("Credit-volume partition withheld: one or more cells or complements are below the publication floor.")
} else {
credit_public |>
  group_by(CreditQuartile) |>
  summarise(
    People = n(),
    `Internal network size` = mean(Internal_network_size),
    `External network size` = mean(External_network_size),
    .groups = "drop"
  ) |>
  pivot_longer(-c(CreditQuartile, People),
               names_to = "Metric", values_to = "Value") |>
  ggplot(aes(CreditQuartile, Value)) +
  geom_col(fill = PAL[2], width = 0.66) +
  geom_text(aes(label = round(Value, 1)), vjust = -0.35, size = 2.8) +
  facet_wrap(~Metric, scales = "free_y", ncol = 2) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.16))) +
  theme_t() +
  theme(axis.text.x = element_text(angle = 25, hjust = 1)) +
  labs(
    title = "Network size by M365 credit-volume group",
    subtitle = "This product-specific lens uses M365 credit groups only",
    x = NULL, y = NULL
  )
}
```

After-hours working {data-icon="fa-clock-o"}
=====================================

Row {data-height=470}
-------------------------------------

### After-hours and weekend collaboration by product mix

```{r after-hours-product-mix, fig.height=4.4, fig.alt="Small bar charts compare after-hours collaboration, after-hours meetings, after-hours email, after-hours chat and weekend collaboration across observed product-mix segments."}
after_hours_summary <- product_mix_public |>
  group_by(ProductUsageMix) |>
  summarise(
    People = n(),
    `After-hours collaboration` = mean(After_hours_collaboration_hours),
    `After-hours meetings` = mean(After_hours_meeting_hours),
    `After-hours email` = mean(After_hours_email_hours),
    `After-hours chat` = mean(After_hours_chat_hours),
    `Weekend collaboration` = mean(Weekend_collaboration_hours),
    .groups = "drop"
  ) |>
  pivot_longer(-c(ProductUsageMix, People),
               names_to = "Metric", values_to = "Value")

if (!nrow(after_hours_summary)) {
  cat("After-hours partition withheld: one or more cells or complements are below the publication floor.")
} else {
ggplot(after_hours_summary, aes(ProductUsageMix, Value, fill = ProductUsageMix)) +
  geom_col(width = 0.65) +
  geom_text(aes(label = round(Value, 2)), vjust = -0.35, size = 2.8) +
  facet_wrap(~Metric, scales = "free_y", ncol = 2) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18))) +
  scale_fill_manual(values = PAL[c(1, 2, 6, 5)], guide = "none") +
  theme_t() +
  theme(axis.text.x = element_text(angle = 25, hjust = 1)) +
  labs(
    title = "After-hours working by observed product mix",
    subtitle = "After-hours collaboration is not total work time and does not imply harm",
    x = NULL, y = "Hours / person / week",
    caption = CHART_CAPTION
  )
}
```

### Interpretation guardrail

After-hours collaboration counts meetings, email, Teams chats, calls and
channels with at least one other person outside configured working hours, with
overlapping activity deduplicated. It is not total work time. Flexible
schedules, on-call duties and time zones may matter, but none is recorded here.

Row {data-height=490}
-------------------------------------

### M365 cost intensity and workload

```{r cost-outcomes, fig.height=4.1, fig.alt="Bars comparing after-hours collaboration, available-to-focus hours and uninterrupted hours across standard-cost and high-cost-per-session profiles."}
if (!nrow(profile_grid)) {
  cat("Profile partition withheld: one or more cells or complements are below the publication floor.")
} else {
profile_grid |>
  select(ConsumptionProfile, People, `Cost-per-session profile`,
         `After-hours hours`, `Available-to-focus hours`,
         `Uninterrupted hours`) |>
  pivot_longer(-c(ConsumptionProfile, People, `Cost-per-session profile`),
               names_to = "Metric", values_to = "Value") |>
  ggplot(
    aes(reorder(ConsumptionProfile, Value), Value,
        fill = `Cost-per-session profile`)
  ) +
  geom_col(width = 0.65) +
  geom_text(aes(label = round(Value, 2)), hjust = -0.12, size = 3) +
  coord_flip() +
  facet_wrap(~Metric, scales = "free_x") +
  scale_fill_manual(
    values = c("High cost per session" = PAL[6],
               "Standard cost per session" = PAL[2],
               "Limited M365 sessions" = PAL[5],
               "No observed M365 credits" = PAL[4])
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18))) +
  theme_t() +
  labs(
    title = "Workload by M365 cost intensity, within credit bands",
    subtitle = "Limited-session people are displayed separately, not assigned a cost-intensity class",
    x = NULL, y = NULL, fill = "Cost-per-session profile"
  )
}
```

Appendix: consumption detail {data-icon="fa-bookmark"}
=====================================

Row {data-height=455}
-------------------------------------

### Service mix

```{r service-mix, fig.height=4.1, fig.alt="Stacked bars comparing the share of Microsoft 365 credits used by service for high-credit and lower-credit users."}
service_people <- m365_service_weekly |>
  left_join(select(person_base, PersonId, CreditBand),
            by = "PersonId") |>
  filter(!is.na(CreditBand), CreditBand != "No observed M365 credits")
service_public <- publication_partition(
  service_people, c("CreditBand", "ServiceName"),
  c("m365_credits", "m365_sessions")
)
service_support <- service_people |>
  distinct(CreditBand, ServiceName, PersonId) |>
  count(CreditBand, ServiceName, name = "Contributors") |>
  left_join(count(person_base, CreditBand, name = "ParentPeople"),
            by = "CreditBand")
if (any(is.na(publication_count(service_support$Contributors,
                               service_support$ParentPeople)))) {
  service_public <- service_public[0, ]
}
if (!nrow(service_public)) {
  cat("The entire service-mix partition is withheld: a service cohort or its complement is below the publication floor. No partial shares or totals are shown.")
} else {
service_public |>
  group_by(CreditBand, ServiceName) |>
  summarise(Credits = sum(m365_credits), People = n_distinct(PersonId),
            .groups = "drop") |>
  group_by(CreditBand) |>
  mutate(Share = Credits / sum(Credits)) |>
  ungroup() |>
  ggplot(aes(CreditBand, Share, fill = reorder(ServiceName, Share))) +
  geom_col(width = 0.68) +
  geom_text(
    aes(label = ifelse(Share >= 0.06, percent(Share, accuracy = 1), "")),
    position = position_stack(vjust = 0.5),
    colour = "grey15", size = 2.9
  ) +
  coord_flip() +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  scale_fill_brewer(palette = "Blues", direction = -1) +
  theme_t() +
  labs(
    title = "Microsoft 365 credit mix by service",
    subtitle = "ServiceName is retained as a separate fact table to avoid multiplying the super panel",
    x = NULL, y = "Share of M365 credits", fill = "Service"
  )
}
```

### M365 credit-volume groups

```{r}
credit_public |>
  group_by(CreditQuartile) |>
  summarise(
    People = n(),
    `Credits / person / week` = mean(weekly_m365_credits),
    `M365 credits / session` =
      safe_divide(sum(total_m365_credits), sum(total_m365_sessions)),
    .groups = "drop"
  ) |>
  mutate(
    `Share of population` = percent(People / sum(People), accuracy = 0.1),
    `Credits / person / week` = round(`Credits / person / week`, 1),
    `M365 credits / session` = if_else(
      is.na(`M365 credits / session`),
      "N/A",
      format(round(`M365 credits / session`, 2), nsmall = 2)
    )
  ) |>
  kable(caption = "M365 credit-volume groups. Quartiles are calculated among people with positive observed M365 credits; people with no observed M365 credits are retained separately.")
```

Row {data-height=115}
-------------------------------------

### Cost-intensity profile availability

```{r cost-profile-note}
if (!nrow(profile_grid)) {
  cat("Profile partition withheld: one or more cost-intensity cells or complements are below the publication floor.")
} else {
  cat("Profile partition is publishable. See the after-hours page for the cost-intensity workload comparison and the baseline table below for profile definitions.")
}
```

Row {data-height=340}
-------------------------------------

### Profile definitions

| Profile component | Definition | Interpretation |
|---|---|---|
| High credit | Upper quartile of average weekly M365 credits among positive M365 users | Higher M365 credit volume |
| Lower credit | Other positive M365 users | Lower M365 credit volume |
| High cost per session | Upper quartile of M365 credits per session among people with at least 10 observed M365 sessions | More credits used per observed M365 session |
| Standard cost per session | Below the high-cost threshold, with enough observed sessions | Lower observed M365 cost intensity |
| Limited M365 sessions | Fewer than 10 observed M365 sessions, including zero-session and GitHub-only people | Insufficient sessions; never included in standard/high-cost comparisons |

The label **high cost per session** describes a measurable consumption pattern.
It is not a wellbeing, productivity or value judgement.

Row {data-height=115}
-------------------------------------

### Profile baseline

```{r profile-table}
if (!nrow(profile_summary)) {
  cat("Profile baseline withheld: the complete ConsumptionProfile by CreditQuartile partition does not meet the publication floor.")
} else {
  profile_summary |>
    kable(digits = 2, caption = "M365 profiles, including the distinct insufficient-session cohort.")
}
```

Methods and appendix {data-icon="fa-book"}
=====================================

Row {data-height=330}
-------------------------------------

### Join and aggregation workflow

```{r workflow, fig.height=3.4}
steps <- tibble(
  x = c(1, 3, 5, 7, 9),
  label = c(
    "Consumption query\nPerson x service x day",
    "Aggregate services\nbefore person-week rates",
    "Person Query\nPerson x week",
    "Join credits by\nPersonId and week",
    "Population,\nservice and function views"
  )
)

ggplot(steps, aes(x, 1)) +
  geom_segment(
    aes(xend = x + 1.35, yend = 1),
    arrow = grid::arrow(length = grid::unit(0.14, "in")),
    colour = PAL[2], linewidth = 0.8
  ) +
  geom_label(
    aes(label = label), fill = "#EAF2F8", colour = PAL[1],
    label.size = 0.25, size = 3.1, lineheight = 0.95
  ) +
  coord_cartesian(xlim = c(0, 10.6), ylim = c(0.6, 1.4), clip = "off") +
  theme_void() +
  labs(title = "Time-aligned person-week panel")
```

Row {data-height=450}
-------------------------------------

### Core rules

| Question | Rule |
|---|---|
| Identity | Take the `PersonId` / `PeopleHistoricalId` crosswalk from Consumption activity files; never derive opaque identifiers by string manipulation |
| Metadata | Join Consumption activity to `PeopleMetaData` with `PeopleHistoricalId`; use person-query attributes for whole-population grouping because a person query alone has no `PeopleHistoricalId` |
| Weekly relationships | Aggregate daily product records to person-week, then join to person-query weeks by `PersonId` and `MetricDate` |
| Service grain | Aggregate across `ServiceName` before computing person-week M365 credits or credits per session |
| Credit quartiles | Use M365 credits only; calculate among positive M365 users and retain people with no observed M365 credits separately |
| Cost intensity | Use M365 credits per observed M365 session; do not infer token mix |
| Focus identity | Check that meetings, scheduled calls and available-to-focus hours reconcile to the configured working-hours basis, and that uninterrupted plus interrupted hours reconcile to available-to-focus hours |
| Network measures | Use weekly person-query network metrics and average them across weeks; never sum them |
| Privacy | At least 10 distinct people per published cell, contributing cohort and non-empty contributor complement. Pool unsafe function cells, then recheck; withhold an entire failing partition consistently across tables, charts, shares and ranks |

Row {data-height=190}
-------------------------------------

### Key crosswalk check

```{r}
unmatched_key_counts |>
  transmute(Check, Status = if_else(Count == 0, "Pass", "Review required")) |>
  kable(caption = "Crosswalk diagnostics are qualitative: no small exception counts or reconstructible crosswalk totals are published.")
```

Row {data-height=430}
-------------------------------------

### Simulation design

This dashboard is a demonstration of analytical possibilities. Every person,
credit, session and Ways of Working value is synthetic.

The current source set represents `r comma(overall$People)` synthetic
people. The person query spans 26 Sunday-start weeks. The M365 and GitHub
credit feeds cover the final `r length(product_weeks)` weeks at weekday daily
grain, which is why this report filters Ways of Working measures to the product
window before relating them to credits.

Available-to-focus hours are simulated on a 40-hour weekly basis here, but that
basis is configurable per tenant through metric rules. The metric describes
calendar time left after meetings and scheduled Teams calls, and should not be
read as coding time or psychological flow.

The following effects are deliberately modest:

- Higher consumption is associated with more chat and less email as a share of
  collaboration modes.
- Total collaboration volume is raised mainly by organisational composition.
- Internal network size has a small residual association after composition is
  considered.
- Cost per M365 session has a weak association with after-hours collaboration
  in some credit bands.
- External network size is a null.

Row {data-height=260}
-------------------------------------

### Synthetic source assets

The uncompressed source-shaped CSVs sit under:

```text
examples/utility-r/_data/
  person-query/PersonQuery.csv
  consumption-query/PeopleMetaData.csv
  consumption-query/PersonM365CreditsMetrics.csv
  consumption-query/PersonGitHubCreditsMetrics.csv
```

They are intended for demonstrations, workshops and adaptation of the template.
They contain no customer data or real identifiers. The GitHub-query files in the
same folder are not required for this consumption report.

Row {data-height=520}
-------------------------------------

### Definitions and limitations

**Credits per session** means Microsoft 365 Copilot credits consumed divided by
observed Microsoft 365 Consumption-query sessions. It is preferred over any
token-normalised measure because token fields are not present in this export.

**Available-to-focus hours** means working hours remaining after excluding
meetings and scheduled Teams calls. `Uninterrupted_hours` and
`Interrupted_hours` partition available-to-focus time in this source. The
working-hours basis is tenant-configurable; the 40-hour identity checked here is
specific to this synthetic dataset.

**Credit concentration** uses only coarse positive-user M365 quartiles, shared
with the credit-group panels. No individual curve, top-1% point, fine histogram,
observed extrema or individual ranking is published. If any quartile or the
no-observed-M365 group falls below the floor, the entire partition is withheld.

**Credit activity segments** are based on active product weeks and observed
M365 credit volume. They are not the standard `vivainsights` action-based segments,
because this source set does not include a Copilot actions metric.

**PeopleHistoricalId** is the documented join key between the Consumption
activity files and `PeopleMetaData`. It must be taken from the Consumption
export rather than derived from `PersonId`. A person-query export alone cannot
be joined to `PeopleMetaData`; the bridge is an activity file that carries both
`PersonId` and `PeopleHistoricalId`.

The dashboard shows associations. It does not establish that AI consumption
caused a working pattern, and it does not observe output quality, task
difficulty or business value. It also makes no wellbeing claims: uninterrupted,
interrupted, available-to-focus, after-hours and weekend measures describe
observed calendar and collaboration patterns only.

Row {data-height=520}
-------------------------------------

### Replacing the synthetic data

1. Replace the four CSVs listed above with schema-aligned governed exports.
2. Validate the identity crosswalk and disclose the match rate.
3. Require independent, week-relevant eligibility and ingestion/coverage evidence
   before zero filling. Missing rows alone do not establish zero activity or no
   provisioning; M365 enabled days do not establish GitHub provisioning. Missing
   person-weeks remain unknown here. Recorded sums per analysis week are lower
   bounds when completeness is unknown, not complete-period consumption.
4. Preserve the daily service grain until after service-level totals are needed;
   aggregate services before computing person-week rates.
5. Keep person-level averaging so each person carries equal analytical weight.
6. Retain the minimum of 10 and whole-partition disclosure gates, including
   contributor complements and shared gates for related tables/charts. Blank
   charts, empty tables and NA values mean withheld (or undefined rates), not
   zero. Overlapping releases require a governed cross-report disclosure review;
   these within-report checks are not a certification of arbitrary real data.
7. Keep M365 Copilot credits and GitHub AI credits in separate panels and
   denominators unless an authoritative common-unit/conversion contract is
   established.
