---
title: "Copilot Consumption and Ways of Working"
subtitle: "The art of the possible with Consumption, GitHub credits and Viva Insights data (fully synthetic)"
date: "Synthetic window: 4 Jan-28 Jun 2026"
output:
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: scroll
    theme: lumen
    self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.width = 8.5, fig.height = 4.5)
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
library(knitr)
library(ggrepel)

MIN_GROUP_N <- 10L
MIN_M365_SESSIONS <- 10L
PAL <- c("#1F4E79", "#2E75B6", "#7FA8D0", "#BDD3E8",
         "#C9C9C9", "#B03A2E", "#5B8C5A", "#7D5BA6")

theme_t <- function() {
  theme_minimal(base_size = 11) +
    theme(
      panel.grid.minor = element_blank(),
      legend.position = "bottom",
      plot.title.position = "plot",
      plot.title = element_text(face = "bold", colour = PAL[1]),
      plot.subtitle = element_text(colour = "grey30",
                                   margin = margin(t = 2, b = 9)),
      plot.caption = element_text(colour = "grey45", size = 8, hjust = 0,
                                  margin = margin(t = 7)),
      plot.margin = margin(t = 10, r = 16, b = 4, l = 4)
    )
}

read_demo <- function(...) {
  read.csv(file.path("_data", ...),
           check.names = FALSE, stringsAsFactors = FALSE)
}

week_start_sunday <- function(x) {
  x <- as.Date(x)
  x - as.integer(format(x, "%w"))
}

clean_label <- function(x, fallback = "Not disclosed") {
  x <- as.character(x)
  if_else(is.na(x) | !nzchar(trimws(x)), fallback, x)
}

safe_divide <- function(num, den) {
  den <- rep_len(den, length(num))
  out <- rep(NA_real_, length(num))
  valid <- !is.na(den) & den > 0
  out[valid] <- num[valid] / den[valid]
  out
}

# Release complete partitions only: dropping one small cell still exposes it
# through the population total, another panel, or a percentage denominator.
publication_partition <- function(data, groups, contributors = character(),
                                  min_n = MIN_GROUP_N) {
  stopifnot(length(min_n) == 1L, is.finite(min_n), min_n >= 10L)
  cells <- data |>
    group_by(across(all_of(groups))) |>
    summarise(People = n_distinct(PersonId), .groups = "drop")
  safe <- nrow(cells) > 0L && all(cells$People >= min_n) &&
    !anyNA(data$PersonId) && all(nzchar(as.character(data$PersonId)))
  for (metric in contributors) {
    support <- data |>
      group_by(across(all_of(groups))) |>
      summarise(
        People = n_distinct(PersonId),
        Positive = n_distinct(PersonId[!is.na(.data[[metric]]) &
                                         .data[[metric]] > 0]),
        .groups = "drop"
      )
    safe <- safe && all(is.finite(data[[metric]])) &&
      all(support$Positive == 0L | support$Positive >= min_n) &&
      all(support$People - support$Positive == 0L |
            support$People - support$Positive >= min_n)
  }
  if (safe) data else data[0, , drop = FALSE]
}

publication_count <- function(n, total, min_n = MIN_GROUP_N) {
  stopifnot(length(min_n) == 1L, is.finite(min_n), min_n >= 10L)
  ifelse(total >= min_n & (n == 0 | n >= min_n) &
           (total - n == 0 | total - n >= min_n), n, NA_real_)
}

consumption_profile <- function(credit_band, sessions, cost_profile) {
  case_when(
    is.na(sessions) | sessions < MIN_M365_SESSIONS ~
      "Limited M365 sessions",
    credit_band == "High credit" &
      cost_profile == "High cost per session" ~
        "High credit, high cost/session",
    credit_band == "High credit" ~ "High credit, standard cost/session",
    credit_band == "Lower credit" &
      cost_profile == "High cost per session" ~
        "Lower credit, high cost/session",
    credit_band == "Lower credit" ~ "Lower credit, standard cost/session",
    TRUE ~ "No observed M365 credits"
  )
}

people_meta <- read_demo("consumption-query", "PeopleMetaData.csv")
m365_daily_raw <- read_demo("consumption-query", "PersonM365CreditsMetrics.csv") |>
  mutate(
    MetricDate = as.Date(MetricDate),
    WeekStart = week_start_sunday(MetricDate)
  )
github_daily_raw <- read_demo("consumption-query", "PersonGitHubCreditsMetrics.csv") |>
  mutate(
    MetricDate = as.Date(MetricDate),
    WeekStart = week_start_sunday(MetricDate)
  )
pq <- read_demo("person-query", "PersonQuery.csv") |>
  mutate(MetricDate = as.Date(MetricDate))

stopifnot(
  max(abs(pq$Meeting_hours + pq$Scheduled_call_hours +
            pq$Available_to_focus_hours - 40), na.rm = TRUE) <= 0.01,
  max(abs(pq$Uninterrupted_hours + pq$Interrupted_hours -
            pq$Available_to_focus_hours), na.rm = TRUE) <= 0.01,
  all(pq$Uninterrupted_hours <= pq$Available_to_focus_hours + 0.01,
      na.rm = TRUE),
  all(pq$Open_1_hour_block <= pq$Available_to_focus_hours + 0.01,
      na.rm = TRUE)
)

person_keys <- bind_rows(
  select(m365_daily_raw, PersonId, PeopleHistoricalId),
  select(github_daily_raw, PersonId, PeopleHistoricalId)
) |>
  filter(!is.na(PersonId), nzchar(PersonId),
         !is.na(PeopleHistoricalId), nzchar(PeopleHistoricalId)) |>
  distinct()

stopifnot(
  nrow(count(person_keys, PersonId) |> filter(n > 1)) == 0,
  nrow(count(person_keys, PeopleHistoricalId) |> filter(n > 1)) == 0
)

credit_key_rows <- bind_rows(
  select(m365_daily_raw, PeopleHistoricalId),
  select(github_daily_raw, PeopleHistoricalId)
)

unmatched_key_counts <- tibble(
  Check = c(
    "PeopleMetaData people with no activity-file crosswalk",
    "Credit activity rows with PeopleHistoricalId absent from PeopleMetaData"
  ),
  Count = c(
    anti_join(people_meta, person_keys, by = "PeopleHistoricalId") |> nrow(),
    anti_join(credit_key_rows, people_meta, by = "PeopleHistoricalId") |> nrow()
  )
)

m365_daily_person <- m365_daily_raw |>
  group_by(PeopleHistoricalId, PersonId, MetricDate, WeekStart) |>
  summarise(
    m365_sessions = sum(`Session count`, na.rm = TRUE),
    m365_credits = sum(`Total Copilot Credits used`, na.rm = TRUE),
    .groups = "drop"
  )

m365_weekly <- m365_daily_person |>
  group_by(PersonId, MetricDate = WeekStart) |>
  summarise(
    m365_sessions = sum(m365_sessions, na.rm = TRUE),
    m365_credits = sum(m365_credits, na.rm = TRUE),
    .groups = "drop"
  )

m365_service_weekly <- m365_daily_raw |>
  group_by(PersonId, MetricDate = WeekStart, ServiceName) |>
  summarise(
    m365_sessions = sum(`Session count`, na.rm = TRUE),
    m365_credits = sum(`Total Copilot Credits used`, na.rm = TRUE),
    .groups = "drop"
  )

github_weekly <- github_daily_raw |>
  group_by(PersonId, MetricDate = WeekStart) |>
  summarise(github_credits = sum(`Total GitHub AI Credits used`, na.rm = TRUE),
            .groups = "drop")

PRODUCT_WINDOW_START <- min(c(m365_weekly$MetricDate, github_weekly$MetricDate),
                            na.rm = TRUE)
PRODUCT_WINDOW_END <- max(c(m365_weekly$MetricDate, github_weekly$MetricDate),
                          na.rm = TRUE)

product_weeks <- sort(unique(c(m365_weekly$MetricDate, github_weekly$MetricDate)))

person_week <- pq |>
  filter(MetricDate %in% product_weeks) |>
  select(PersonId, MetricDate, Collaboration_hours,
         Meeting_hours, Email_hours, Chat_hours, Scheduled_call_hours,
         Unscheduled_call_hours, Channel_message_hours,
         Active_connected_hours, External_collaboration_hours,
         Available_to_focus_hours, Uninterrupted_hours, Interrupted_hours,
         Open_1_hour_block,
         Recurring_meeting_hours, Conflicting_meeting_hours,
         Meeting_hours_with_six_or_fewer_hours_of_advanced_notice,
         Emails_sent, Chats_sent, Meetings, Calls,
         After_hours_collaboration_hours, After_hours_meeting_hours,
         After_hours_email_hours, After_hours_chat_hours,
         Weekend_collaboration_hours, Collaboration_span,
         Internal_network_size, External_network_size, Strong_ties,
         Diverse_ties, Network_outside_organization,
         Total_Copilot_enabled_days) |>
  left_join(select(m365_weekly, PersonId, MetricDate,
                   m365_sessions, m365_credits),
            by = c("PersonId", "MetricDate")) |>
  left_join(select(github_weekly, PersonId, MetricDate, github_credits),
            by = c("PersonId", "MetricDate")) |>
  mutate(
    credits_per_m365_session = safe_divide(m365_credits, m365_sessions)
  )

WOW_METRICS <- c(
  "Collaboration_hours", "Active_connected_hours",
  "Email_hours", "Chat_hours", "Meeting_hours",
  "Unscheduled_call_hours", "After_hours_collaboration_hours",
  "Collaboration_span", "Weekend_collaboration_hours",
  "Available_to_focus_hours", "Uninterrupted_hours", "Interrupted_hours",
  "Internal_network_size", "External_network_size"
)

metric_labels <- c(
  Collaboration_hours = "Collaboration hours / person / week",
  Active_connected_hours = "Active connected hours / person / week",
  Email_hours = "Email hours / person / week",
  Chat_hours = "Chat hours / person / week",
  Meeting_hours = "Meeting hours / person / week",
  Unscheduled_call_hours = "Unscheduled call hours / person / week",
  After_hours_collaboration_hours =
    "After-hours collaboration hours / person / week",
  Collaboration_span = "Collaboration span (hours)",
  Weekend_collaboration_hours =
    "Weekend collaboration hours / person / week",
  Available_to_focus_hours = "Available-to-focus hours / person / week",
  Uninterrupted_hours = "Uninterrupted hours / person / week",
  Interrupted_hours = "Interrupted hours / person / week",
  Internal_network_size = "Internal network size (people)",
  External_network_size = "External network size (people)"
)

person_consumption <- person_week |>
  group_by(PersonId) |>
  summarise(
    # Recorded sums per analysis week are lower bounds when coverage is unknown.
    # Missing person-weeks above remain missing, never measured zero.
    weekly_m365_credits = sum(m365_credits, na.rm = TRUE) / n(),
    weekly_github_credits = sum(github_credits, na.rm = TRUE) / n(),
    weekly_m365_sessions = sum(m365_sessions, na.rm = TRUE) / n(),
    active_weeks = sum(m365_credits > 0, na.rm = TRUE),
    total_m365_credits = sum(m365_credits, na.rm = TRUE),
    total_github_credits = sum(github_credits, na.rm = TRUE),
    total_m365_sessions = sum(m365_sessions, na.rm = TRUE),
    product_weeks = n(),
    credits_per_m365_session =
      safe_divide(sum(m365_credits, na.rm = TRUE), sum(m365_sessions, na.rm = TRUE)),
    .groups = "drop"
  ) |>
  mutate(
    positive_consumer = total_m365_credits > 0,
    active_week_share = active_weeks / product_weeks,
    positive_quartile = if_else(
      positive_consumer,
      ntile(if_else(positive_consumer, weekly_m365_credits, NA_real_), 4L),
      NA_integer_
    ),
    CreditQuartile = factor(
      case_when(
        !positive_consumer ~ "No observed M365 credits",
        positive_quartile == 1L ~ "Q1 lowest",
        positive_quartile == 2L ~ "Q2",
        positive_quartile == 3L ~ "Q3",
        TRUE ~ "Q4 highest"
      ),
      levels = c("No observed M365 credits", "Q1 lowest", "Q2", "Q3", "Q4 highest")
    )
  )

high_cost_cut <- quantile(
  person_consumption$credits_per_m365_session[
    person_consumption$total_m365_sessions >= MIN_M365_SESSIONS
  ],
  0.75, na.rm = TRUE
)

person_attributes <- pq |>
  arrange(PersonId, MetricDate) |>
  group_by(PersonId) |>
  summarise(
    Organization = first(clean_label(Organization)),
    FunctionType = first(clean_label(FunctionType)),
    LevelDesignation = first(clean_label(LevelDesignation)),
    IsManager = first(clean_label(IsManager)),
    Team = first(clean_label(Team)),
    Role = first(clean_label(Role)),
    Seniority = first(clean_label(Seniority)),
    Tenure = first(clean_label(Tenure)),
    IsDeveloper = first(clean_label(IsDeveloper)),
    IsCopilotLicensed = any(Total_Copilot_enabled_days > 0, na.rm = TRUE),
    .groups = "drop"
  )

person_base <- person_attributes |>
  left_join(person_keys, by = "PersonId") |>
  left_join(person_consumption, by = "PersonId") |>
  mutate(
    across(c(weekly_m365_credits,
             weekly_github_credits, weekly_m365_sessions,
             total_m365_credits, total_github_credits,
             total_m365_sessions, active_weeks, active_week_share),
           \(x) replace_na(x, 0)),
    positive_consumer = replace_na(positive_consumer, FALSE),
    CreditQuartile = replace_na(as.character(CreditQuartile),
                                "No observed M365 credits"),
    CreditQuartile = factor(
      CreditQuartile,
      levels = c("No observed M365 credits", "Q1 lowest", "Q2", "Q3", "Q4 highest")
    ),
    CreditBand = case_when(
      CreditQuartile == "Q4 highest" ~ "High credit",
      positive_consumer ~ "Lower credit",
      TRUE ~ "No observed M365 credits"
    ),
    CostPerSessionProfile = case_when(
      total_m365_sessions < MIN_M365_SESSIONS ~ "Limited M365 sessions",
      !positive_consumer ~ "No observed M365 credits",
      credits_per_m365_session >= high_cost_cut ~ "High cost per session",
      TRUE ~ "Standard cost per session"
    ),
    ConsumptionProfile = consumption_profile(
      CreditBand, total_m365_sessions, CostPerSessionProfile
    ),
    CreditActivitySegment = case_when(
      !positive_consumer ~ "Non-user",
      active_weeks >= 10 & CreditBand == "High credit" ~
        "Power User",
      active_weeks >= 8 ~ "Habitual User",
      active_weeks >= 3 ~ "Novice User",
      TRUE ~ "Low User"
    ),
    CreditActivitySegment = factor(
      CreditActivitySegment,
      levels = c("Power User", "Habitual User", "Novice User",
                 "Low User", "Non-user"),
      ordered = TRUE
    ),
    IsCopilotLicensed = if_else(IsCopilotLicensed, "true", "false")
  )

person_wow <- person_week |>
  group_by(PersonId) |>
  summarise(across(all_of(WOW_METRICS), \(x) mean(x, na.rm = TRUE)),
            .groups = "drop") |>
  left_join(person_base, by = "PersonId")

if (n_distinct(person_base$PersonId) < MIN_GROUP_N) {
  stop("Publication withheld: population does not meet the minimum group size.")
}

m365_contributors <- c("total_m365_credits", "total_m365_sessions")
pooled_functions <- person_wow |>
  group_by(FunctionType) |>
  summarise(
    People = n_distinct(PersonId),
    CreditPeople = n_distinct(PersonId[total_m365_credits > 0]),
    SessionPeople = n_distinct(PersonId[total_m365_sessions > 0]),
    .groups = "drop"
  ) |>
  filter(People < MIN_GROUP_N |
           is.na(publication_count(CreditPeople, People)) |
           is.na(publication_count(SessionPeople, People))) |>
  pull(FunctionType)
function_people <- person_wow |>
  mutate(FunctionType = if_else(FunctionType %in% pooled_functions,
                                "Other functions (pooled)", FunctionType))
function_public <- publication_partition(
  function_people, "FunctionType", m365_contributors
)
# Pool unsafe function cells first, then recheck the complete partition;
# a still-small pooled remainder withholds every function output.
credit_public <- publication_partition(
  person_wow, "CreditQuartile", m365_contributors
)
# Both segment outputs use the same joint gate; a baseline must not restore
# the denominator of a suppressed segment-by-quartile composition.
segment_public <- publication_partition(
  person_wow, c("CreditActivitySegment", "CreditQuartile"),
  m365_contributors
)
profile_public <- publication_partition(
  person_wow, c("ConsumptionProfile", "CreditQuartile"), m365_contributors
)
github_public <- publication_partition(
  person_base, character(), "total_github_credits"
)
overall_public <- publication_partition(
  person_base, character(), m365_contributors
)

function_summary <- function_public |>
  group_by(FunctionType) |>
  summarise(
    People = n(),
    `Weekly credits` = mean(weekly_m365_credits),
    `M365 credits` = mean(weekly_m365_credits),
    `M365 sessions` = mean(weekly_m365_sessions),
    `M365 credits / session` =
      safe_divide(sum(total_m365_credits), sum(total_m365_sessions)),
    `Collaboration hours` = mean(Collaboration_hours),
    `After-hours hours` = mean(After_hours_collaboration_hours),
    `Collaboration span` = mean(Collaboration_span),
    `Available-to-focus hours` = mean(Available_to_focus_hours),
    `Uninterrupted hours` = mean(Uninterrupted_hours),
    `Interrupted hours` = mean(Interrupted_hours),
    `Internal network size` = mean(Internal_network_size),
    .groups = "drop"
  )

FUNCTION_ORDER <- function_summary |>
  arrange(`Weekly credits`) |>
  pull(FunctionType)

overall <- person_base |>
  summarise(
    People = n(),
    Developers = publication_count(sum(IsDeveloper == "Yes"), n()),
    Licensed = publication_count(sum(IsCopilotLicensed == "true"), n()),
    Consumers = publication_count(sum(positive_consumer), n()),
    `Weekly credits` = if (nrow(overall_public)) mean(weekly_m365_credits)
      else NA_real_,
    `M365 credits / session` = if (nrow(overall_public))
      safe_divide(sum(total_m365_credits), sum(total_m365_sessions))
      else NA_real_
  )

CHART_CAPTION <- sprintf(
  paste0(
    "Synthetic data | n = %s people | ",
    "product window %s to %s"
  ),
  comma(overall$People),
  format(PRODUCT_WINDOW_START, "%d %b %Y"),
  format(PRODUCT_WINDOW_END, "%d %b %Y")
)

std_diff <- function(x, grp) {
  keep <- !is.na(x) & !is.na(grp)
  x <- x[keep]
  grp <- grp[keep]
  a <- x[grp]
  b <- x[!grp]
  if (length(a) < MIN_GROUP_N || length(b) < MIN_GROUP_N) return(NA_real_)
  pooled <- sqrt(((length(a) - 1) * var(a) + (length(b) - 1) * var(b)) /
                   (length(a) + length(b) - 2))
  if (is.na(pooled) || pooled == 0) return(NA_real_)
  (mean(a) - mean(b)) / pooled
}
```

<style>
body {
  font-family: "Segoe UI Variable", "Segoe UI", system-ui, sans-serif;
  color: #374151;
  background: #F4F4F4;
}
.navbar {
  min-height: 52px;
}
.navbar-brand {
  font-weight: 600;
}
.chart-wrapper {
  border: 1px solid #E5E7EB;
  border-radius: 8px;
  box-shadow: none;
}
.chart-title {
  color: #374151;
  font-weight: 500;
}
.value-box {
  min-height: 108px !important;
  height: 108px !important;
  color: #111827 !important;
  background: #FFFFFF !important;
  border: 1px solid #E5E7EB;
  border-top: 4px solid #3B82F6;
  border-radius: 8px;
  box-shadow: none;
}
.value-box-info { border-top-color: #14B8A6; }
.value-box-success { border-top-color: #8B5CF6; }
.value-box-warning { border-top-color: #F59E0B; }
.value-box > .inner {
  padding: 13px 18px !important;
}
.value-box .value {
  color: #111827 !important;
  font-size: 30px !important;
  line-height: 1.05;
}
.value-box .caption {
  color: #374151 !important;
  font-size: 13px !important;
  line-height: 1.25;
}
.value-box .icon i {
  top: 15px !important;
  right: 16px !important;
  font-size: 52px !important;
  color: rgba(17, 24, 39, 0.09) !important;
}
.insight-banner {
  margin: 8px 8px 0 0;
  padding: 12px 18px;
  border-left: 4px solid #3B82F6;
  border-radius: 8px;
  background: #EFF6FF;
  color: #1E3A5F;
  font-size: 15px;
}
.synthetic-pill {
  display: inline-block;
  margin-right: 8px;
  padding: 2px 8px;
  border: 1px solid #F59E0B;
  border-radius: 999px;
  background: #FFF7E6;
  color: #92400E;
  font-size: 12px;
  font-weight: 600;
}
table {
  font-variant-numeric: tabular-nums;
}
th {
  background: #F9FAFB;
}
.chart-stage img {
  max-width: 100% !important;
  height: auto !important;
}
@media (max-width: 767px) {
  body { padding-left: 0 !important; }
  html, body, #dashboard-container, .tab-content,
  .dashboard-page-wrapper, .dashboard-column-orientation,
  .dashboard-column, .chart-wrapper, .value-box {
    min-width: 0 !important;
    max-width: 100% !important;
    width: 100% !important;
    box-sizing: border-box !important;
  }
  .dashboard-row {
    min-width: 0 !important;
    flex-direction: column !important;
  }
  .chart-wrapper {
    min-width: 0 !important;
    width: calc(100% - 8px) !important;
  }
  .value-box {
    min-height: 92px !important;
    height: 92px !important;
  }
  .value-box .value { font-size: 27px !important; }
  .chart-stage { overflow-x: auto !important; }
  table { font-size: 12px; }
}
</style>

Overview {data-icon="fa-dashboard"}
=====================================

<div class="insight-banner">
<span class="synthetic-pill">Synthetic demonstration</span>
Simulated values in real export schemas: a weekly Person query, daily Microsoft 365
credit consumption by service, daily GitHub AI credits, and Consumption metadata
joined through the activity-file crosswalk.
Main volume panels use M365 credits only; GitHub AI credits are shown separately.
Empty panels and NA values denote withholding or undefined rates, not zero.
</div>

Row {data-height=120}
-------------------------------------

### People represented

```{r}
flexdashboard::valueBox(
  comma(overall$People), "People in the analysis",
  icon = "fa-users", color = "primary"
)
```

### Observed M365 enabled population

```{r}
flexdashboard::valueBox(
  percent(overall$Licensed / overall$People, accuracy = 0.1),
  "Any M365 enabled days observed; NA means withheld",
  icon = "fa-id-card", color = "info"
)
```

### M365 credits per person per week

```{r}
flexdashboard::valueBox(
  format(round(overall$`Weekly credits`, 1), big.mark = ",", nsmall = 1),
  "M365 Copilot credits / person / week",
  icon = "fa-calculator", color = "success"
)
```

### M365 credits per session

```{r}
flexdashboard::valueBox(
  format(round(overall$`M365 credits / session`, 2), nsmall = 2),
  "Credits / observed M365 session",
  icon = "fa-tachometer", color = "warning"
)
```

Row {data-height=510}
-------------------------------------

### Consumption baselines by function

```{r function-baseline, fig.height=4.4, fig.alt="Horizontal bars comparing average weekly M365 credits and sessions per person across disclosure-safe function cohorts."}
if (!nrow(function_summary)) {
  cat("Function partition withheld: one or more cells or complements are below the publication floor.")
} else {
baseline_long <- function_summary |>
  select(FunctionType, People, `Weekly credits`, `M365 sessions`) |>
  pivot_longer(c(`Weekly credits`, `M365 sessions`),
               names_to = "Measure", values_to = "Value") |>
  mutate(
    Measure = recode(
      Measure,
      `Weekly credits` = "Credits / person / week",
      `M365 sessions` = "M365 sessions / person / week"
    ),
    Label = if_else(
      Measure == "Credits / person / week",
      format(round(Value, 1), nsmall = 1),
      format(round(Value, 1), nsmall = 1)
    ),
    FunctionType = factor(FunctionType, levels = FUNCTION_ORDER)
  )

ggplot(baseline_long,
       aes(FunctionType, Value, fill = Measure)) +
  geom_col(position = position_dodge(width = 0.72), width = 0.66) +
  geom_text(aes(label = Label),
            position = position_dodge(width = 0.72),
            hjust = -0.10, size = 3) +
  coord_flip() +
  facet_wrap(~Measure, scales = "free_x") +
  scale_fill_manual(values = PAL[c(1, 3)], guide = "none") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18))) +
  theme_t() +
  labs(
    subtitle = paste0(
      "Small cells/complements are pooled; the complete partition requires n >= ",
      MIN_GROUP_N
    ),
    x = NULL, y = NULL,
    caption = CHART_CAPTION
  )
}
```

### Baseline table

```{r}
function_summary |>
  arrange(desc(`Weekly credits`)) |>
  transmute(
    Function = FunctionType,
    People,
    `M365 credits / person / week` = round(`M365 credits`, 1),
    `M365 sessions / person / week` = round(`M365 sessions`, 1),
    `M365 credits / session` = round(`M365 credits / session`, 2)
  ) |>
  kable(
    caption = "Function baseline; unsafe contributor/complement cells are pooled before release. Credits per session uses observed Microsoft 365 sessions, not tokens."
  )
```

Row {data-height=335}
-------------------------------------

### Population-level highlights

**Consumption volume and cost intensity are separate.** All main credit-volume
bands, rankings and comparisons use **M365 Copilot credits only**. M365 credits
per session is an intensity measure, not a currency price. GitHub AI credits
have their own panel: no common unit or conversion has been confirmed, so the
two measures are never added, compared as shares, or put in a shared denominator.

**Service mix is now observable.** The Consumption query exposes ServiceName, so
the report can show whether credits are concentrated in Analyst, Cowork,
Researcher or WorkIQ rather than relying on a fabricated task-type dimension.

**Ways of Working comparisons use the aligned product window.** The person query
has 26 weeks of collaboration history, while the credit feeds have 13 weeks of
weekday daily records. Relationships in this report use only the product weeks.
Available-to-focus hours are calendar availability within configured working
hours, not a measure of coding flow or psychological focus.

### How to read cost intensity

**M365 credits per session** is the number of Microsoft 365 Copilot credits used
per observed Consumption-query session. It is a cost-intensity measure, not a
productivity or value measure. It does not establish that a group is more or
less efficient, because the report does not observe output quality, task
difficulty or business value.

Credit distribution {data-icon="fa-chart-area"}
=====================================

Row {data-height=470}
-------------------------------------

### M365 consumption in coarse groups

```{r credit-concentration, fig.height=4.4, fig.alt="Bars showing M365 credit shares for disclosure-safe positive-user quartiles, not individual ranks."}
concentration <- credit_public |>
  filter(positive_consumer) |>
  group_by(CreditQuartile) |>
  summarise(
    People = n_distinct(PersonId),
    Credits = sum(weekly_m365_credits),
    .groups = "drop"
  ) |>
  mutate(
    `Share of users` = safe_divide(People, sum(People)),
    `Share of M365 credits` = safe_divide(Credits, sum(Credits))
  )

ggplot(concentration, aes(CreditQuartile, `Share of M365 credits`)) +
  geom_col(fill = PAL[2], width = 0.68) +
  geom_text(aes(label = percent(`Share of M365 credits`, accuracy = 1)),
            vjust = -0.3) +
  scale_y_continuous(labels = percent_format(),
                     expand = expansion(mult = c(0, 0.15))) +
  theme_t() +
  labs(
    title = "M365 credit concentration by positive-user quartile",
    subtitle = "Whole partition withheld if any cell is too small; no individual ranks or top-1% points",
    x = NULL, y = "Share of positive users' M365 credits",
    caption = CHART_CAPTION
  )
```

### Concentration bands

```{r}
band_table <- concentration |>
  transmute(
    Band = CreditQuartile, People,
    `Share of users` = percent(`Share of users`, accuracy = 1),
    `Share of M365 credits` = percent(`Share of M365 credits`, accuracy = 1)
  )

band_table |>
  kable(caption = "Positive M365 users in coarse quartiles; an empty table means the complete partition is withheld.")
```

Row {data-height=440}
-------------------------------------

### Average M365 volume in the same coarse groups

```{r credit-histogram, fig.height=4.1, fig.alt="Average weekly M365 credits in the same disclosure-safe quartiles as the concentration table."}
ggplot(concentration, aes(CreditQuartile, Credits / People)) +
  geom_col(fill = PAL[2], width = 0.68) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.10))) +
  theme_t() +
  labs(
    title = "M365 volume by positive-user quartile",
    subtitle = "No fine histogram bins, observed extrema or individual data points are published",
    x = NULL, y = "M365 credits / person / week",
    caption = CHART_CAPTION
  )
```

### GitHub AI credits — separate unit

```{r credit-source-split}
if (!nrow(github_public)) {
  cat("GitHub AI credit panel withheld: population, contributors or their complement are below the publication floor.")
} else {
github_public |>
  summarise(
    People = n(),
    `People with observed GitHub AI credits` = sum(total_github_credits > 0),
    `GitHub AI credits / person / week` = round(mean(weekly_github_credits), 1)
  ) |>
  kable(caption = "GitHub AI credits only. The denominator is the synthetic analysis population, not a provisioned or licensed GitHub cohort.")
}
```

Absence of GitHub rows does **not** establish absence of provisioning. M365
enabled days do not establish GitHub eligibility or provisioning. Missing person-weeks
remain unknown. Published credit/session figures sum only supplied records and divide
by analysis weeks; they are recorded-volume lower bounds, not verified complete-period
consumption. “No observed credits” means no positive supplied record, not verified inactivity.

Credit relationships
=====================================

Row {data-height=470}
-------------------------------------

### Ways of Working metrics by credit-consumption group

```{r credit-heatmap, fig.height=4.4, fig.alt="Heatmap of standardised Ways of Working metric differences for five credit-consumption groups."}
credit_levels <- c("No observed M365 credits", "Q1 lowest", "Q2", "Q3", "Q4 highest")

scan <- expand_grid(
  Metric = WOW_METRICS,
  CreditQuartile = credit_levels
) |>
  rowwise() |>
  mutate(d = std_diff(
    credit_public[[Metric]],
    credit_public$CreditQuartile == CreditQuartile
  )) |>
  ungroup() |>
  filter(!is.na(d)) |>
  mutate(
    Metric = unname(metric_labels[Metric]),
    CreditQuartile = factor(CreditQuartile, levels = rev(credit_levels))
  )

ggplot(scan, aes(CreditQuartile, reorder(Metric, d), fill = d)) +
  geom_tile(colour = "white", linewidth = 0.6) +
  geom_text(aes(label = sprintf("%+.2f", d)), size = 3) +
  scale_fill_gradient2(
    low = PAL[6], mid = "#F5F2EA", high = PAL[1],
    midpoint = 0, limits = c(-1.3, 1.3), oob = squish
  ) +
  theme_t() +
  theme(legend.position = "right") +
  labs(
    title = "Standardised difference against everyone outside the group",
    subtitle = "Positive values indicate the group scores above the rest of the population on that measure",
    x = NULL, y = NULL, fill = "Std. difference"
  )
```

### Credit groups

```{r}
credit_public |>
  group_by(CreditQuartile) |>
  summarise(
    People = n(),
    `Credits / person / week` = mean(weekly_m365_credits),
    `M365 credits / session` =
      safe_divide(sum(total_m365_credits), sum(total_m365_sessions)),
    .groups = "drop"
  ) |>
  mutate(
    `Share of population` = percent(People / sum(People), accuracy = 0.1),
    `Credits / person / week` = round(`Credits / person / week`, 1),
    `M365 credits / session` = if_else(
      is.na(`M365 credits / session`),
      "N/A",
      format(round(`M365 credits / session`, 2), nsmall = 2)
    )
  ) |>
  kable(caption = "Credit-volume groups. Quartiles are calculated among positive users; non-users are retained separately.")
```

Row {data-height=455}
-------------------------------------

### Collaboration mode mix by credit-consumption group

```{r mode-mix, fig.height=4.1, fig.alt="One hundred percent stacked bars comparing email, chat, meeting and unscheduled-call shares across credit groups."}
modes <- c("Email_hours", "Chat_hours", "Meeting_hours",
           "Unscheduled_call_hours")

mode_share <- credit_public |>
  group_by(CreditQuartile) |>
  summarise(People = n(), across(all_of(modes), mean), .groups = "drop") |>
  pivot_longer(all_of(modes), names_to = "Mode", values_to = "Hours") |>
  group_by(CreditQuartile) |>
  mutate(Share = Hours / sum(Hours)) |>
  ungroup() |>
  mutate(Mode = recode(
    Mode,
    Email_hours = "Email",
    Chat_hours = "Chat",
    Meeting_hours = "Meetings",
    Unscheduled_call_hours = "Unscheduled calls"
  ))

ggplot(mode_share, aes(CreditQuartile, Share, fill = Mode)) +
  geom_col(width = 0.68) +
  geom_text(
    aes(label = ifelse(Share >= 0.06, percent(Share, accuracy = 1), "")),
    position = position_stack(vjust = 0.5),
    colour = ifelse(mode_share$Mode %in% c("Meetings", "Unscheduled calls"),
                    "#1F2937", "white"),
    size = 3
  ) +
  coord_flip() +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  scale_fill_manual(values = PAL[1:4]) +
  theme_t() +
  labs(
    title = "Share of collaboration hours by mode",
    subtitle = "Mode hours are shown before deduplication, so shares may not sum exactly across modes",
    x = NULL, y = NULL, fill = "Collaboration mode"
  )
```

### Workload, focus and network

```{r credit-outcomes, fig.height=4.1, fig.alt="Small bar charts comparing after-hours collaboration, available-to-focus hours, uninterrupted hours, interrupted hours and network metrics across credit groups."}
if (!nrow(credit_public)) {
  cat("Credit-group partition withheld: one or more cells or complements are below the publication floor.")
} else {
credit_public |>
  group_by(CreditQuartile) |>
  summarise(
    People = n(),
    `After-hours collaboration hours / person / week` =
      mean(After_hours_collaboration_hours),
    `Available-to-focus hours / person / week` =
      mean(Available_to_focus_hours),
    `Uninterrupted hours / person / week` = mean(Uninterrupted_hours),
    `Interrupted hours / person / week` = mean(Interrupted_hours),
    `Internal network size (people)` = mean(Internal_network_size),
    `External network size (people)` = mean(External_network_size),
    .groups = "drop"
  ) |>
  pivot_longer(-c(CreditQuartile, People),
               names_to = "Metric", values_to = "Value") |>
  ggplot(aes(CreditQuartile, Value)) +
  geom_col(fill = PAL[2], width = 0.66) +
  geom_text(aes(label = round(Value, 1)), vjust = -0.35, size = 2.8) +
  facet_wrap(~Metric, scales = "free_y", ncol = 2) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.16))) +
  theme_t() +
  theme(axis.text.x = element_text(angle = 25, hjust = 1)) +
  labs(
    title = "Workload, focus and network by credit-volume group",
    subtitle = "Network measures come directly from the weekly person query and are averaged, not summed",
    x = NULL, y = NULL
  )
}
```

Function drill-down {data-icon="fa-sitemap"}
=====================================

Row {data-height=505}
-------------------------------------

### Consumption and collaboration volume by function

```{r bubble-collaboration, fig.height=4.5, fig.alt="Bubble plot of weekly M365 credits and collaboration hours by function. Bubble size represents all people in the function and colour represents Microsoft 365 credits per session."}
ggplot(
  function_summary,
  aes(`Weekly credits`, `Collaboration hours`,
      size = People, colour = `M365 credits / session`,
      label = FunctionType)
) +
  geom_point(alpha = 0.85) +
  geom_text_repel(size = 3.2, colour = "grey20", seed = 2026,
                  min.segment.length = 0, max.overlaps = Inf,
                  box.padding = 0.6, point.padding = 0.3) +
  scale_size_area(max_size = 14) +
  scale_colour_gradient(low = PAL[3], high = PAL[6], na.value = PAL[5]) +
  scale_x_continuous(expand = expansion(mult = 0.18)) +
  scale_y_continuous(expand = expansion(mult = 0.18)) +
  theme_t() +
  labs(
    title = "Credit consumption against collaboration hours, by function",
    subtitle = "Bubble size shows all people; colour shows Microsoft 365 credits per observed session",
    x = "M365 credits / person / week",
    y = "Collaboration hours / user / week",
    colour = "M365 credits / session",
    size = "People",
    caption = CHART_CAPTION
  )
```

### Reading the chart

The upper-right functions combine high credit volume with high collaboration
volume. That does not imply that one causes the other. Their role mix,
seniority and working practices may raise both.

The colour adds a third dimension: cost per observed Microsoft 365 session.
That can indicate more credit-intensive sessions, but it should be interpreted
with service mix, entitlement and task context rather than treated as a
performance score.

Row {data-height=490}
-------------------------------------

### Workload and credit intensity by function

```{r bubble-workload, fig.height=4.4, fig.alt="Bubble plot of M365 credits and after-hours collaboration by function. Bubble size represents all people and colour represents Microsoft 365 credits per session."}
ggplot(
  function_summary,
  aes(`Weekly credits`, `After-hours hours`,
      size = People, colour = `M365 credits / session`,
      label = FunctionType)
) +
  geom_point(alpha = 0.85) +
  geom_text_repel(size = 3.2, colour = "grey20", seed = 2026,
                  min.segment.length = 0, max.overlaps = Inf,
                  box.padding = 0.6, point.padding = 0.3) +
  scale_size_area(max_size = 14) +
  scale_colour_gradient(low = PAL[3], high = PAL[6], na.value = PAL[5]) +
  scale_x_continuous(expand = expansion(mult = 0.18)) +
  scale_y_continuous(expand = expansion(mult = 0.18)) +
  theme_t() +
  labs(
    title = "Credit consumption against after-hours workload, by function",
    subtitle = "Same bubble-size and colour encoding as the chart above, for comparison across functions",
    x = "M365 credits / person / week",
    y = "After-hours hours / user / week",
    colour = "M365 credits / session",
    size = "People",
    caption = CHART_CAPTION
  )
```

### Function comparison matrix

```{r function-heatmap, fig.height=4.4, fig.alt="Indexed heatmap comparing M365 consumption and Ways of Working measures across disclosure-safe function cohorts."}
heat_metrics <- c(
  "Weekly credits", "M365 credits / session", "Collaboration hours",
  "After-hours hours", "Available-to-focus hours", "Uninterrupted hours",
  "Interrupted hours", "Internal network size"
)

function_summary |>
  mutate(FunctionType = factor(FunctionType, levels = FUNCTION_ORDER)) |>
  select(FunctionType, all_of(heat_metrics)) |>
  pivot_longer(-FunctionType, names_to = "Metric", values_to = "Value") |>
  filter(!is.na(Value)) |>
  mutate(Metric = recode(
    Metric,
    `Weekly credits` = "Credits / user / week",
    `M365 credits / session` = "M365 credits / session",
    `Collaboration hours` = "Collaboration hours / person / week",
    `After-hours hours` = "After-hours hours / person / week",
    `Available-to-focus hours` =
      "Available-to-focus hours / person / week",
    `Uninterrupted hours` = "Uninterrupted hours / person / week",
    `Interrupted hours` = "Interrupted hours / person / week",
    `Internal network size` = "Internal network size (people)"
  )) |>
  group_by(Metric) |>
  mutate(Index = 100 * Value / mean(Value)) |>
  ungroup() |>
  ggplot(aes(FunctionType, Metric, fill = Index)) +
  geom_tile(colour = "white", linewidth = 0.7) +
  geom_text(aes(label = round(Index)), size = 3) +
  scale_fill_gradient2(
    low = "#E8EEF5", mid = "#F5F2EA", high = PAL[1], midpoint = 100
  ) +
  theme_t() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1),
        legend.position = "right") +
  labs(
    title = "Each metric indexed to the function average (100)",
    subtitle = "A compact view of consumption, focus and network measures across functions",
    x = NULL, y = NULL, fill = "Index"
  )
```

Consumption mix {data-icon="fa-adjust"}
=====================================

Row {data-height=480}
-------------------------------------

### Credit volume by cost-per-session profile

```{r cost-profile, fig.height=4.3, fig.alt="Grouped bars comparing weekly credit volume for high cost per session and standard cost per session profiles within each credit band."}
profile_summary <- profile_public |>
  group_by(ConsumptionProfile) |>
  summarise(
    People = n(),
    `Credits / person / week` = mean(weekly_m365_credits),
    `M365 credits / session` =
      if (isTRUE(first(ConsumptionProfile) == "Limited M365 sessions")) NA_real_
      else safe_divide(sum(total_m365_credits), sum(total_m365_sessions)),
    `After-hours hours` = mean(After_hours_collaboration_hours),
    `Available-to-focus hours` = mean(Available_to_focus_hours),
    `Uninterrupted hours` = mean(Uninterrupted_hours),
    .groups = "drop"
  )

profile_grid <- profile_summary |>
  mutate(
    CreditBand = case_when(
      ConsumptionProfile == "Limited M365 sessions" ~ "Limited M365 sessions",
      grepl("^High credit", ConsumptionProfile) ~ "High credit",
      grepl("^Lower credit", ConsumptionProfile) ~ "Lower credit",
      TRUE ~ "No observed M365 credits"
    ),
    `Cost-per-session profile` = case_when(
      ConsumptionProfile == "Limited M365 sessions" ~ "Limited M365 sessions",
      grepl("high cost/session", ConsumptionProfile) ~ "High cost per session",
      grepl("standard cost/session", ConsumptionProfile) ~ "Standard cost per session",
      TRUE ~ "No observed M365 credits"
    )
  )

ggplot(
  profile_grid,
  aes(CreditBand, `Credits / person / week`,
      fill = `Cost-per-session profile`)
) +
  geom_col(position = position_dodge(width = 0.7), width = 0.6) +
  geom_text(
    aes(label = round(`Credits / person / week`, 1)),
    position = position_dodge(width = 0.7), vjust = -0.4, size = 3.2
  ) +
  scale_fill_manual(values = c(
    "High cost per session" = PAL[6],
    "Standard cost per session" = PAL[3],
    "Limited M365 sessions" = PAL[5],
    "No observed M365 credits" = PAL[4]
  )) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18))) +
  theme_t() +
  labs(
    title = "Credit volume and M365 cost intensity",
    subtitle = "Insufficient-session people remain separate and are not standard-cost comparators",
    x = "M365 credit-volume band", y = "M365 credits / person / week",
    fill = "Cost-per-session profile"
  )
```

### Profile definitions

| Profile component | Definition | Interpretation |
|---|---|---|
| High credit | Upper quartile of average weekly M365 credits among positive M365 users | Higher M365 credit volume |
| Lower credit | Other positive M365 users | Lower M365 credit volume |
| High cost per session | Upper quartile of M365 credits per session among people with at least 10 observed M365 sessions | More credits used per observed M365 session |
| Standard cost per session | Below the high-cost threshold, with enough observed sessions | Lower observed M365 cost intensity |
| Limited M365 sessions | Fewer than 10 observed M365 sessions, including zero-session and GitHub-only people | Insufficient sessions; never included in standard/high-cost comparisons |

The label **high cost per session** describes a measurable consumption pattern.
It is not a wellbeing, productivity or value judgement.

```{r profile-table}
profile_summary |>
  kable(digits = 2, caption = "M365 profiles, including the distinct insufficient-session cohort. Empty means the entire profile partition is withheld.")
```

Row {data-height=455}
-------------------------------------

### Workload by cost-per-session profile within credit bands

```{r cost-outcomes, fig.height=4.1, fig.alt="Bars comparing after-hours collaboration, available-to-focus hours and uninterrupted hours across standard-cost and high-cost-per-session profiles."}
if (!nrow(profile_grid)) {
  cat("Profile partition withheld: one or more cells or complements are below the publication floor.")
} else {
profile_grid |>
  select(ConsumptionProfile, People, `Cost-per-session profile`,
         `After-hours hours`, `Available-to-focus hours`,
         `Uninterrupted hours`) |>
  pivot_longer(-c(ConsumptionProfile, People, `Cost-per-session profile`),
               names_to = "Metric", values_to = "Value") |>
  ggplot(
    aes(reorder(ConsumptionProfile, Value), Value,
        fill = `Cost-per-session profile`)
  ) +
  geom_col(width = 0.65) +
  geom_text(aes(label = round(Value, 2)), hjust = -0.12, size = 3) +
  coord_flip() +
  facet_wrap(~Metric, scales = "free_x") +
  scale_fill_manual(
    values = c("High cost per session" = PAL[6],
               "Standard cost per session" = PAL[2],
               "Limited M365 sessions" = PAL[5],
               "No observed M365 credits" = PAL[4])
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18))) +
  theme_t() +
  labs(
    title = "Workload by M365 cost intensity, within credit bands",
    subtitle = "Limited-session people are displayed separately, not assigned a cost-intensity class",
    x = NULL, y = NULL, fill = "Cost-per-session profile"
  )
}
```

### Service mix

```{r service-mix, fig.height=4.1, fig.alt="Stacked bars comparing the share of Microsoft 365 credits used by service for high-credit and lower-credit users."}
service_people <- m365_service_weekly |>
  left_join(select(person_base, PersonId, CreditBand),
            by = "PersonId") |>
  filter(!is.na(CreditBand), CreditBand != "No observed M365 credits")
service_public <- publication_partition(
  service_people, c("CreditBand", "ServiceName"),
  c("m365_credits", "m365_sessions")
)
service_support <- service_people |>
  distinct(CreditBand, ServiceName, PersonId) |>
  count(CreditBand, ServiceName, name = "Contributors") |>
  left_join(count(person_base, CreditBand, name = "ParentPeople"),
            by = "CreditBand")
if (any(is.na(publication_count(service_support$Contributors,
                               service_support$ParentPeople)))) {
  service_public <- service_public[0, ]
}
if (!nrow(service_public)) {
  cat("The entire service-mix partition is withheld: a service cohort or its complement is below the publication floor. No partial shares or totals are shown.")
} else {
service_public |>
  group_by(CreditBand, ServiceName) |>
  summarise(Credits = sum(m365_credits), People = n_distinct(PersonId),
            .groups = "drop") |>
  group_by(CreditBand) |>
  mutate(Share = Credits / sum(Credits)) |>
  ungroup() |>
  ggplot(aes(CreditBand, Share, fill = reorder(ServiceName, Share))) +
  geom_col(width = 0.68) +
  geom_text(
    aes(label = ifelse(Share >= 0.06, percent(Share, accuracy = 1), "")),
    position = position_stack(vjust = 0.5),
    colour = "grey15", size = 2.9
  ) +
  coord_flip() +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  scale_fill_brewer(palette = "Blues", direction = -1) +
  theme_t() +
  labs(
    title = "Microsoft 365 credit mix by service",
    subtitle = "ServiceName is a real Consumption-query dimension",
    x = NULL, y = "Share of M365 credits", fill = "Service"
  )
}
```

Usage segments {data-icon="fa-users"}
=====================================

Row {data-height=480}
-------------------------------------

### Credit-volume distribution within activity segments

```{r segment-credit, fig.height=4.3, fig.alt="Stacked bars showing credit-consumption quartiles within Power, Habitual, Novice, Low and Non-user credit activity segments."}
segment_credit <- segment_public |>
  count(CreditActivitySegment, CreditQuartile, name = "People") |>
  group_by(CreditActivitySegment) |>
  mutate(Share = People / sum(People)) |>
  ungroup()

if (!nrow(segment_credit)) {
  cat("The entire segment/quartile partition and its segment baseline are withheld: one or more cells are below the publication floor.")
} else {
ggplot(segment_credit, aes(CreditActivitySegment, Share,
                           fill = CreditQuartile)) +
  geom_col(width = 0.68) +
  coord_flip() +
  scale_x_discrete(limits = rev(levels(person_base$CreditActivitySegment))) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  scale_fill_manual(values = PAL[c(5, 4, 3, 2, 1)]) +
  theme_t() +
  labs(
    title = "Credit-volume distribution within credit activity segments",
    subtitle = paste0(
      "Segments are based on active M365 weeks and observed M365 credits ",
      "from ", format(PRODUCT_WINDOW_START, "%d %b %Y"), " to ",
      format(PRODUCT_WINDOW_END, "%d %b %Y")
    ),
    x = NULL, y = NULL, fill = "Credit-volume group"
  )
}
```

### Why replace standard usage segments?

The `vivainsights::identify_usage_segments()` helper expects a sustained
activity metric such as Copilot actions. This export set does not contain a
Copilot actions column in either the Consumption query or the selected person
query. The report therefore uses a transparent credit activity segmentation:
active product weeks plus credit volume. It does not fabricate an actions field.

Row {data-height=400}
-------------------------------------

### Segment baseline

```{r}
segment_public |>
  group_by(CreditActivitySegment) |>
  summarise(
    People = n(),
    `Active weeks` = mean(active_weeks),
    `M365 credits / person / week` = mean(weekly_m365_credits),
    `M365 credits / session` =
      safe_divide(sum(total_m365_credits), sum(total_m365_sessions)),
    .groups = "drop"
  ) |>
  arrange(CreditActivitySegment) |>
  mutate(
    across(c(`Active weeks`, `M365 credits / person / week`),
           \(x) round(x, 2)),
    `M365 credits / session` = if_else(
      is.na(`M365 credits / session`),
      "N/A",
      format(round(`M365 credits / session`, 2), nsmall = 2)
    )
  ) |>
  kable(caption = paste0(
    "Segment baseline across ", length(product_weeks),
    " product weeks. Non-users have no observed M365 credits, not necessarily no GitHub activity. Empty means the joint segment/quartile partition is withheld."
  ))
```

### Interpretation guardrail

Neither activity segment nor credit group is a performance category. They
describe observed use in the product window. Results should be reported for
privacy-safe cohorts and interpreted alongside role, entitlement, service mix
and organisational context.

Methods and appendix {data-icon="fa-book"}
=====================================

Row {data-height=410}
-------------------------------------

### Join and aggregation workflow

```{r workflow, fig.height=3.4}
steps <- tibble(
  x = c(1, 3, 5, 7, 9),
  label = c(
    "Consumption query\nPerson x service x day",
    "Aggregate services\nbefore person-week rates",
    "Person Query\nPerson x week",
    "Join credits by\nPersonId and week",
    "Population,\nservice and function views"
  )
)

ggplot(steps, aes(x, 1)) +
  geom_segment(
    aes(xend = x + 1.35, yend = 1),
    arrow = grid::arrow(length = grid::unit(0.14, "in")),
    colour = PAL[2], linewidth = 0.8
  ) +
  geom_label(
    aes(label = label), fill = "#EAF2F8", colour = PAL[1],
    label.size = 0.25, size = 3.1, lineheight = 0.95
  ) +
  coord_cartesian(xlim = c(0, 10.6), ylim = c(0.6, 1.4), clip = "off") +
  theme_void() +
  labs(title = "Time-aligned person-week panel")
```

### Core rules

| Question | Rule |
|---|---|
| Identity | Take the `PersonId` / `PeopleHistoricalId` crosswalk from Consumption activity files; never derive opaque identifiers by string manipulation |
| Metadata | Join Consumption activity to `PeopleMetaData` with `PeopleHistoricalId`; use person-query attributes for whole-population grouping because a person query alone has no `PeopleHistoricalId` |
| Weekly relationships | Aggregate daily product records to person-week, then join to person-query weeks by `PersonId` and `MetricDate` |
| Service grain | Aggregate across `ServiceName` before computing person-week M365 credits or credits per session |
| Credit quartiles | Use M365 credits only; calculate among positive M365 users and retain people with no observed M365 credits separately |
| Cost intensity | Use M365 credits per observed M365 session; do not infer token mix |
| Focus identity | Check that meetings, scheduled calls and available-to-focus hours reconcile to the configured working-hours basis, and that uninterrupted plus interrupted hours reconcile to available-to-focus hours |
| Network measures | Use weekly person-query network metrics and average them across weeks; never sum them |
| Privacy | At least 10 distinct people per published cell, contributing cohort and non-empty contributor complement. Pool unsafe function cells, then recheck; withhold an entire failing partition consistently across tables, charts, shares and ranks |

### Key crosswalk check

```{r}
unmatched_key_counts |>
  transmute(Check, Status = if_else(Count == 0, "Pass", "Review required")) |>
  kable(caption = "Crosswalk diagnostics are qualitative: no small exception counts or reconstructible crosswalk totals are published.")
```

Row {data-height=390}
-------------------------------------

### Simulation design

This dashboard is a demonstration of analytical possibilities. Every person,
credit, session and Ways of Working value is synthetic.

The current source set represents `r comma(overall$People)` synthetic
people. The person query spans 26 Sunday-start weeks. The M365 and GitHub
credit feeds cover the final `r length(product_weeks)` weeks at weekday daily
grain, which is why this report filters Ways of Working measures to the product
window before relating them to credits.

Available-to-focus hours are simulated on a 40-hour weekly basis here, but that
basis is configurable per tenant through metric rules. The metric describes
calendar time left after meetings and scheduled Teams calls, and should not be
read as coding time or psychological flow.

The following effects are deliberately modest:

- Higher consumption is associated with more chat and less email as a share of
  collaboration modes.
- Total collaboration volume is raised mainly by organisational composition.
- Internal network size has a small residual association after composition is
  considered.
- Cost per M365 session has a weak association with after-hours collaboration
  in some credit bands.
- External network size is a null.

### Synthetic source assets

The uncompressed source-shaped CSVs sit under:

```text
examples/utility-r/_data/
  person-query/PersonQuery.csv
  consumption-query/PeopleMetaData.csv
  consumption-query/PersonM365CreditsMetrics.csv
  consumption-query/PersonGitHubCreditsMetrics.csv
```

They are intended for demonstrations, workshops and adaptation of the template.
They contain no customer data or real identifiers. The GitHub-query files in the
same folder are not required for this consumption report.

Row {data-height=430}
-------------------------------------

### Definitions and limitations

**Credits per session** means Microsoft 365 Copilot credits consumed divided by
observed Microsoft 365 Consumption-query sessions. It is preferred over any
token-normalised measure because token fields are not present in this export.

**Available-to-focus hours** means working hours remaining after excluding
meetings and scheduled Teams calls. `Uninterrupted_hours` and
`Interrupted_hours` partition available-to-focus time in this source. The
working-hours basis is tenant-configurable; the 40-hour identity checked here is
specific to this synthetic dataset.

**Credit concentration** uses only coarse positive-user M365 quartiles, shared
with the credit-group panels. No individual curve, top-1% point, fine histogram,
observed extrema or individual ranking is published. If any quartile or the
no-observed-M365 group falls below the floor, the entire partition is withheld.

**Credit activity segments** are based on active product weeks and observed
M365 credit volume. They are not the standard `vivainsights` action-based segments,
because this source set does not include a Copilot actions metric.

**PeopleHistoricalId** is the documented join key between the Consumption
activity files and `PeopleMetaData`. It must be taken from the Consumption
export rather than derived from `PersonId`. A person-query export alone cannot
be joined to `PeopleMetaData`; the bridge is an activity file that carries both
`PersonId` and `PeopleHistoricalId`.

The dashboard shows associations. It does not establish that AI consumption
caused a working pattern, and it does not observe output quality, task
difficulty or business value. It also makes no wellbeing claims: uninterrupted,
interrupted, available-to-focus, after-hours and weekend measures describe
observed calendar and collaboration patterns only.

### Replacing the synthetic data

1. Replace the four CSVs listed above with schema-aligned governed exports.
2. Validate the identity crosswalk and disclose the match rate.
3. Require independent, week-relevant eligibility and ingestion/coverage evidence
   before zero filling. Missing rows alone do not establish zero activity or no
   provisioning; M365 enabled days do not establish GitHub provisioning. This
   synthetic report assumes zero for absent rows solely for fixture arithmetic,
   not because the export proves completeness. Do not reuse that assumption for
   real data.
4. Preserve the daily service grain until after service-level totals are needed;
   aggregate services before computing person-week rates.
5. Keep person-level averaging so each person carries equal analytical weight.
6. Retain the minimum of 10 and whole-partition disclosure gates, including
   contributor complements and shared gates for related tables/charts. Blank
   charts, empty tables and NA values mean withheld (or undefined rates), not
   zero. Overlapping releases require a governed cross-report disclosure review;
   these within-report checks are not a certification of arbitrary real data.
7. Keep M365 Copilot credits and GitHub AI credits in separate panels and
   denominators unless an authoritative common-unit/conversion contract is
   established.
