---
title: "Developer Experience and Copilot"
output:
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: scroll
    theme: lumen
    self_contained: true
    mathjax: null
    fig_mobile: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE,
  fig.width=11, fig.height=4.2, dpi=120, fig.retina=1, out.width='100%', fig.align='center')
source('github-developer-experience-helpers.R', local=knitr::knit_global())
```

<style>
:root { --primary:#4f46e5; --bg:#f8f9fa; --border:#e5e7eb; }
body { font-family:'Segoe UI',Arial,sans-serif; color:#253047; background:var(--bg); font-size:14px; }
.navbar { background:#fff; border-bottom:1px solid var(--border); }
.navbar-brand { color:#242742 !important; font-weight:650; font-size:17px; }
.navbar-nav>li>a { color:#475569 !important; font-size:13px; }
.navbar-nav>.active>a { background:#eef2ff !important; color:#3730a3 !important; }
.chart-wrapper { border:1px solid var(--border); border-radius:8px; box-shadow:0 1px 3px #18203308; margin-bottom:12px; background:white; height:auto !important; min-height:0 !important; }
.chart-title { font-size:15px; font-weight:600; padding:12px 18px; color:#303755; border-bottom:1px solid #f1f3f5; }
.dashboard-row { height:auto !important; }
.dashboard-page, .dashboard-row, .dashboard-column, .chart-wrapper, .chart-stage, .chart-shim { min-width:0 !important; max-width:100%; }
.chart-stage { height:auto !important; flex:0 0 auto !important; padding:14px 20px !important; overflow:auto !important; }
.chart-shim { position:relative !important; inset:auto !important; }
.image-container { position:static !important; background:none !important; }
.image-container img { opacity:1 !important; }
.chart-stage img { display:block; width:100%; height:auto; max-width:1320px; margin:0 auto; }
.chart-stage p { max-width:1100px; line-height:1.55; }
.kpis { display:grid; grid-template-columns:repeat(4,minmax(0,1fr)); gap:12px; margin:14px 0; }
.kpi { border:1px solid var(--border); border-radius:8px; padding:12px 16px; background:white; }
.kpi span,.kpi small { display:block; color:#475569; font-size:12px; }
.kpi strong { display:block; font-size:27px; font-weight:650; color:#242742; margin:3px 0; }
.pill { display:inline-block; border:1px solid #c7d2fe; border-radius:20px; padding:2px 9px; color:#3730a3; background:#eef2ff; font-size:12px; }
.leadline { font-size:18px; color:#20243b; max-width:1080px; line-height:1.5; }
.context { color:#526079; font-size:12px; margin:8px 0; }
.definitions { border:1px solid var(--border); background:#f8fafc; border-radius:8px; padding:12px 16px; margin:10px 0; line-height:1.55; }
details { margin:12px 0; border:1px solid var(--border); border-radius:8px; padding:10px 14px; }
summary { cursor:pointer; font-weight:600; color:#3730a3; }
details[open] summary { margin-bottom:12px; }
.table-scroll { overflow:auto; max-height:440px; margin:12px 0; }
table { width:100%; font-size:13px; font-variant-numeric:tabular-nums; border-collapse:collapse; }
th { position:sticky; top:0; background:#f1f5f9; z-index:1; color:#303755; }
th,td { padding:9px 12px !important; border-bottom:1px solid #e5e7eb; vertical-align:top; }
caption { text-align:left; color:#526079; font-size:12px; padding:8px 0; }
.framework { display:grid; grid-template-columns:repeat(3,minmax(0,1fr)); gap:12px; }
.framework>div { background:#f8fafc; border:1px solid var(--border); border-radius:8px; padding:14px; line-height:1.5; }
a { color:#4338ca; }
@media(min-width:1200px) { .dashboard-page-wrapper { max-width:1440px; margin:0 auto; } }
@media(max-width:767px) {
  body { padding-top:59px !important; }
  .kpis { grid-template-columns:repeat(2,minmax(0,1fr)); }
  .framework { grid-template-columns:1fr; }
  .chart-stage { padding:12px !important; }
  .chart-stage img { min-width:700px; }
}
</style>

<script>
document.addEventListener('click', function(event) {
  if (window.innerWidth < 768 && event.target.closest('.navbar-nav a')) {
    window.jQuery('#navbar').collapse('hide');
  }
});
</script>

1 Landscape
=====================================

Row
-------------------------------------

### Who is represented, and what does their working week look like?

<span class="pill">Synthetic demo · Real export schema</span>
<span class="context">Baseline: `r period_label` · `r BASELINE_WEEKS` complete Person Query weeks · Product feeds observed on weekdays in `r product_label`</span>

<p class="leadline">The baseline covers <strong>`r num(nrow(baseline))` developers across `r n_distinct(baseline$Team)` teams</strong>, with all eight Person Query weeks recorded for every developer. The median developer averaged <strong>`r f1(median(baseline$Collaboration_hours))` collaboration hours</strong>, including <strong>`r f1(median(baseline$Meeting_hours))` meeting hours</strong>, <strong>`r f1(median(baseline$Available_to_focus_hours))` available-to-focus hours</strong>, <strong>`r f1(median(baseline$Uninterrupted_hours))` uninterrupted hours</strong>, and <strong>`r f1(median(baseline$Interrupted_hours))` interrupted hours</strong> per week.</p>

```{r landscape-kpis, results='asis'}
cat('<div class="kpis">')
card('Developer roster', num(nrow(baseline)), paste(num(sum(people$IsDeveloper)), 'developers;', num(nrow(people)), 'people in source population'))
card('Collaboration hours / week', f1(median(baseline$Collaboration_hours)), 'Median of person-level eight-week averages')
card('Meeting hours / week', f1(median(baseline$Meeting_hours)), 'Median of person-level eight-week averages')
card('Available-to-focus hours / week', f1(median(baseline$Available_to_focus_hours)), 'Calendar availability; tenant working-hours rules apply')
card('Uninterrupted hours / week', f1(median(baseline$Uninterrupted_hours)), 'Partitioned with interrupted hours')
cat('</div>')
```

This is a manager view of working conditions and recorded Copilot use from committed synthetic CSVs whose column names match Viva Insights query exports. The roster contains `r num(nrow(people))` synthetic people, of whom `r num(sum(people$IsDeveloper))` are developers. Other job-family records are excluded by the `IsDeveloper` attribute, not by product activity. The helper does not generate or overwrite data. Microsoft 365 collection completeness is unknown by default: its complete-window rates and common-product comparisons are unavailable, not zero. Matching a fixture filename or schema never certifies coverage.

<div class="definitions">
<strong>Reading the working week.</strong> Collaboration hours include meeting hours, scheduled calls, email and chat activity, so they must not be added to meeting hours. Available-to-focus hours are the hours remaining during working hours after excluding meetings and scheduled Teams calls. Uninterrupted hours and interrupted hours partition available-to-focus hours. The working-hours basis is configurable per tenant through metric rules, so the 40-hour basis in this synthetic file is not universal. Available-to-focus hours are calendar availability; they do not measure coding time or psychological flow.
</div>

Row
-------------------------------------

### Team and role composition

```{r roster-composition, fig.height=4.4, fig.alt='Six team bars showing role composition. Every team contains 60 developers. Percentage labels and an exact table accompany the chart.'}
stack_plot(team_context, 'Roles represented in each team',
  'Roster composition provides context for coordination needs.',
  base_caption(unit='share of each team, 60 developers per team'))
```

<details><summary>Exact role mix, seniority and tenure</summary>

```{r composition-details, results='asis'}
html_table(composition_table(team_context), 'Role shares use the 60 developers in each team as denominator.')
html_table(composition_counts |> transmute(Attribute, Category, `Developers (count)`=People,
  `Share of developers (%)`=pct(People/nrow(baseline))), 'Role, seniority and tenure are illustrative HR attributes carried by the query exports.')
```

</details>

Row
-------------------------------------

### Team league table for workload and recorded Copilot use

<p class="leadline">Sorted by collaboration hours, <strong>`r team_league$Team[1]`</strong> has the highest median collaboration at <strong>`r f1(team_league$Collaboration_hours[1])` hours/person/week</strong>. Highest reportable GitHub recorded-active share: <strong>`r leader_label(team_league$Team, team_league$GH_active_share)`</strong>, <strong>`r pct(suppressWarnings(max(team_league$GH_active_share, na.rm=TRUE)))`</strong>. `r if (all(is.na(team_league$M365_active_share))) 'Microsoft 365 recorded-active share is unavailable for every team: Microsoft 365 collection completeness is not independently established by these exports, so the value is unresolved measurement, not zero use.' else paste0('Highest reportable Microsoft 365 recorded-active share: <strong>', leader_label(team_league$Team, team_league$M365_active_share), '</strong>, <strong>', pct(suppressWarnings(max(team_league$M365_active_share, na.rm=TRUE))), '</strong>.')` Unavailable means missing coverage evidence or privacy suppression, not inactivity.</p>

```{r landscape-tables, results='asis'}
html_table(team_league |>
  transmute(Rank, Team, `Developers (count)`=Developers,
    `Collaboration hours`=f1(Collaboration_hours), `Meeting hours`=f1(Meeting_hours),
    `Available-to-focus hours`=f1(Available_to_focus_hours),
    `Uninterrupted hours`=f1(Uninterrupted_hours), `Interrupted hours`=f1(Interrupted_hours),
    `After-hours collaboration`=f1(After_hours_collaboration_hours),
    `GitHub valid developers`=num(GH_valid_n), `GitHub active (%)`=pct_or_reason(GH_active_share, GH_valid_n),
    `GitHub credits / person / week`=f1(GH_intensity),
    `M365 valid developers`=num(M365_valid_n), `M365 active (%)`=pct_or_reason(M365_active_share, M365_valid_n),
    `M365 credits / person / week`=f1(M365_intensity)),
  paste('Sorted by collaboration hours descending | Hours/person/week and credits/person/week |', period_label,
        '| Microsoft 365 columns read 0 valid developers and an unavailable credit rate because no developer has independently established Microsoft 365 coverage in this window: that is unresolved measurement, not zero use.'))
html_table(work_summary, paste('Population percentiles across', num(nrow(baseline)), 'developers’ eight-week averages | Hours/person/week |', period_label))
```

Row
-------------------------------------

### Ranked team view of headline metrics

```{r team-league-visual, fig.height=12.2, fig.alt='Small-multiple ranked bars compare six teams on workload, available-to-focus time, uninterrupted time, interrupted time, after-hours collaboration and product-specific recorded use.'}
plot_league <- team_league_plot |>
  group_by(Metric) |>
  arrange(Value, .by_group=TRUE) |>
  mutate(Team_metric=factor(paste(Team, Metric, sep=' | '), levels=paste(Team, Metric, sep=' | '))) |>
  ungroup()
ggplot(plot_league, aes(x=Value, y=Team_metric)) +
  geom_col(data=filter(plot_league, is.finite(Value)), fill='#2563eb', width=.62) +
  geom_text(data=filter(plot_league, is.finite(Value)),
    aes(label=Label), hjust=-.10, size=3.1, colour='#253047') +
  geom_text(data=filter(plot_league, !is.finite(Value)),
    aes(x=0, label='Unavailable / withheld'), hjust=0, size=3.1, colour='#526079') +
  facet_wrap(~Metric, scales='free', ncol=2, labeller=label_wrap_gen(26)) +
  scale_y_discrete(labels=function(x) sub(' \\| .*', '', x)) +
  scale_x_continuous(expand=expansion(mult=c(0,.25))) + theme_report() +
  chart_labs('Ranked team metrics for the baseline window',
    'Each panel is ranked independently. Product denominators use developers with valid observation across the full baseline window.',
    paste(period_label, '| 60 developers per team | Product intensity uses credits, not fabricated action counts.'),
    x=NULL, y=NULL)
```

**Manager next step.** Use the team and role composition when discussing the distribution of working hours. Agree which coordination or tooling question needs investigation before setting a target. No performance judgement or individual action is supported by these data.

2 Focus and coordination
=====================================

Row
-------------------------------------

### How much sustained work appears alongside coordination?

<span class="pill">Synthetic demo · Real export schema</span>
<span class="context">`r period_label` · All `r num(nrow(baseline))` developers · `r BASELINE_WEEKS` complete Person Query weeks each</span>

<p class="leadline"><strong>`r team_league$Team[which.max(team_league$Meeting_hours)]`</strong> has the highest median meeting load at <strong>`r f1(max(team_league$Meeting_hours))` hours/person/week</strong>. <strong>`r team_league$Team[which.max(team_league$Available_to_focus_hours)]`</strong> has the most available-to-focus time at <strong>`r f1(max(team_league$Available_to_focus_hours))` hours/person/week</strong>, and <strong>`r team_league$Team[which.max(team_league$Uninterrupted_hours)]`</strong> has the most uninterrupted time at <strong>`r f1(max(team_league$Uninterrupted_hours))` hours/person/week</strong>. These are calendar and collaboration associations, not evidence of coding time or delivery quality.</p>

<div class="definitions">
<strong>Definitions, paraphrased from Microsoft Learn.</strong> <strong>Available-to-focus hours</strong> are hours remaining during working hours after excluding meetings and scheduled Teams calls for focused work. <strong>Uninterrupted hours</strong> are one-hour-or-longer blocks of uninterrupted time. <strong>Interrupted hours</strong> are available-to-focus time interrupted by email, Teams chat, unscheduled calls or Teams channel activity. <strong>Open 1-hour block</strong> is a calendar count without scheduled meetings during the workday. The working-hours basis is set by tenant metric rules; the 40-hour basis in this synthetic data is not universal.
</div>

Row
-------------------------------------

### Meeting, available-to-focus, uninterrupted and interrupted-time distributions

```{r focus-intervals, fig.height=4.4, fig.alt='Team percentile intervals compare meeting hours, available-to-focus hours, uninterrupted hours and interrupted hours.'}
interval_plot(baseline, c('Meeting_hours','Available_to_focus_hours','Uninterrupted_hours','Interrupted_hours'),
  title='Meeting, available-to-focus, uninterrupted and interrupted-time distributions by team',
  subtitle='Dot: median. Line: 25th to 75th percentile of person-level eight-week averages. Separate horizontal scales.')
```

Row
-------------------------------------

### Meeting characteristics that may warrant a conversation

```{r meeting-characteristics, results='asis'}
characteristics <- interval_data(baseline, c('Recurring_meeting_hours','Conflicting_meeting_hours','Meeting_hours_with_six_or_fewer_hours_of_advanced_notice')) |>
  group_by(Metric) |> arrange(desc(p50), .by_group=TRUE) |> ungroup() |>
  transmute(Team=Group, `Meeting characteristic`=Metric, `25th percentile`=f1(p25), Median=f1(p50), `75th percentile`=f1(p75))
html_table(characteristics, paste('Hours/person/week |', period_label, '| 60 developers per team. Characteristics overlap and must not be added.'))
```

<div class="definitions">
<strong>Recurring meeting hours:</strong> meetings set to recur. <strong>Conflicting meeting hours:</strong> only the overlapping portion of calendar meetings. <strong>Short-notice meeting hours:</strong> meetings scheduled six hours or less before their start. These are separate, potentially overlapping characteristics, not components of an additive stack.
</div>

**Manager next step.** Use these distributions to ask which meetings unblock work and which interrupt it. Combine the answer with feedback on review wait times and cognitive load before testing a calendar change.

3 Sustainable workload
=====================================

Row
-------------------------------------

### How is collaboration outside working hours distributed and repeated?

<span class="pill">Synthetic demo · Real export schema</span>
<span class="context">`r period_label` · All `r num(nrow(baseline))` developers · `r BASELINE_WEEKS` observed weeks each</span>

<p class="leadline"><strong>`r team_league$Team[which.max(team_league$After_hours_collaboration_hours)]`</strong> has the highest median after-hours collaboration at <strong>`r f1(max(team_league$After_hours_collaboration_hours))` hours/person/week</strong>. The population median is <strong>`r f1(median(baseline$After_hours_collaboration_hours))` hours</strong>, and the upper quartile begins at <strong>`r f1(quantile(baseline$After_hours_collaboration_hours,.75))` hours</strong>. These describe recorded collaboration outside scheduled hours, with no wellbeing outcome available.</p>

<div class="definitions">
<strong>After-hours collaboration, paraphrased from Microsoft Learn:</strong> time in meetings, email, Teams chats, calls and channels with at least one other person outside working hours, with overlapping activity deduplicated. It is not total work time. Flexible schedules, on-call duties and time zones may matter, but none is recorded here.
</div>

Row
-------------------------------------

### After-hours collaboration across teams

```{r afterhours-distribution, fig.height=4.2, fig.alt='Six team medians and interquartile ranges of after-hours collaboration hours per developer per week.'}
interval_plot(baseline, 'After_hours_collaboration_hours',
  title='After-hours collaboration distribution by team',
  subtitle='Dot: median. Line: 25th to 75th percentile of each developer’s eight-week average. No health bands are assigned.')
```

Row
-------------------------------------

### Persistence over the observed weeks

```{r persistence, fig.height=4.3, fig.alt='Bar chart of developers with zero, one to three, or four to eight weeks meeting an illustrative after-hours convention. Counts and percentages are labelled.'}
persistence <- baseline |> mutate(Band=case_when(Weeks_above==0 ~ '0 weeks',
  Weeks_above<PERSISTENCE_WEEKS ~ paste('1 to', PERSISTENCE_WEEKS-1, 'weeks'),
  TRUE ~ paste(PERSISTENCE_WEEKS, 'to', BASELINE_WEEKS, 'weeks'))) |>
  count(Band, name='People') |> mutate(Share=People/nrow(baseline))
stopifnot(sum(persistence$People)==nrow(baseline))
persistence <- disclose_partition(persistence)
if (!nrow(persistence)) unavailable_plot('Persistence composition withheld') else
ggplot(persistence, aes(x=People, y=factor(Band, levels=rev(c('0 weeks', paste('1 to', PERSISTENCE_WEEKS-1, 'weeks'), paste(PERSISTENCE_WEEKS, 'to', BASELINE_WEEKS, 'weeks')))))) +
  geom_col(fill='#c7d2fe', width=.55) +
  geom_text(aes(label=paste0(num(People), ' (', pct(Share), ')')), hjust=-.12, size=3.7) +
  scale_x_continuous(expand=expansion(mult=c(0,.23))) + theme_report() +
  chart_labs('How often the illustrative convention was met',
    paste('Convention:', AFTER_HOURS_CONVENTION, 'or more after-hours collaboration hours in a week. No burnout diagnosis is implied.'),
    base_caption(unit=paste('developer counts and share of', num(nrow(baseline)))), x='Developers (count)', y='Weeks at or above convention')
```

**Manager next step.** Before drawing a workload conclusion, validate work-schedule settings and seek anonymous feedback on sustainability and disconnection. Do not target individual developers from these aggregates.

4 AI use and query coverage
=====================================

Row
-------------------------------------

### What is observed, eligible and missing?

<span class="pill">Synthetic demo · Real export schema</span>
<span class="context">`r period_label` · same `r num(nrow(baseline))`-developer roster · GitHub rows distinguish zero from missing</span>

<p class="leadline">Common complete-window observation is unavailable without independent Microsoft 365 collection evidence. GitHub activity rows establish observation, not licensing or provisioning; absence leaves status unknown. Microsoft 365 eligibility uses that week's <code>Total_Copilot_enabled_days</code>: static metadata cannot override a zero-enabled week. The product windows deliberately start later than the Person Query history.</p>

<div class="definitions">
<strong>Zero versus missing, and two different coverage questions.</strong> The synthetic GitHub activity file includes explicit zero rows for inactive weekdays: these are measured zero, whereas an absent row leaves status unknown. The GitHub credit file is <strong>sparse</strong> — it carries a row only where billable usage occurred — so credit coverage is resolved separately from activity coverage and never invalidates an observed activity week. Microsoft 365 consumption rows cannot certify their own collection completeness. The product calendar and static licence metadata are not completeness evidence. Missing Microsoft 365 values stay <code>NA</code> unless independent person-week coverage and positive enabled days justify a zero; the default report has no such evidence, which is why every Microsoft 365 complete-window measure on this page and in the team league table reads as unavailable rather than zero.
</div>

Row
-------------------------------------

### Joint footprint with privacy-safe exceptions retained

```{r joint-footprint, fig.height=3.8, fig.alt='A labelled stacked bar shows product-use and observation status only when every positive cell meets the privacy floor.'}
joint_stack <- joint_counts_public |> transmute(Group='Developer roster', Category=Display, People, Share)
stack_plot(joint_stack, 'Recorded product use and eligibility or observation status',
  'Only and neither require established eligibility and observation. Any sub-10 cell withholds the entire partition.',
  base_caption(unit=paste('share of', num(nrow(baseline)), 'developers')))
```

<details><summary>Product denominators and coverage derivation</summary>

```{r joint-details, results='asis'}
html_table(joint_counts_public |> transmute(`Published tool-use status`=Display, `Developers (count)`=People, `Share of roster (%)`=pct(Share)),
  'The entire partition is withheld if any positive cell is below the privacy floor; totals cannot reveal a withheld remainder.')
eligibility_table <- bind_rows(lapply(c('GH','M365'), function(p) {
  valid_col <- if (p=='GH') 'GH_activity_valid' else 'M365_valid'
  b <- baseline_panel |> group_by(PersonId) |>
    summarise(eligible=all(.data[[paste0(p,'_eligible')]]),
              valid=all(.data[[valid_col]]), .groups='drop') |>
    mutate(Status=case_when(
      is.na(eligible) ~ 'Eligibility or observation unknown',
      !eligible ~ 'Not eligible for the full window',
      !valid ~ 'Eligible; completeness unresolved or incomplete',
      TRUE ~ 'Eligible and valid'))
  b |> count(Status, name='People') |> disclose_partition() |>
    mutate(Product=if(p=='GH') 'GitHub Copilot' else 'Microsoft 365 Copilot', .before=1)
}))
html_table(eligibility_table, 'Each reportable product partition reconciles to the roster. An unsafe partition is withheld in full, including its complementary cells. GitHub validity here is activity-row coverage; credit coverage is resolved separately below.')
html_table(baseline_panel |> group_by(PersonId) |>
  summarise(activity=all(GH_activity_valid), credit=all(GH_credit_valid), .groups='drop') |>
  count(`GitHub activity coverage`=if_else(activity,'Resolved','Unresolved'),
        `GitHub credit coverage`=if_else(credit,'Resolved','Unresolved'), name='People') |>
  disclose_partition() |>
  transmute(`GitHub activity coverage`, `GitHub credit coverage`, `Developers (count)`=People),
  'The credit export is sparse, so a week with no billable usage resolves to a measured zero rather than invalidating the observed activity week. Sparse credit rows never void observed activity.')
html_table(coverage_derivation, 'No standalone coverage export is invented. Weekly eligibility, activity coverage and credit coverage are separate requirements.')
```

</details>

Row
-------------------------------------

### Product-specific frequency, intensity and services

```{r product-usage, results='asis'}
html_table(product_summary |> arrange(desc(`Recorded active (%)`)) |> mutate(across(-Product, as.character)) |>
  pivot_longer(-Product, names_to='Product measure', values_to='Reported value') |>
  pivot_wider(names_from=Product, values_from='Reported value'),
  paste(period_label, '| Each product uses its own eligible and valid population. People with no recorded use remain in the frequency and intensity denominator. Microsoft 365 rows are unavailable because no developer has independently established Microsoft 365 coverage; GitHub credit measures use the credit-resolved population, which is derived separately from activity coverage.'))
html_table(m365_service_mix |> transmute(Service=ServiceName, `People (count)`=People,
  `Sessions`=num(Sessions), `Credits`=num(Credits), `Credit share (%)`=pct(Share)),
  'Recorded Microsoft 365 service rows over baseline weekdays. These observed sums and shares do not establish complete collection or complete-window population rates.')
```

The concentration row is the share of GitHub credits or Microsoft 365 credits attributable to the highest-volume 10% of active developers for that product. Identities and individual rankings are never displayed. Credits, sessions, completion suggestions and chat requests are different units and are not added together.

5 GitHub Copilot breakdowns
=====================================

Row
-------------------------------------

### What GitHub Copilot features, models and languages appear in the real breakdown files?

<span class="pill">Synthetic demo · Real export schema</span>
<span class="context">Daily GitHub breakdowns · `r period_label` weekdays · synthetic allocation, not verified real-export semantics</span>

<p class="leadline">In this synthetic fixture only, the breakdown files are daily margins of a shared Feature × Language × Model allocation. Largest reportable feature: <strong>`r leader_label(feature_mix$Feature, feature_mix$Share)`</strong>, <strong>`r pct(feature_mix$Share[1])`</strong>. Largest reportable model: <strong>`r leader_label(model_mix$Model, model_mix$Share)`</strong>, <strong>`r pct(model_mix$Share[1])`</strong>. These illustrative shares do not validate real-export model attribution.</p>

<div class="definitions">
<strong>Synthetic allocation invariant.</strong> The generator constructs all four breakdown totals from accepted completions plus user-initiated chats and checks its own internal consistency. This is not an authoritative real-export metric definition, and the reader does not reject inputs whose totals differ. Model attribution is illustrative, including attribution of code completions to models. An exact authoritative metric-definition source is required before enforcing these equalities or interpreting completion-model allocation on real data. <code>Share</code> is calculated in this report within each supplied breakdown; it is not an export column.
</div>

Row
-------------------------------------

### Feature, model and language shares

```{r github-margins, fig.height=8.4, fig.alt='Three panels show GitHub Copilot feature, model and language shares over the baseline window.'}
margin_plot <- bind_rows(
  feature_mix |> transmute(Margin='Feature', Category=Feature, Share),
  model_mix |> transmute(Margin='Model', Category=Model, Share),
  language_mix |> transmute(Margin='Language', Category=Language, Share)) |>
  group_by(Margin) |> slice_max(Share, n=12, with_ties=FALSE) |>
  arrange(Share, .by_group=TRUE) |>
  mutate(Category_margin=factor(paste(Category, Margin, sep=' | '), levels=paste(Category, Margin, sep=' | '))) |>
  ungroup()
if (!nrow(margin_plot)) unavailable_plot('GitHub breakdowns unavailable or withheld') else
ggplot(margin_plot, aes(x=Share, y=Category_margin)) +
  geom_col(fill='#2563eb', width=.65) +
  geom_text(aes(label=pct(Share)), hjust=-.1, size=3.1) +
  facet_wrap(~Margin, scales='free_y', ncol=1) +
  scale_y_discrete(labels=function(x) sub(' \\| .*', '', x)) +
  scale_x_continuous(labels=label_percent(), expand=expansion(mult=c(0,.18))) +
  theme_report() +
  chart_labs('GitHub Copilot usage margins from daily breakdown exports',
    'All linked margins and cross-tabs are withheld if any cell has fewer than 10 people.',
    paste(period_label, '| Shared allocation and completion-model attribution are synthetic illustrations only.'),
    x='Share of the supplied breakdown usage', y=NULL)
```

Row
-------------------------------------

### Cross-tabs: model × feature and language × model

```{r github-crosstabs, results='asis'}
top_model_feature <- model_feature_mix |>
  arrange(desc(Count)) |>
  slice_head(n=15) |>
  transmute(Model, Feature, `People (count)`=People, `Usage count`=num(Count), `Share within model (%)`=pct(Share))
top_language_model <- language_model_mix |>
  arrange(desc(Count)) |>
  slice_head(n=15) |>
  transmute(Language, Model, `People (count)`=People, `Usage count`=num(Count), `Share within language (%)`=pct(Share))
html_table(top_model_feature, paste('Largest reportable Model x Feature cells |', period_label, '| Daily model-feature export.'))
html_table(top_language_model, paste('Largest reportable Language x Model cells |', period_label, '| Daily language-model export.'))
```

<details><summary>Language × feature detail</summary>

```{r language-feature-detail, results='asis'}
html_table(language_feature_mix |>
  arrange(desc(Count)) |>
  slice_head(n=20) |>
  transmute(Language, Feature, `People (count)`=People, `Usage count`=num(Count), `Share within language (%)`=pct(Share)),
  paste('Largest reportable Language x Feature cells |', period_label, '| Daily language-feature export.'))
```

</details>

**Manager next step.** Use these breakdowns to ask which workflows each feature supports. Do not treat agent-type features or model choice as maturity, quality, productivity or time-saved measures.

6 Working patterns and evaluation
=====================================

Row
-------------------------------------

### How do working conditions differ across recorded tool-use groups?

<span class="pill">Synthetic demo · Real export schema</span>
<span class="context">`r period_label` · Comparison population: `r num(nrow(gh_observed))` developers with complete GitHub activity coverage · Same eight-week baseline</span>

<p class="leadline">The GitHub activity export resolves its own observation from explicit weekday zero rows, so working conditions can be compared across <strong>`r num(sum(gh_observed$GH_use == 'GitHub use recorded'))` developers with recorded GitHub use</strong> and <strong>`r num(sum(gh_observed$GH_use == 'No GitHub use recorded'))` observed developers with none</strong>. A <em>two-product</em> comparison remains unavailable by default, because Microsoft 365 completeness is not established by these exports and missing coverage is not recorded non-use. If independent evidence is supplied during an adaptation, compare only people observed throughout the same window and apply the privacy floor to every group. All of these comparisons remain descriptive, not causal.</p>

<div class="definitions">
<strong>Comparison population:</strong> developers valid for the Person Query and for GitHub activity coverage throughout the baseline. The `r num(sum(baseline$GH_use == 'GitHub observation unresolved'))` developers whose GitHub observation is unresolved stay in the `r num(nrow(baseline))`-person landscape and are excluded here rather than being recoded as having no use. This is not statistical matching. Group membership uses the same window as working conditions, so temporal direction cannot be established.
</div>

Row
-------------------------------------

### Selected working-condition distributions

```{r group-intervals, fig.height=4.4, fig.alt='Medians and interquartile ranges for collaboration, meeting, available-to-focus, uninterrupted and after-hours collaboration across reportable recorded GitHub-use groups.'}
interval_plot(gh_observed, c('Collaboration_hours','Meeting_hours','Available_to_focus_hours','Uninterrupted_hours','After_hours_collaboration_hours'), group='GH_use',
  title='Working-condition distributions by recorded GitHub Copilot use',
  subtitle='Dot: median. Line: middle half of each reportable group’s person-level averages. Separate scales. Differences are unadjusted.')
```

<details><summary>Exact reportable group sizes and percentile values</summary>

```{r group-percentiles, results='asis'}
sizes <- gh_observed |> count(GH_use, name='People', .drop=TRUE) |> filter(People >= MIN_GROUP_N)
html_table(interval_data(gh_observed, c('Collaboration_hours','Meeting_hours','Available_to_focus_hours','Uninterrupted_hours','After_hours_collaboration_hours'), 'GH_use') |>
  left_join(sizes, by=c('Group'='GH_use'), relationship='many-to-one') |>
  filter(!is.na(People)) |>
  group_by(Metric) |> arrange(desc(p50), .by_group=TRUE) |> ungroup() |>
  transmute(Group, Metric, `Developers (count)`=People, `25th percentile`=f1(p25), Median=f1(p50), `75th percentile`=f1(p75)),
  paste('Hours/person/week |', period_label, '| Each person contributes one eight-week average. Sub-10 use groups are withheld.'))
```

</details>

Row
-------------------------------------

### Role and team context for the comparison

```{r role-group-context, fig.height=3.8, fig.alt='Role composition within reportable recorded GitHub-use groups, with percentage labels and an exact table.'}
stack_plot(role_context, 'Role composition within the comparison groups',
  'Group sizes and roles can shape the comparison. Any group with a role cell below 10 is omitted from this breakdown.',
  base_caption(sum(role_context$People), unit='share of developers within each displayed use group'))
```

<details><summary>Exact role shares and team composition</summary>

```{r group-context-tables, results='asis'}
html_table(composition_table(role_context), 'A role breakdown is withheld in full for any group containing a cell below 10 people.')
html_table(composition_table(team_joint_context), 'A team breakdown is withheld in full for any group containing a cell below 10 people. Shares use all developers in the displayed group.')
```

</details>

**Interpretation.** Team, role, seniority, task selection and tenure can influence working patterns and tool use. These charts are intentionally limited to agreed measures. No broad metric scan, adjustment model, causal claim or preferred tool-use group is presented.

Row
-------------------------------------

### Weekly trends and valid denominators

```{r product-trends, fig.height=4.2, fig.alt='Weekly trends show collaboration medians and product recorded-active rates with valid denominators.'}
trend |> select(Week, GH_rate, M365_rate) |>
  pivot_longer(-Week, names_to='Product', values_to='Rate') |>
  mutate(Product=recode(Product, GH_rate='GitHub Copilot', M365_rate='Microsoft 365 Copilot')) |>
  filter(!is.na(Rate)) |>
  ggplot(aes(Week, Rate, colour=Product)) + geom_line(linewidth=.9) +
  scale_colour_manual(values=c('#2563eb','#0f766e')) +
  scale_y_continuous(labels=label_percent(), limits=c(0,1)) +
  scale_x_date(date_breaks='1 month', date_labels='%d %b') + theme_report() +
  chart_labs('Recorded active developers as a share of valid weekly observation',
    'At least one positive activity day that week. Missing coverage is not inactivity; Microsoft 365 rates are unavailable by default without independent completeness evidence.',
    paste(history_label, '| Denominator: product-eligible developers with valid observation in that week. UTC week starts.'),
    x='Week beginning (2026)', y='Recorded active (%)')
```

<details><summary>Coverage and numerators for every weekly window</summary>

```{r weekly-coverage, results='asis'}
html_table(trend |> transmute(`Week beginning (UTC)`=format(Week,'%Y-%m-%d'),
  `PQ developers (count)`=num(PQ_n), `GH valid (count)`=num(GH_n), `GH active (count)`=num(GH_active),
  `GH active (%)`=pct_or_reason(GH_rate, GH_n), `M365 valid (count)`=num(M365_n),
  `M365 active (count)`=num(M365_active), `M365 active (%)`=pct_or_reason(M365_rate, M365_n),
  `Both valid (count)`=num(Joint_n)),
  'A count of 0 is an observed absence of valid weekly observation, not a withheld value: the GitHub feed opens 13 weeks after the Person Query history, and Microsoft 365 has no valid weeks at all because its completeness is not independently established. A rate over an empty denominator is undefined rather than withheld.')
```

</details>

Row
-------------------------------------

### Evaluate a documented change with developer feedback and delivery outcomes

<div class="framework">
<div><strong>SPACE: multidimensional productivity</strong><p>Consider satisfaction and wellbeing, performance, activity, communication and collaboration, and efficiency and flow together. This report contributes activity and working-condition signals. It cannot establish overall productivity or wellbeing.</p></div>
<div><strong>DevEx: feedback loops, cognitive load and flow</strong><p>Combine developer feedback with system measures. Add perceived review delays, cognitive load and ability to focus alongside measured build and review waits. Calendar-derived uninterrupted time cannot establish a psychological flow state.</p></div>
<div><strong>DORA: delivery at service level</strong><p>Pair speed and delivery quality for one application or service. Add change lead time, deployment frequency, failed deployment recovery time, change fail rate and deployment rework rate. None is available in the current sources.</p></div>
</div>

**Decision supported now.** Select a workflow question and close its measurement gaps. An intervention effect, productivity gain or wellbeing conclusion requires additional evidence. No operational action is justified by synthetic values.

Appendix
=====================================

Row
-------------------------------------

### Definitions, sources and reporting contracts

<span class="pill">Synthetic demo · Real export schema</span>

**Scope and units.** Pages use the last eight complete Person Query weeks, `r period_label`, and product weekdays that fall in the final 13-week product window. The full Person Query source window is `r history_label`, with Sunday-start weeks. Dates are ISO dates interpreted using a UTC convention. This is an illustrative date contract, not evidence about developers’ time zones. Available-to-focus hours are based on working hours configured through tenant metric rules; the 40-hour basis in this synthetic file is not universal.

**Aggregation.** Working-condition values are weekly Person Query metrics. Baseline charts first average each person’s eight weekly values and then compute group percentiles. Each developer has equal weight. Product-active rates divide active eligible, observed developers by all eligible, observed developers for that product and window. Daily Microsoft 365 service rows are aggregated before person-day and person-week analysis.

**Source definitions.** Metric descriptions are paraphrased from the [Microsoft Viva Insights metrics reference](https://learn.microsoft.com/en-us/viva/insights/advanced/reference/metrics). The source manifest is `_data/README.md`; the report reads the committed CSVs listed there.

```{r source-contracts, results='asis'}
html_table(source_contracts, 'CSV files read by github-developer-experience-helpers.R. No zip, external download or data-generation step is run by this report.')
```

**Joins and missingness.** Person Query anchors the roster and weekly panel. Consumption metadata is keyed by `PeopleHistoricalId`, which is an opaque key read from the activity-file crosswalk and never reconstructed from `PersonId`; the Person Query is used for HR attributes because it already carries `PersonId`. Daily feeds aggregate to unique person-week keys before left joins. One-to-one and many-to-one join assertions prevent row amplification. `0` is an observed or justified zero. `NA` is unresolved measurement and is never silently replaced outside valid observation. GitHub activity coverage and GitHub credit coverage are separate flags, each gating only the measures it governs.

Row
-------------------------------------

### Privacy, reproducibility and validation

**Privacy floor.** At least 10 distinct people are required for a published group. Percentile visuals use `vivainsights::create_boxplot(..., mingroup = 10, return = "table")`. Linked HR margins and cross-tabs share one release gate: every cell of every published margin must clear the floor, so no published margin can be differenced against another to expose a sub-floor group. Joint-use and eligibility partitions are withheld in full when unsafe. Linked GitHub margins and cross-tabs are withheld together, and their release additionally requires that no person's pattern of cell membership is shared by fewer than 10 people. Service releases check positive contributors and overlapping membership-pattern complements within the developer roster. Product rates check both numerator and complement; unavailable values are labelled, not plotted as zero. Missing category labels are shown as “Unknown / not supplied”. No individual points, identifiers or individual leaderboards appear in the HTML.

**Simulation boundary.** The CSV values are synthetic, but the report no longer invents CSV schemas or writes files. It does not use a standalone eligibility or coverage export, fabricated Microsoft 365 action categories, model or language share files, or an agent-use indicator that is absent from the real GitHub activity export. It parses the real `Agent adoption` field and derives all shares locally.

**Coverage boundary.** There is no default fixture-completeness exception or filename-based trust. `derive_m365_coverage(person_weeks, evidence = NULL)` returns unknown completeness unless a caller explicitly supplies independently obtained, person-week-specific evidence with its source. That is an in-memory analysis input, not a fabricated export column or file. A calendar, static licence status, observed positive rows, or missing rows cannot supply that evidence. A future fixture-only opt-in would need validated provenance tied to the exact files; assumptions must not survive replacement or row deletion.

```{r validation-table, results='asis'}
html_table(validation_summary, 'Automated data-contract checks executed during this render.')
```

<details><summary>Runtime versions and package choice</summary>

```{r versions, results='asis'}
html_table(tibble(Component=c('R','dplyr','tidyr','ggplot2','vivainsights','rmarkdown','flexdashboard'),
  Version=c(as.character(getRversion()), vapply(c('dplyr','tidyr','ggplot2','vivainsights','rmarkdown','flexdashboard'),
    function(p) as.character(packageVersion(p)), character(1)))))
```

The [vivainsights function index](https://microsoft.github.io/vivainsights/llms.txt) supports the person-averaged boxplot summary used for percentile calculations. Custom joint coverage aggregation is necessary because single-product usage segments do not establish comparable observation for two products.

</details>

Row
-------------------------------------

### Research framework and future extensions

| Framework | What this report can contribute | Evidence still needed |
|---|---|---|
| SPACE | Activity, collaboration and calendar-derived working conditions across multiple dimensions | Satisfaction and wellbeing surveys, performance outcomes, direct evidence of efficiency and flow |
| DevEx | Context for feedback loops, cognitive load and flow discussions | Anonymous developer feedback, build/test waits, review waits, task context and perceived ability to focus |
| DORA | A framework for evaluating delivery change, without an outcome claim | Service boundaries, production deployments, change lead times, recovery events, failed changes and deployment rework |

- [The SPACE of Developer Productivity](https://queue.acm.org/detail.cfm?id=3454124), ACM Queue, 2021. Productivity is multidimensional and cannot be reduced to one activity measure.
- [DevEx: What Actually Drives Productivity](https://queue.acm.org/detail.cfm?id=3595878), ACM Queue, 2023. Connect feedback loops, cognitive load and flow with developer feedback and system measures.
- [DORA’s software delivery performance metrics](https://dora.dev/guides/dora-metrics/). Current guidance describes five metrics across throughput and instability, best applied to one application or service at a time.
- [Viva Insights metrics reference](https://learn.microsoft.com/en-us/viva/insights/advanced/reference/metrics). Primary reference for the paraphrased collaboration and focus definitions.
