---
title: "GitHub Copilot and Developer Productivity"
subtitle: "The art of the possible with GitHub Copilot and Viva Insights data (fully synthetic)"
date: "Synthetic window: 4 Jan-28 Jun 2026"
output:
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: scroll
    theme: lumen
    self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.width = 8.5, fig.height = 4.5)
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
library(knitr)
library(vivainsights)
library(ggrepel)
library(lubridate)

set.seed(2026)

MIN_GROUP_N <- 10L   # minimum distinct people before a cohort is reported

PAL  <- c("#1F4E79", "#2E75B6", "#7FA8D0", "#BDD3E8", "#C9C9C9",
          "#B03A2E", "#5B8C5A", "#7D5BA6")
PAL5 <- c(PAL[1], PAL[2], PAL[3], "#A9BFA0", PAL[5])

theme_t <- function() {
  theme_minimal(base_size = 11) +
    theme(
      panel.grid.minor = element_blank(),
      legend.position = "bottom",
      legend.title = element_blank(),
      plot.title.position = "plot",
      plot.title = element_text(face = "bold", colour = PAL[1]),
      plot.subtitle = element_text(colour = "grey30",
                                   margin = margin(t = 2, b = 9)),
      plot.caption = element_text(colour = "grey45", size = 8, hjust = 0,
                                  margin = margin(t = 7)),
      plot.margin = margin(t = 10, r = 16, b = 4, l = 4)
    )
}

# Standardised difference between two vectors, used throughout for comparability
# across metrics on different scales.
std_diff <- function(a, b) {
  s <- sqrt(((length(a) - 1) * var(a) + (length(b) - 1) * var(b)) /
              (length(a) + length(b) - 2))
  if (is.na(s) || s == 0) return(NA_real_)
  (mean(a) - mean(b)) / s
}
```

```{r simulate-people}
N_DEV   <- 900L
N_OTHER <- 180L   # non-developers who also touch repositories
N_WEEKS <- 26L
weeks   <- seq(as.Date("2026-01-04"), by = "week", length.out = N_WEEKS)
days    <- seq(min(weeks), max(weeks) + 6, by = "day")

teams  <- c("Platform", "Payments", "Mobile", "Data Engineering",
            "Developer Experience", "Identity")
TEAM_PAL <- setNames(PAL[c(1, 2, 3, 7, 4, 8)], teams)
tenure_levels <- c("Under 2 years", "2 to 5 years", "Over 5 years")

people_all <- tibble(
  PersonId = sprintf("P%05d", seq_len(N_DEV + N_OTHER)),
  # Non-developers sit in the same teams and appear in the GitHub extract when
  # they review code, which is exactly why tool presence cannot define the
  # developer population.
  JobFamily = c(rep("Software Engineering", N_DEV),
                sample(c("Product Management", "Design", "Technical Writing"),
                       N_OTHER, replace = TRUE, prob = c(0.55, 0.30, 0.15))),
  Team   = sample(teams, N_DEV + N_OTHER, replace = TRUE,
                  prob = c(0.22, 0.18, 0.16, 0.18, 0.12, 0.14)),
  Tenure = factor(sample(tenure_levels, N_DEV + N_OTHER, replace = TRUE,
                         prob = c(0.34, 0.42, 0.24)), levels = tenure_levels)
) |>
  mutate(
    IsDeveloper = JobFamily == "Software Engineering",
    # A licence is not universal, and an unlicensed person is absent from the
    # activity export for a completely different reason than an inactive one.
    IsCopilotLicensed = rbinom(n(), 1, if_else(IsDeveloper, 0.93, 0.55)) == 1,
    tenure_z = as.numeric(Tenure) - 2,
    # Tenure raises meeting load and AI appetite at the same time. This is the
    # planted confounder.
    meeting_latent = 0.55 * tenure_z + rnorm(n(), 0, 0.85),
    ai_latent      = 0.30 * tenure_z + rnorm(n(), 0, 0.90),
    # Preference for conversational rather than inline assistance, and for
    # agentic flows. Independent of how much AI a person uses in total.
    chat_pref  = rnorm(n(), 0, 1),
    agent_pref = rnorm(n(), 0, 1),
    # Independent person-level drivers for the metrics that carry no injected
    # relationship with AI use. Keeping these separate from meeting_latent is
    # what allows the nulls in the effect register to actually come out null.
    conflict_latent = rnorm(n(), 0, 1),
    genload_latent  = rnorm(n(), 0, 1),
    afterhours_latent = rnorm(n(), 0, 1),
    span_latent     = rnorm(n(), 0, 1),
    net_int_latent  = rnorm(n(), 0, 1),
    net_ext_latent  = rnorm(n(), 0, 1),
    # Share latents so the meeting sub-components vary independently rather
    # than moving as one block.
    s_recurring = plogis(rnorm(n(), 0.1, 0.6)),
    s_large     = plogis(rnorm(n(), -1.4, 0.6)),
    s_small     = plogis(rnorm(n(), -0.8, 0.6)),
    s_notice    = plogis(rnorm(n(), -1.9, 0.7)),
    s_selforg   = plogis(rnorm(n(), -1.3, 0.7)),
    # A licensed person who never engages. This is what creates attrition
    # between the licensed population and the activity export.
    never_uses = rbinom(n(), 1, plogis(-1.7 - 0.9 * ai_latent)) == 1,
    p_active_day = if_else(never_uses, 0, plogis(0.30 + 0.75 * ai_latent))
  )

devs <- filter(people_all, IsDeveloper)
```

```{r simulate-person-query}
# Weekly Person Query for the whole population. Meeting load and focus time are
# generated first, because the AI activity below is written to depend on them
# rather than the reverse.
pq <- expand_grid(PersonId = people_all$PersonId, MetricDate = weeks) |>
  left_join(select(people_all, PersonId, Team, Tenure, JobFamily, IsDeveloper,
                   IsCopilotLicensed, meeting_latent, tenure_z,
                   conflict_latent, genload_latent, afterhours_latent,
                   span_latent, net_int_latent, net_ext_latent,
                   s_recurring, s_large, s_small, s_notice, s_selforg),
            by = "PersonId") |>
  mutate(
    wk_noise = rnorm(n(), 0, 1),

    # --- Meeting load -----------------------------------------------------
    Meeting_hours = pmax(0.5, 7.4 + 2.5 * meeting_latent + 1.5 * wk_noise),
    Meetings      = pmax(1, round(Meeting_hours * 1.35 + rnorm(n(), 0, 1.6))),
    # Sub-components of meeting hours. Each carries its own person-level share,
    # so they vary independently instead of moving as one block.
    Recurring_meeting_hours       = pmax(0, Meeting_hours * s_recurring * runif(n(), 0.7, 1.3)),
    Large_and_long_meeting_hours  = pmax(0, Meeting_hours * s_large * runif(n(), 0.5, 1.6)),
    Small_and_short_meeting_hours = pmax(0, Meeting_hours * s_small * runif(n(), 0.6, 1.5)),
    Meeting_hours_with_six_or_fewer_hours_of_advanced_notice =
      pmax(0, Meeting_hours * s_notice * runif(n(), 0.4, 1.8)),
    Time_in_self_organized_meetings = pmax(0, Meeting_hours * s_selforg * runif(n(), 0.5, 1.6)),
    # Registered as nulls, so these are driven by their own latents rather than
    # by meeting_latent. Otherwise the planted confounder would leak into every
    # metric in the report and nothing would come out null.
    Conflicting_meeting_hours    = pmax(0, 1.05 + 0.42 * conflict_latent + rnorm(n(), 0, 0.35)),
    Generated_load_Meeting_hours = pmax(0, 6.40 + 1.90 * genload_latent + rnorm(n(), 0, 1.30)),

    # --- Focus time -------------------------------------------------------
    # Uninterrupted hours fall as meeting load rises, which is mechanical rather
    # than behavioural and should be described that way.
    Uninterrupted_hours = pmax(0.5, 16.5 - 0.85 * Meeting_hours + rnorm(n(), 0, 2.4)),
    Interrupted_hours   = pmax(0.5, 9.0 + 0.35 * Meeting_hours + rnorm(n(), 0, 2.0)),
    Available_to_focus_hours = Uninterrupted_hours + Interrupted_hours,
    Open_1_hour_block   = pmax(0, round(9.5 - 0.55 * Meeting_hours + rnorm(n(), 0, 2.2))),

    # --- Ways of working --------------------------------------------------
    # Collaboration hours contain meeting hours by definition, so this one is
    # expected to track meeting load. The rest are registered nulls.
    Collaboration_hours = pmax(2, Meeting_hours + 5.4 + 0.9 * wk_noise),
    After_hours_collaboration_hours = pmax(0, 1.40 + 0.38 * afterhours_latent +
                                             rnorm(n(), 0, 0.45)),
    Collaboration_span  = pmax(6, 9.00 + 0.55 * span_latent + rnorm(n(), 0, 0.80)),
    Internal_network_size = pmax(5, round(38 + 8 * net_int_latent + rnorm(n(), 0, 6))),
    External_network_size = pmax(0, round(9 + 2.6 * net_ext_latent + rnorm(n(), 0, 2)))
  )
```

```{r simulate-github}
# Daily GitHub Copilot activity. Sparse by construction, so that the aggregation
# rule has real work to do.
gh_daily <- expand_grid(
    PersonId = filter(people_all, IsCopilotLicensed)$PersonId,
    MetricDate = days
  ) |>
  filter(!lubridate::wday(MetricDate) %in% c(1, 7)) |>
  left_join(select(people_all, PersonId, ai_latent, chat_pref, agent_pref,
                   p_active_day), by = "PersonId") |>
  # Attach the meeting load and focus of the week the day falls in.
  mutate(week_start = weeks[findInterval(MetricDate, weeks)]) |>
  left_join(select(pq, PersonId, MetricDate, Meeting_hours,
                   Uninterrupted_hours, Open_1_hour_block),
            by = c("PersonId" = "PersonId", "week_start" = "MetricDate")) |>
  filter(!is.na(Meeting_hours)) |>
  mutate(
    mtg_z   = as.numeric(scale(Meeting_hours)),
    unint_z = as.numeric(scale(Uninterrupted_hours)),
    open_z  = as.numeric(scale(Open_1_hour_block)),

    active = rbinom(n(), 1, p_active_day),

    # Total intensity falls weakly with meeting load, which is the substitution
    # reading, and rises with uninterrupted time.
    intensity = exp(2.05 + 0.62 * ai_latent - 0.10 * mtg_z + 0.14 * unint_z),

    # Chat share rises with meeting load and falls when open blocks are
    # plentiful. A fragmented day favours inline completions over a deliberate
    # conversation with the assistant.
    chat_share = plogis(-0.95 + 0.55 * chat_pref + 0.22 * mtg_z - 0.12 * open_z),

    suggested = active * rpois(n(), pmin(intensity, 120)),
    accept_p  = plogis(0.05 + 0.20 * unint_z + 0.25 * ai_latent + rnorm(n(), 0, 0.3)),
    accepted  = rbinom(n(), suggested, accept_p),
    chat_requests = active * rpois(n(), pmin(intensity * chat_share, 60)),
    agent_adoption = active == 1 &
      rbinom(n(), 1, plogis(-1.1 + 0.85 * agent_pref + 0.35 * ai_latent)) == 1
  ) |>
  # Sparsity. An inactive day is simply absent from the export, which is what a
  # real extract looks like.
  filter(active == 1) |>
  transmute(PersonId, MetricDate, week_start,
            `Code completions suggested` = suggested,
            `Code completions accepted`  = accepted,
            `User-initiated chat requests` = chat_requests,
            `Agent adoption` = agent_adoption)
```

```{r simulate-breakdowns}
models    <- c("GPT-4.1", "GPT-5 mini", "Claude Sonnet", "o4-mini", "Default")
languages <- c("TypeScript", "Python", "Java", "Go", "C#", "SQL")

# Model mix is a team-level habit with person-level noise. Some teams lean on
# frontier models and some stay on the default.
team_model_bias <- tibble(
  Team = teams,
  frontier_lean = c(0.9, 0.2, -0.4, 0.6, 1.1, -0.7)
)

model_mix <- devs |>
  left_join(team_model_bias, by = "Team") |>
  rowwise() |>
  mutate(w = list({
    base <- c(0.9 + 0.55 * frontier_lean, 0.8, 0.7 + 0.45 * frontier_lean,
              0.6, 1.4 - 0.60 * frontier_lean)
    base <- pmax(base, 0.05) * exp(rnorm(5, 0, 0.35))
    base / sum(base)
  })) |>
  ungroup() |>
  select(PersonId, Team, w) |>
  unnest_wider(w, names_sep = "_") |>
  setNames(c("PersonId", "Team", models))

language_mix <- devs |>
  rowwise() |>
  mutate(w = list({
    b <- exp(rnorm(6, 0, 0.9)); b / sum(b)
  })) |>
  ungroup() |>
  select(PersonId, w) |>
  unnest_wider(w, names_sep = "_") |>
  setNames(c("PersonId", languages))
```

```{r save-synthetic-data}
DATA_DIR <- file.path("_data", "github")
SOURCE_DIR <- file.path(DATA_DIR, "github-query")
PERSON_DIR <- file.path(DATA_DIR, "person-query")
REFERENCE_DIR <- file.path(DATA_DIR, "reference")
dir.create(SOURCE_DIR, showWarnings = FALSE, recursive = TRUE)
dir.create(PERSON_DIR, showWarnings = FALSE, recursive = TRUE)
dir.create(REFERENCE_DIR, showWarnings = FALSE, recursive = TRUE)

model_mix_long <- model_mix |>
  pivot_longer(all_of(models), names_to = "Model", values_to = "Share")

language_mix_long <- language_mix |>
  pivot_longer(all_of(languages), names_to = "Language", values_to = "Share")

write.csv(
  select(people_all, PersonId, Team, Tenure, JobFamily, IsDeveloper,
         IsCopilotLicensed),
  file.path(REFERENCE_DIR, "people-snapshot.csv"),
  row.names = FALSE
)
write.csv(
  gh_daily,
  file.path(SOURCE_DIR, "activity-daily.csv"),
  row.names = FALSE
)
write.csv(
  model_mix_long,
  file.path(SOURCE_DIR, "model-mix.csv"),
  row.names = FALSE
)
write.csv(
  language_mix_long,
  file.path(SOURCE_DIR, "language-mix.csv"),
  row.names = FALSE
)
pq_export <- pq |>
  select(
    PersonId, MetricDate, Team, Tenure, JobFamily, IsDeveloper,
    IsCopilotLicensed, Meeting_hours, Meetings, Recurring_meeting_hours,
    Conflicting_meeting_hours, Large_and_long_meeting_hours,
    Small_and_short_meeting_hours,
    Meeting_hours_with_six_or_fewer_hours_of_advanced_notice,
    Time_in_self_organized_meetings, Generated_load_Meeting_hours,
    Uninterrupted_hours, Interrupted_hours, Available_to_focus_hours,
    Open_1_hour_block, Collaboration_hours,
    After_hours_collaboration_hours, Collaboration_span,
    Internal_network_size, External_network_size
  )

write.csv(
  pq_export,
  file.path(PERSON_DIR, "person-query-weekly.csv"),
  row.names = FALSE
)
```

```{r weekly-panel}
gh_weekly <- gh_daily |>
  group_by(PersonId, MetricDate = week_start) |>
  summarise(
    observed_days = n(),
    suggested = sum(`Code completions suggested`),
    accepted  = sum(`Code completions accepted`),
    chat      = sum(`User-initiated chat requests`),
    agent_days = sum(`Agent adoption`),
    .groups = "drop"
  )

panel <- pq |>
  filter(IsDeveloper, IsCopilotLicensed) |>
  inner_join(gh_weekly, by = c("PersonId", "MetricDate")) |>
  mutate(
    total_ai      = accepted + chat,
    ai_per_day    = total_ai / observed_days,
    accept_rate   = if_else(suggested > 0, accepted / suggested, NA_real_),
    chat_share_ai = if_else(total_ai > 0, chat / total_ai, NA_real_),
    agent_rate    = agent_days / observed_days
  )

dev_summary <- panel |>
  group_by(PersonId, Team, Tenure) |>
  summarise(
    weeks_observed = n(),
    observed_days  = sum(observed_days),
    total_ai_volume = sum(total_ai),
    ai_per_day     = sum(total_ai) / sum(observed_days),
    accept_rate    = sum(accepted) / pmax(sum(suggested), 1),
    chat_share_ai  = sum(chat) / pmax(sum(accepted) + sum(chat), 1),
    agent_rate     = sum(agent_days) / sum(observed_days),
    across(c(Meeting_hours, Meetings, Recurring_meeting_hours,
             Conflicting_meeting_hours, Large_and_long_meeting_hours,
             Small_and_short_meeting_hours,
             Meeting_hours_with_six_or_fewer_hours_of_advanced_notice,
             Time_in_self_organized_meetings, Generated_load_Meeting_hours,
             Uninterrupted_hours, Interrupted_hours, Available_to_focus_hours,
             Open_1_hour_block, Collaboration_hours,
             After_hours_collaboration_hours, Collaboration_span,
             Internal_network_size, External_network_size), mean),
    .groups = "drop"
  ) |>
  # Funnel step 5. Eight weeks of observation before a developer is analysed.
  filter(weeks_observed >= 8)
```

```{r archetypes}
q <- function(x, p) as.numeric(quantile(x, p, na.rm = TRUE))

dev_summary <- dev_summary |>
  mutate(
    Archetype = case_when(
      ai_per_day    <  q(ai_per_day, 0.25)                      ~ "Light",
      agent_rate    >= q(agent_rate, 0.80)                      ~ "Agent-led",
      chat_share_ai >= q(chat_share_ai, 0.70)                   ~ "Chat-led",
      chat_share_ai <= q(chat_share_ai, 0.30)                   ~ "Completion-led",
      TRUE                                                      ~ "Balanced"
    ),
    Archetype = factor(Archetype,
                       levels = c("Light", "Completion-led", "Balanced",
                                  "Chat-led", "Agent-led"))
  )

arch_tbl <- dev_summary |>
  group_by(Archetype) |>
  summarise(
    People = n(),
    `AI interactions per observed day` = round(mean(ai_per_day), 1),
    `Acceptance rate` = percent(mean(accept_rate), accuracy = 0.1),
    `Chat share of AI activity` = percent(mean(chat_share_ai), accuracy = 0.1),
    `Days with agent use` = percent(mean(agent_rate), accuracy = 0.1),
    .groups = "drop"
  ) |>
  filter(People >= MIN_GROUP_N)

REPORTABLE_ARCH <- arch_tbl$Archetype
dev_rep <- filter(dev_summary, Archetype %in% REPORTABLE_ARCH)

heavy <- dev_rep |> filter(ai_per_day >= q(ai_per_day, 0.75))
light <- dev_rep |> filter(ai_per_day <= q(ai_per_day, 0.25))

MEETING_METRICS <- c(
  "Meeting_hours", "Meetings", "Recurring_meeting_hours",
  "Conflicting_meeting_hours", "Large_and_long_meeting_hours",
  "Small_and_short_meeting_hours",
  "Meeting_hours_with_six_or_fewer_hours_of_advanced_notice",
  "Time_in_self_organized_meetings", "Generated_load_Meeting_hours"
)
FOCUS_METRICS <- c("Uninterrupted_hours", "Interrupted_hours",
                   "Available_to_focus_hours", "Open_1_hour_block")
WOW_METRICS <- c("Collaboration_hours", "After_hours_collaboration_hours",
                 "Collaboration_span", "Internal_network_size",
                 "External_network_size")

METRIC_LABELS <- c(
  Meeting_hours = "Meeting hours",
  Meetings = "Meetings",
  Recurring_meeting_hours = "Recurring meeting hours",
  Conflicting_meeting_hours = "Conflicting meeting hours",
  Large_and_long_meeting_hours = "Large and long meetings",
  Small_and_short_meeting_hours = "Small and short meetings",
  Meeting_hours_with_six_or_fewer_hours_of_advanced_notice = "Meetings with <6h notice",
  Time_in_self_organized_meetings = "Self-organised meetings",
  Generated_load_Meeting_hours = "Generated meeting load",
  Uninterrupted_hours = "Uninterrupted hours",
  Interrupted_hours = "Interrupted hours",
  Available_to_focus_hours = "Available-to-focus hours",
  Open_1_hour_block = "Open one-hour blocks",
  Collaboration_hours = "Collaboration hours",
  After_hours_collaboration_hours = "After-hours collaboration",
  Collaboration_span = "Collaboration span",
  Internal_network_size = "Internal network size",
  External_network_size = "External network size"
)

scan_table <- function(metrics) {
  tibble(Metric = metrics) |>
    rowwise() |>
    mutate(
      `Heaviest quartile` = round(mean(heavy[[Metric]]), 2),
      `Lightest quartile` = round(mean(light[[Metric]]), 2),
      d = round(std_diff(heavy[[Metric]], light[[Metric]]), 2)
    ) |>
    ungroup() |>
    mutate(Label = unname(METRIC_LABELS[Metric]))
}

meeting_scan <- scan_table(MEETING_METRICS)
focus_scan   <- scan_table(FOCUS_METRICS)
wow_scan     <- scan_table(WOW_METRICS)

team_baseline <- dev_rep |>
  group_by(Team) |>
  summarise(
    People = n(),
    `AI interactions / day` = mean(ai_per_day),
    `Acceptance rate` = mean(accept_rate),
    `Chat share of AI activity` = mean(chat_share_ai),
    `Agent-use rate` = mean(agent_rate),
    `Meeting hours` = mean(Meeting_hours),
    `Uninterrupted hours` = mean(Uninterrupted_hours),
    .groups = "drop"
  ) |>
  filter(People >= MIN_GROUP_N)

TEAM_ORDER <- team_baseline |> arrange(`AI interactions / day`) |> pull(Team)

OBS_WINDOW_START <- min(weeks)
OBS_WINDOW_END <- max(weeks) + 6
CHART_CAPTION <- sprintf(
  "Synthetic data | n = %s developers | %s to %s",
  comma(nrow(dev_rep)),
  format(OBS_WINDOW_START, "%d %b %Y"),
  format(OBS_WINDOW_END, "%d %b %Y")
)
```

<style>
body {
  font-family: "Segoe UI Variable", "Segoe UI", system-ui, sans-serif;
  color: #374151;
  background: #F4F4F4;
}
.navbar {
  min-height: 52px;
}
.navbar-brand {
  font-weight: 600;
}
.chart-wrapper {
  border: 1px solid #E5E7EB;
  border-radius: 8px;
  box-shadow: none;
}
.chart-title {
  color: #374151;
  font-weight: 500;
}
.value-box {
  min-height: 108px !important;
  height: 108px !important;
  color: #111827 !important;
  background: #FFFFFF !important;
  border: 1px solid #E5E7EB;
  border-top: 4px solid #3B82F6;
  border-radius: 8px;
  box-shadow: none;
}
.value-box-info { border-top-color: #14B8A6; }
.value-box-success { border-top-color: #8B5CF6; }
.value-box-warning { border-top-color: #F59E0B; }
.value-box > .inner {
  padding: 13px 18px !important;
}
.value-box .value {
  color: #111827 !important;
  font-size: 30px !important;
  line-height: 1.05;
}
.value-box .caption {
  color: #374151 !important;
  font-size: 13px !important;
  line-height: 1.25;
}
.value-box .icon i {
  top: 15px !important;
  right: 16px !important;
  font-size: 52px !important;
  color: rgba(17, 24, 39, 0.09) !important;
}
.insight-banner {
  margin: 8px 8px 0 0;
  padding: 12px 18px;
  border-left: 4px solid #3B82F6;
  border-radius: 8px;
  background: #EFF6FF;
  color: #1E3A5F;
  font-size: 15px;
}
.synthetic-pill {
  display: inline-block;
  margin-right: 8px;
  padding: 2px 8px;
  border: 1px solid #F59E0B;
  border-radius: 999px;
  background: #FFF7E6;
  color: #92400E;
  font-size: 12px;
  font-weight: 600;
}
table {
  font-variant-numeric: tabular-nums;
}
th {
  background: #F9FAFB;
}
.chart-stage img {
  max-width: 100% !important;
  height: auto !important;
}
@media (max-width: 767px) {
  body { padding-left: 0 !important; }
  html, body, #dashboard-container, .tab-content,
  .dashboard-page-wrapper, .dashboard-column-orientation,
  .dashboard-column, .chart-wrapper, .value-box {
    min-width: 0 !important;
    max-width: 100% !important;
    width: 100% !important;
    box-sizing: border-box !important;
  }
  .dashboard-row {
    min-width: 0 !important;
    flex-direction: column !important;
  }
  .chart-wrapper {
    min-width: 0 !important;
    width: calc(100% - 8px) !important;
  }
  .value-box {
    min-height: 92px !important;
    height: 92px !important;
  }
  .value-box .value { font-size: 27px !important; }
  .chart-stage { overflow-x: auto !important; }
  table { font-size: 12px; }
}
</style>

Overview {data-icon="fa-dashboard"}
=====================================

<div class="insight-banner">
<span class="synthetic-pill">Synthetic demonstration</span>
Meeting load and focus time are the two dimensions where AI use moves. Ways of
working is largely flat for this developer population, which is the expected
pattern once the analysis is restricted to engineering roles.
</div>

Row {data-height=120}
-------------------------------------

### Developers represented

```{r}
flexdashboard::valueBox(
  comma(nrow(dev_rep)), "Developers meeting the observation threshold",
  icon = "fa-users", color = "primary"
)
```

### AI interactions per day

```{r}
flexdashboard::valueBox(
  format(round(mean(dev_rep$ai_per_day), 1), nsmall = 1),
  "AI interactions / observed day (average)",
  icon = "fa-bolt", color = "info"
)
```

### Acceptance rate

```{r}
flexdashboard::valueBox(
  percent(mean(dev_rep$accept_rate, na.rm = TRUE), accuracy = 0.1),
  "Accepted share of code completions suggested",
  icon = "fa-check", color = "success"
)
```

### Chat share of AI activity

```{r}
flexdashboard::valueBox(
  percent(mean(dev_rep$chat_share_ai, na.rm = TRUE), accuracy = 0.1),
  "Chat's share of total AI interactions",
  icon = "fa-comments", color = "warning"
)
```

Row {data-height=510}
-------------------------------------

### Team baseline: AI usage and meeting load

```{r team-baseline-chart, fig.height=4.4, fig.alt="Horizontal bars comparing average AI interactions per day and meeting hours per developer across six teams."}
baseline_long <- team_baseline |>
  select(Team, People, `AI interactions / day`, `Meeting hours`) |>
  pivot_longer(c(`AI interactions / day`, `Meeting hours`),
               names_to = "Measure", values_to = "Value") |>
  mutate(Label = format(round(Value, 1), nsmall = 1),
         Team = factor(Team, levels = TEAM_ORDER))

ggplot(baseline_long, aes(Team, Value, fill = Measure)) +
  geom_col(position = position_dodge(width = 0.72), width = 0.66) +
  geom_text(aes(label = Label),
            position = position_dodge(width = 0.72),
            hjust = -0.10, size = 3) +
  coord_flip() +
  facet_wrap(~Measure, scales = "free_x") +
  scale_fill_manual(values = PAL[c(1, 3)], guide = "none") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.18))) +
  theme_t() +
  labs(
    subtitle = "Developers meeting the observation threshold, by team",
    x = NULL, y = NULL, caption = CHART_CAPTION
  )
```

### Baseline table

```{r}
team_baseline |>
  arrange(desc(`AI interactions / day`)) |>
  transmute(
    Team, People,
    `AI interactions / day` = round(`AI interactions / day`, 1),
    `Acceptance rate` = percent(`Acceptance rate`, accuracy = 0.1),
    `Chat share` = percent(`Chat share of AI activity`, accuracy = 0.1),
    `Agent-use rate` = percent(`Agent-use rate`, accuracy = 0.1),
    `Meeting hours` = round(`Meeting hours`, 1),
    `Uninterrupted hours` = round(`Uninterrupted hours`, 1)
  ) |>
  kable(caption = "Team baseline, established before any cross-metric association is examined.")
```

Row {data-height=335}
-------------------------------------

### Population-level highlights

**This population is developers only.** A GitHub Copilot extract also captures
reviewers, product managers and technical writers. Restricting to confirmed
engineering roles is what makes the team baseline and every comparison below
meaningful; the full population funnel is in the appendix.

**Meeting load and focus time are where AI use moves.** Total AI activity falls
as meeting hours rise, and its chat share rises at the same time. Ways of
working, by contrast, is close to flat, which is the expected result for a
population restricted to a single job family.

**A team-level habit is model choice.** Some teams lean heavily on the default
model rather than a frontier option, which is a prompt for a conversation with
that team rather than a performance judgement on any individual.

### About this population and metric

**AI interactions per observed day** sums accepted code completions and chat
requests, then divides by the number of days a developer actually appears in
the extract, rather than by every calendar day in the window. A calendar-day
denominator would understate light and intermittent users, because a missing
person-day is unobserved rather than zero.

See the appendix for the full population funnel, the join rules, and the
effects deliberately placed in the simulation.

Usage distribution {data-icon="fa-chart-area"}
=====================================

Row {data-height=470}
-------------------------------------

### Concentration of AI activity across developers

```{r ai-concentration, fig.height=4.4, fig.alt="Concentration curve showing the cumulative share of AI activity accounted for by the highest-usage share of developers."}
ai_rank <- dev_rep |>
  arrange(desc(total_ai_volume)) |>
  mutate(
    n = n(),
    `Share of developers` = row_number() / n,
    `Share of AI activity` = cumsum(total_ai_volume) / sum(total_ai_volume)
  )

marker_x <- c(0.10, 0.25)
marker_y <- vapply(marker_x, function(p) {
  ai_rank$`Share of AI activity`[which.min(abs(ai_rank$`Share of developers` - p))]
}, numeric(1))
markers <- tibble(x = marker_x, y = marker_y,
                  lbl = percent(marker_y, accuracy = 1))

ggplot(ai_rank, aes(`Share of developers`, `Share of AI activity`)) +
  geom_area(fill = PAL[3], alpha = 0.35) +
  geom_line(colour = PAL[1], linewidth = 1) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "grey55") +
  geom_point(data = markers, aes(x, y), colour = PAL[6], size = 2.6,
            inherit.aes = FALSE) +
  geom_text(data = markers, aes(x, y, label = lbl), inherit.aes = FALSE,
            vjust = -1.0, hjust = -0.15, size = 3.3, colour = PAL[6]) +
  scale_x_continuous(labels = percent_format(accuracy = 1),
                     expand = expansion(mult = c(0.01, 0.06))) +
  scale_y_continuous(labels = percent_format(accuracy = 1),
                     expand = expansion(mult = c(0.01, 0.10))) +
  theme_t() +
  labs(
    title = "Cumulative share of AI activity by developer rank",
    subtitle = paste0(
      "Developers ranked from the highest to the lowest total AI activity; ",
      "the dashed line marks equal usage. A curve further above the line ",
      "indicates more concentrated usage"
    ),
    x = "Share of developers (ranked highest to lowest)",
    y = "Cumulative share of AI activity",
    caption = CHART_CAPTION
  )
```

### Concentration bands

```{r}
band_levels <- c("Top 1%", "Next 4% (top 1-5%)", "Next 5% (top 5-10%)",
                 "Next 15% (top 10-25%)", "Next 25% (top 25-50%)",
                 "Bottom 50%")

band_table <- ai_rank |>
  mutate(
    Band = case_when(
      `Share of developers` <= 0.01 ~ band_levels[1],
      `Share of developers` <= 0.05 ~ band_levels[2],
      `Share of developers` <= 0.10 ~ band_levels[3],
      `Share of developers` <= 0.25 ~ band_levels[4],
      `Share of developers` <= 0.50 ~ band_levels[5],
      TRUE ~ band_levels[6]
    ),
    Band = factor(Band, levels = band_levels)
  ) |>
  group_by(Band) |>
  summarise(
    People = n(),
    `AI activity (band total)` = sum(total_ai_volume),
    .groups = "drop"
  ) |>
  arrange(Band) |>
  mutate(
    `Share of developers` = percent(People / sum(People), accuracy = 1),
    `Share of AI activity` = percent(
      `AI activity (band total)` / sum(`AI activity (band total)`),
      accuracy = 1
    )
  ) |>
  select(Band, People, `Share of developers`, `Share of AI activity`)

band_table |>
  kable(caption = paste0(
    "Developers ranked by total AI activity over the observation window (n = ",
    comma(sum(band_table$People)), ")."
  ))
```

Row {data-height=440}
-------------------------------------

### Distribution of AI interactions per observed day

```{r ai-histogram, fig.height=4.1, fig.alt="Histogram of AI interactions per observed day per developer, with median and mean reference lines."}
med_ai <- median(dev_rep$ai_per_day)
mean_ai <- mean(dev_rep$ai_per_day)
ref_lines <- tibble(stat = c("Median", "Mean"), value = c(med_ai, mean_ai))

ggplot(dev_rep, aes(ai_per_day)) +
  geom_histogram(bins = 40, fill = PAL[2], colour = "white", linewidth = 0.2) +
  geom_vline(data = ref_lines, aes(xintercept = value, colour = stat),
            linetype = "dashed", linewidth = 0.7, show.legend = TRUE) +
  scale_colour_manual(values = c(Median = PAL[1], Mean = PAL[6])) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.10))) +
  theme_t() +
  labs(
    title = "Distribution of AI interactions per observed day",
    subtitle = paste0(
      "One point per developer meeting the observation threshold; a long ",
      "right tail indicates a small group of high-usage developers, and a ",
      "more symmetric shape indicates more uniform usage"
    ),
    x = "AI interactions / observed day", y = "Number of developers",
    colour = NULL, caption = CHART_CAPTION
  )
```

### Why look at the distribution

An average such as "AI interactions per observed day" can be pulled upward by a
small number of heavy users. The concentration curve answers the practical
question directly: what share of AI activity does the top share of developers
account for? The histogram shows the shape that produces that concentration.

This has a practical implication for the team baseline on the Overview page. A
team average can look high or low mainly because a handful of its developers
sit in the top usage band, rather than because the team as a whole uses AI
differently.

Meeting load {data-icon="fa-calendar"}
=====================================

Row {data-height=440}
-------------------------------------

### Meeting-load metrics by AI-usage quartile

```{r meeting-scan-chart, fig.height=4.2, fig.alt="Bar chart of standardised differences in meeting-load metrics between the heaviest and lightest AI-usage quartile."}
ggplot(meeting_scan, aes(reorder(Label, d), d, fill = d > 0)) +
  geom_col(width = 0.68) +
  geom_hline(yintercept = 0, colour = "grey40") +
  scale_fill_manual(values = c(`TRUE` = PAL[2], `FALSE` = PAL[1]), guide = "none") +
  coord_flip() +
  theme_t() +
  labs(
    title = "Meeting-load metrics by AI-usage quartile",
    subtitle = "Standardised difference between the heaviest and lightest AI-usage quartile; positive means the heaviest quartile is higher",
    x = NULL, y = "Standardised difference", caption = CHART_CAPTION
  )
```

### Meeting load by AI usage quartile

```{r}
meeting_scan |>
  select(Metric = Label, `Heaviest quartile`, `Lightest quartile`,
         `Standardised difference` = d) |>
  kable(caption = "Meeting-load metrics by AI-usage quartile (mean hours per person-week unless the metric is a count).")
```

Row {data-height=430}
-------------------------------------

### Substitution or compensation

The negative association between meeting hours and total AI activity admits two
readings, and this design cannot separate them on its own.

**Substitution.** Developers with lighter meeting loads have more coding time,
so they generate more AI interactions. The causal arrow runs from calendar to
tool, and the tool usage is a symptom rather than a cause.

**Compensation.** Developers reach for the assistant when their day is
fragmented, because it lowers the cost of picking work back up after an
interruption. The causal arrow runs from calendar to tool again, but the
predicted sign is the opposite.

The mix evidence, shown alongside, is what tells the two apart, and it points
in different directions for different measures: total volume falls with
meeting load, consistent with substitution, while chat's share of that smaller
volume rises, consistent with compensation.

### AI volume and chat share across meeting-load quartiles

```{r substitution-plot, fig.height=3.6, fig.alt="Two line charts showing AI interactions per day and chat share of AI activity across four meeting-load quartiles."}
mtg_q <- dev_rep |>
  mutate(`Meeting load quartile` = factor(ntile(Meeting_hours, 4),
                                          levels = 1:4,
                                          labels = c("Q1 lightest", "Q2", "Q3",
                                                     "Q4 heaviest"))) |>
  group_by(`Meeting load quartile`) |>
  summarise(
    People = n(),
    `AI interactions per day` = round(mean(ai_per_day), 1),
    `Chat share of AI activity` = mean(chat_share_ai),
    `Acceptance rate` = mean(accept_rate),
    .groups = "drop"
  ) |>
  filter(People >= MIN_GROUP_N)

mtg_q |>
  select(`Meeting load quartile`, `AI interactions per day`,
         `Chat share of AI activity`) |>
  mutate(`Chat share of AI activity` = `Chat share of AI activity` * 100) |>
  pivot_longer(-`Meeting load quartile`) |>
  ggplot(aes(`Meeting load quartile`, value, group = name, colour = name)) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2.4) +
  facet_wrap(~name, scales = "free_y") +
  scale_colour_manual(values = PAL[c(1, 2)], guide = "none") +
  theme_t() +
  labs(
    title = "AI volume and chat share across meeting-load quartiles",
    subtitle = "Two views of the same quartiles: total AI interactions per day, and chat's share of that activity",
    x = NULL, y = NULL, caption = CHART_CAPTION
  )
```

Row {data-height=330}
-------------------------------------

### Meeting load by agent use

```{r}
agent_q <- dev_rep |>
  mutate(`Agent use quartile` = factor(ntile(agent_rate, 4), levels = 1:4,
                                       labels = c("Q1 lowest", "Q2", "Q3",
                                                  "Q4 highest"))) |>
  group_by(`Agent use quartile`) |>
  summarise(People = n(),
            `Meeting hours` = round(mean(Meeting_hours), 2),
            `Recurring meeting hours` = round(mean(Recurring_meeting_hours), 2),
            `Generated load meeting hours` = round(mean(Generated_load_Meeting_hours), 2),
            .groups = "drop") |>
  filter(People >= MIN_GROUP_N)

kable(agent_q, caption = "Meeting-load metrics by agent-use quartile (a genuine null in this simulation).")
```

### Reading the null

Agent adoption carries no meeting-load signal here. That is a genuine null and
it belongs in the report alongside the metrics that moved. The most common
failure in an adoption report is to show only the dimensions that moved, which
turns a scan across many metrics into a search for the ones that happened to
look interesting.

Focus time {data-icon="fa-clock-o"}
=====================================

Row {data-height=420}
-------------------------------------

### Focus-time metrics by AI-usage quartile

```{r}
focus_scan |>
  select(Metric = Label, `Heaviest quartile`, `Lightest quartile`,
         `Standardised difference` = d) |>
  kable(caption = "Focus-time metrics by AI-usage quartile.")
```

### Mechanical against behavioural

Uninterrupted hours fall as meeting hours rise more or less by arithmetic,
because both are drawn from the same working day. That relationship is not a
finding about behaviour. The two charts alongside look instead at whether
contiguous time changes how a developer uses the assistant, once meeting load
itself is set aside.

`Available-to-focus hours` has shown unexplained step changes in production
tenancies, appearing simultaneously in populations that had no exposure to
whatever intervention was being studied. Read it alongside `Uninterrupted
hours` rather than on its own, and check for a level shift across the whole
population before attributing any movement in it to a programme.

Row {data-height=420}
-------------------------------------

### Acceptance rate by uninterrupted-hours quartile

```{r focus-completions, fig.height=3.9, fig.alt="Bar chart of completion acceptance rate across four uninterrupted-hours quartiles."}
dev_rep |>
  mutate(`Uninterrupted hours quartile` =
           factor(ntile(Uninterrupted_hours, 4), levels = 1:4,
                  labels = c("Q1 least", "Q2", "Q3", "Q4 most"))) |>
  group_by(`Uninterrupted hours quartile`) |>
  summarise(People = n(),
            value = mean(accept_rate), .groups = "drop") |>
  filter(People >= MIN_GROUP_N) |>
  ggplot(aes(`Uninterrupted hours quartile`, value)) +
  geom_col(fill = PAL[1], width = 0.65) +
  geom_text(aes(label = percent(value, accuracy = 0.1)), vjust = -0.5, size = 3.2) +
  scale_y_continuous(labels = percent_format(accuracy = 1),
                     expand = expansion(mult = c(0, 0.14))) +
  theme_t() +
  labs(
    title = "Acceptance rate by uninterrupted-hours quartile",
    subtitle = "Association only; contiguous time and receptiveness to inline suggestions may share a cause",
    x = NULL, y = "Accepted share of suggestions", caption = CHART_CAPTION
  )
```

### Chat share by open-block quartile

```{r fragmentation, fig.height=3.9, fig.alt="Bar chart of chat share of AI activity across four open-one-hour-block quartiles."}
dev_rep |>
  mutate(`Open one-hour blocks` =
           factor(ntile(Open_1_hour_block, 4), levels = 1:4,
                  labels = c("Q1 fewest", "Q2", "Q3", "Q4 most"))) |>
  group_by(`Open one-hour blocks`) |>
  summarise(People = n(), value = mean(chat_share_ai), .groups = "drop") |>
  filter(People >= MIN_GROUP_N) |>
  ggplot(aes(`Open one-hour blocks`, value)) +
  geom_col(fill = PAL[2], width = 0.65) +
  geom_text(aes(label = percent(value, accuracy = 0.1)), vjust = -0.5, size = 3.2) +
  scale_y_continuous(labels = percent_format(accuracy = 1),
                     expand = expansion(mult = c(0, 0.14))) +
  theme_t() +
  labs(
    title = "Chat share by open-block quartile",
    subtitle = "A fragmented day appears to favour inline completions over deliberate conversation",
    x = NULL, y = "Chat share of AI activity", caption = CHART_CAPTION
  )
```

Team and model view {data-icon="fa-sitemap"}
=====================================

Row {data-height=470}
-------------------------------------

### Usage archetype mix by team

```{r arch-by-team, fig.height=4.4, fig.alt="Stacked bar chart of usage archetype mix across six teams."}
team_arch <- dev_rep |>
  count(Team, Archetype) |>
  group_by(Team) |>
  mutate(total = sum(n), share = n / total) |>
  ungroup() |>
  filter(total >= MIN_GROUP_N)

ggplot(team_arch, aes(Team, share, fill = Archetype)) +
  geom_col(width = 0.7) +
  scale_fill_manual(values = PAL5) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  coord_flip() +
  theme_t() +
  labs(
    title = "Usage archetype mix by team",
    subtitle = "Teams below the reporting threshold are suppressed",
    x = NULL, y = "Share of developers", caption = CHART_CAPTION
  )
```

### AI usage against meeting load, by team

```{r team-bubble, fig.height=4.4, fig.alt="Bubble chart of meeting hours against AI interactions per day by team, with bubble size representing the number of developers."}
ggplot(
  team_baseline,
  aes(`Meeting hours`, `AI interactions / day`,
      size = People, colour = Team, label = Team)
) +
  geom_point(alpha = 0.85) +
  geom_text_repel(size = 3.2, colour = "grey20", seed = 2026,
                  min.segment.length = 0, max.overlaps = Inf,
                  box.padding = 0.6, point.padding = 0.3,
                  show.legend = FALSE) +
  scale_size_area(max_size = 14) +
  scale_colour_manual(values = TEAM_PAL, guide = "none") +
  scale_x_continuous(expand = expansion(mult = 0.18)) +
  scale_y_continuous(expand = expansion(mult = 0.18)) +
  theme_t() +
  labs(
    title = "AI usage against meeting load, by team",
    subtitle = "Bubble size shows the number of developers represented in each team",
    x = "Meeting hours / developer / week",
    y = "AI interactions / observed day",
    size = "Developers", caption = CHART_CAPTION
  )
```

Row {data-height=460}
-------------------------------------

### Model mix by team

```{r model-mix, fig.height=4.2, fig.alt="Stacked bar chart of model mix across six teams."}
model_long <- model_mix |>
  filter(PersonId %in% dev_rep$PersonId) |>
  pivot_longer(all_of(models), names_to = "Model", values_to = "share") |>
  group_by(Team, Model) |>
  summarise(People = n_distinct(PersonId), share = mean(share), .groups = "drop") |>
  filter(People >= MIN_GROUP_N)

ggplot(model_long, aes(Team, share, fill = Model)) +
  geom_col(width = 0.7) +
  scale_fill_manual(values = PAL5) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  coord_flip() +
  theme_t() +
  labs(
    title = "Model mix by team",
    subtitle = "Reliance on the default model varies widely between teams",
    x = NULL, y = "Share of model-attributed activity", caption = CHART_CAPTION
  )
```

### Reliance on the default model

```{r default-reliance}
model_long |>
  filter(Model == "Default") |>
  arrange(desc(share)) |>
  transmute(Team, `Share on the default model` = percent(share, accuracy = 0.1)) |>
  kable(caption = "Reliance on the default model (a high share is a prompt for a conversation rather than a problem in itself).")
```

A team sitting almost entirely on the default model may be perfectly well
served by it. It may equally be a team that has never been shown how to pick a
model for a task. The measure identifies where to ask rather than what to
conclude.

Row {data-height=380}
-------------------------------------

### Language mix across the developer population

```{r language-mix, fig.height=3.5, fig.alt="Horizontal bar chart of programming language mix across the developer population."}
language_mix |>
  filter(PersonId %in% dev_rep$PersonId) |>
  pivot_longer(all_of(languages), names_to = "Language", values_to = "share") |>
  group_by(Language) |>
  summarise(share = mean(share), .groups = "drop") |>
  ggplot(aes(reorder(Language, share), share)) +
  geom_col(fill = PAL[2], width = 0.65) +
  geom_text(aes(label = percent(share, accuracy = 0.1)), hjust = -0.12, size = 3.1) +
  scale_y_continuous(labels = percent_format(accuracy = 1),
                     expand = expansion(mult = c(0, 0.16))) +
  coord_flip() +
  theme_t() +
  labs(
    title = "Language mix across the developer population",
    subtitle = "Useful for targeting enablement rather than for assessing anyone",
    x = NULL, y = "Share of language-attributed activity", caption = CHART_CAPTION
  )
```

### Reading team and model views

Model choice is a team habit more than an individual one, and it is the
closest thing in this data to a measure of how deliberately a team is using
the tooling. Language mix is a targeting tool for enablement content rather
than an assessment of any individual or team.

Ways of working and screening {data-icon="fa-balance-scale"}
=====================================

Row {data-height=440}
-------------------------------------

### Ways-of-working metrics by AI-usage quartile

```{r wow-scan-chart, fig.height=4.2, fig.alt="Bar chart of standardised differences in ways-of-working metrics between the heaviest and lightest AI-usage quartile."}
ggplot(wow_scan, aes(reorder(Label, d), d, fill = d > 0)) +
  geom_col(width = 0.6) +
  geom_hline(yintercept = 0, colour = "grey40") +
  scale_fill_manual(values = c(`TRUE` = PAL[2], `FALSE` = PAL[1]), guide = "none") +
  coord_flip() +
  theme_t() +
  labs(
    title = "Ways-of-working metrics by AI-usage quartile",
    subtitle = "Mostly flat, which is the expected result for a developer population",
    x = NULL, y = "Standardised difference", caption = CHART_CAPTION
  )
```

### Ways of working by AI usage quartile

```{r}
wow_scan |>
  select(Metric = Label, `Heaviest quartile`, `Lightest quartile`,
         `Standardised difference` = d) |>
  kable(caption = "Ways-of-working metrics by AI-usage quartile.")
```

Most of this dimension is flat. Meeting load and focus time are where a coding
assistant would plausibly show up, and reporting the flat dimensions alongside
the moving ones is what stops the report reading as advocacy. These metrics
are shared with the companion consumption template, so a developer population
can be compared with the wider licensed population on the same terms.

Row {data-height=430}
-------------------------------------

### Which working-pattern metrics separate heavier from lighter AI users

```{r iv-analysis, fig.height=4.0, fig.alt="Bar chart ranking working-pattern metrics by Information Value in separating heavier from lighter AI users."}
iv_data <- dev_rep |>
  mutate(HeavyUser = as.integer(ai_per_day >= median(ai_per_day))) |>
  select(HeavyUser, all_of(c(MEETING_METRICS, FOCUS_METRICS, WOW_METRICS))) |>
  as.data.frame()

iv_res <- create_IV(iv_data, outcome = "HeavyUser", return = "summary")

as_tibble(iv_res) |>
  arrange(desc(IV)) |>
  head(12) |>
  mutate(Variable = coalesce(unname(METRIC_LABELS[Variable]),
                             gsub("_", " ", Variable))) |>
  ggplot(aes(reorder(Variable, IV), IV)) +
  geom_col(fill = PAL[1], width = 0.68) +
  coord_flip() +
  theme_t() +
  labs(
    title = "Which working-pattern metrics separate heavier from lighter AI users",
    subtitle = "Information Value is a screening rank rather than an effect size",
    x = NULL, y = "Information Value", caption = CHART_CAPTION
  )
```

### Reading the ranking

Read this as a shortlist rather than a result. In a realistic developer
population the values are small, and every one of them sits below the
conventional 0.1 threshold at which Information Value is treated as weakly
predictive. That is itself the finding: working patterns separate heavy from
light AI users only faintly, so any claim that AI use can be read off a
calendar, or a calendar off AI use, is not supported.

The metrics at the top are the ones worth taking into a properly specified
model or a conversation with the teams. The metrics at the bottom are the ones
to stop reporting.

Rigour checks {data-icon="fa-flask"}
=====================================

Row {data-height=380}
-------------------------------------

### Tenure mix by archetype

```{r composition}
dev_rep |>
  count(Archetype, Tenure) |>
  group_by(Archetype) |>
  mutate(share = n / sum(n)) |>
  ungroup() |>
  select(-n) |>
  pivot_wider(names_from = Tenure, values_from = share, values_fill = 0) |>
  mutate(across(where(is.numeric), \(x) percent(x, accuracy = 0.1))) |>
  kable(caption = "Tenure mix by archetype (the archetypes are not drawn from the same population).")
```

### Why adjustment can widen a gap

Tenure drives both meeting load and AI appetite in this simulation, so part of
the raw association between them is a fact about who the heavier users are
rather than about what they do. The adjusted gaps alongside are **larger**
than the unadjusted ones, which surprises people who expect adjustment only
ever to shrink things. It happens when a confounder pushes in the opposite
direction to the effect: more senior developers use AI more and also sit in
more meetings, so seniority masks the underlying negative relationship
between meeting load and AI use. Removing it lets the relationship show.

Adjusting for tenure and team does not make the comparison causal. It removes
two known sources of confounding and leaves every unknown one in place.

Row {data-height=380}
-------------------------------------

### Heavier against lighter AI users, before and after adjustment

```{r adjusted}
adj <- dev_rep |>
  mutate(Heavy = as.integer(ai_per_day >= median(ai_per_day)))

adj_compare <- function(metric) {
  m1 <- lm(reformulate("Heavy", metric), data = adj)
  m2 <- lm(reformulate(c("Heavy", "Tenure", "Team"), metric), data = adj)
  tibble(
    Metric = gsub("_", " ", metric),
    Model  = c("Unadjusted", "Adjusted for tenure and team"),
    Gap    = round(c(coef(m1)[["Heavy"]], coef(m2)[["Heavy"]]), 2),
    `Lower 95%` = round(c(confint(m1)["Heavy", 1], confint(m2)["Heavy", 1]), 2),
    `Upper 95%` = round(c(confint(m1)["Heavy", 2], confint(m2)["Heavy", 2]), 2)
  )
}

bind_rows(adj_compare("Meeting_hours"),
          adj_compare("Uninterrupted_hours")) |>
  kable(caption = "Heavier against lighter AI users, before and after accounting for who is in each group.")
```

### A stronger design: within-person change

Comparing archetypes is the weakest available design, because developers
chose their own archetype. A within-person comparison is stronger, since each
developer acts as their own control and every stable attribute, including
tenure, team and seniority, cancels out. What it cannot remove is anything
that changed at the same time as AI use, and in a real deployment a great deal
changes at once.

Row {data-height=440}
-------------------------------------

### Within-person change: meeting hours against AI activity

```{r within-person, fig.height=4.0, fig.alt="Scatter plot of within-person change in meeting hours against within-person change in AI interactions per day."}
halves <- panel |>
  mutate(half = if_else(MetricDate < median(weeks), "First half", "Second half")) |>
  group_by(PersonId, half) |>
  summarise(ai = sum(total_ai) / sum(observed_days),
            mtg = mean(Meeting_hours),
            unint = mean(Uninterrupted_hours), .groups = "drop") |>
  pivot_wider(names_from = half, values_from = c(ai, mtg, unint)) |>
  filter(!is.na(`ai_First half`), !is.na(`ai_Second half`)) |>
  mutate(
    d_ai  = `ai_Second half` - `ai_First half`,
    d_mtg = `mtg_Second half` - `mtg_First half`,
    d_unint = `unint_Second half` - `unint_First half`
  )

ggplot(halves, aes(d_mtg, d_ai)) +
  geom_point(alpha = 0.28, colour = PAL[2], size = 1.5) +
  geom_smooth(method = "lm", se = TRUE, colour = PAL[1], linewidth = 0.8) +
  theme_t() +
  labs(
    title = "Within-person change: meeting hours against AI activity",
    subtitle = "Each point is one developer; stable attributes cancel out by construction",
    x = "Change in weekly meeting hours", y = "Change in AI interactions per observed day",
    caption = CHART_CAPTION
  )
```

### Within-person change model

```{r within-person-model}
wp <- lm(d_ai ~ d_mtg + d_unint, data = halves)

tibble(
  Term = c("Change in meeting hours", "Change in uninterrupted hours"),
  Coefficient = round(coef(wp)[c("d_mtg", "d_unint")], 3),
  `Lower 95%` = round(confint(wp)[c("d_mtg", "d_unint"), 1], 3),
  `Upper 95%` = round(confint(wp)[c("d_mtg", "d_unint"), 2], 3)
) |>
  kable(caption = "Within-person change model. Still an association, but a much better controlled one.")
```

Methods and appendix {data-icon="fa-book"}
=====================================

Row {data-height=360}
-------------------------------------

### Purpose and safe use

This is a **fully synthetic template**. It builds tables shaped like a GitHub
Copilot activity export and a weekly Viva Insights Person Query, then runs the
analysis an engineering organisation would run against the real thing. It
reads no customer data and contains none. Every relationship in the charts was
placed there by the simulation, so nothing here is a finding about any
organisation.

The question the template answers is this. **Does the way developers spend
their week shape how they use AI, and does the way they use AI show up in how
they spend their week?** It looks at three dimensions: meeting load, focus
time, and ways of working (collaboration volume, mode mix and workload,
carried over from the companion consumption template).

Nothing here supports a causal claim, and the template carries **no delivery
outcomes**: no pull requests merged, no lead time, no deployment frequency and
no change failure rate. Those live in engineering systems rather than Viva
Insights, and pairing them with individual behavioural metrics invites exactly
the individual-performance reading these datasets should not be used for.

### Population scope and what this template does not claim

This report covers **developers only**. The companion template,
`copilot-consumption-ways-of-working-simulation.Rmd`, covers the whole
Copilot-licensed population using a Microsoft Copilot consumption extract. The
two should not be merged: a GitHub Copilot extract describes people writing
code, and a consumption extract describes everybody with a licence. Averaging
a developer population together with a sales population produces a number
that describes neither.

Light meeting loads may leave room for AI use, or AI use may free up time that
then gets spent elsewhere. The template names that ambiguity rather than
resolving it, and the rigour-checks page shows what a stronger design would
need.

Row {data-height=380}
-------------------------------------

### The population funnel

```{r funnel}
funnel <- tribble(
  ~Step, ~Rule, ~`Why it matters`,
  "1. All employees",
  "Everyone in the Person Query.",
  "The widest frame. Includes people who will never touch a repository.",

  "2. Copilot licensed",
  "IsCopilotLicensed is TRUE in the people metadata.",
  "A null in a Copilot metric means unlicensed. A zero means licensed and inactive. Collapsing the two understates adoption.",

  "3. Present in the GitHub extract",
  "PersonId appears at least once in the GitHub activity export.",
  "Necessary but not sufficient. A product manager reviewing a pull request can appear here.",

  "4. Developer confirmed",
  "Job family or function in the Person Query identifies an engineering role.",
  "This is the analytical population. It is defined from HR attributes rather than from tool presence.",

  "5. Sufficient observation",
  "At least eight weeks with any observed activity in the window.",
  "Protects against reading a joiner or a leaver as a light user."
)

kable(funnel, caption = "The developer population funnel. Steps 3 and 4 are different tests and both are needed.")
```

### Funnel applied to the synthetic population

```{r funnel-counts}
gh_people <- unique(gh_weekly$PersonId)

tibble(
  Step = c("1. All employees",
           "2. Copilot licensed",
           "3. Present in the GitHub extract",
           "4. Developer confirmed by job family",
           "5. Observed in at least eight weeks"),
  People = c(
    nrow(people_all),
    sum(people_all$IsCopilotLicensed),
    n_distinct(gh_people),
    n_distinct(intersect(gh_people, devs$PersonId)),
    nrow(dev_summary)
  )
) |>
  mutate(`Share of all employees` = percent(People / nrow(people_all),
                                            accuracy = 0.1)) |>
  kable(caption = "The funnel applied to the synthetic population. Steps 3 and 4 remove different people.")
```

Row {data-height=350}
-------------------------------------

### Who else appears in a GitHub extract

```{r funnel-leakage}
people_all |>
  filter(PersonId %in% gh_people) |>
  count(JobFamily, name = "People") |>
  arrange(desc(People)) |>
  mutate(`Share of the GitHub extract` = percent(People / length(gh_people),
                                                 accuracy = 0.1)) |>
  kable(caption = "Job family of everyone appearing in the GitHub extract. Tool presence is not a developer test.")
```

A meaningful share of the extract is not developers at all. Product managers,
designers and technical writers appear because they review code and open
issues, and their meeting load and focus time look nothing like an engineer's.
The reverse leak matters just as much: developers who hold a licence but never
engage drop out at step 3, and if the population is defined from the extract
they vanish from the denominator entirely. They are the most interesting group
in an adoption report, because they are the ones the enablement effort has not
reached.

### Observed days per developer

```{r sparsity-check, fig.height=3.4, fig.alt="Histogram of observed days per developer in the GitHub extract, with a dashed line marking the weekdays in the window."}
lic_devs <- filter(people_all, IsDeveloper, IsCopilotLicensed)

obs <- gh_daily |>
  count(PersonId, name = "observed_days") |>
  right_join(select(lic_devs, PersonId), by = "PersonId") |>
  mutate(observed_days = coalesce(observed_days, 0L))

weekdays_in_window <- sum(!lubridate::wday(days) %in% c(1, 7))

ggplot(obs, aes(observed_days)) +
  geom_histogram(bins = 30, fill = PAL[2], colour = "white") +
  geom_vline(xintercept = weekdays_in_window, linetype = "dashed",
             colour = PAL[1]) +
  theme_t() +
  labs(
    title = "Observed days per developer in the GitHub extract",
    subtitle = paste0("The dashed line marks the ", weekdays_in_window,
                      " weekdays in the window; almost nobody reaches it"),
    x = "Days with any observed activity", y = "Developers", caption = CHART_CAPTION
  )
```

The gap between the histogram and the dashed line is the whole reason the
sparse panel rule matters. Treating every unobserved day as a zero would
divide most developers' activity by a denominator two or three times larger
than the number of days actually observed. This template carries an
observed-days count through every aggregation and uses per-observed-day rates
wherever a rate is reported.

Row {data-height=380}
-------------------------------------

### Workflow: how the three sources combine

```{r workflow, fig.height=3.1}
nodes <- tribble(
  ~x, ~y, ~label, ~fill,
  1, 3, "GitHub Copilot\nactivity\nPersonId x day", PAL[1],
  1, 2, "Feature, model and\nlanguage breakdowns\nPersonId x day x dim", PAL[2],
  1, 1, "Person Query\nPersonId x week", PAL[3],
  2.5, 2, "Weekly\ndeveloper panel", PAL[2],
  4, 2.6, "Usage archetype\nassignment", PAL[1],
  4, 1.4, "Meeting load and\nfocus time analysis", PAL[1]
)

edges <- tribble(
  ~x, ~y, ~xend, ~yend,
  1.45, 3,   2.05, 2.15,
  1.45, 2,   2.05, 2,
  1.45, 1,   2.05, 1.85,
  2.95, 2.1, 3.55, 2.5,
  2.95, 1.9, 3.55, 1.5
)

ggplot() +
  geom_segment(data = edges, aes(x, y, xend = xend, yend = yend),
               colour = "grey55", linewidth = 0.4,
               arrow = arrow(length = unit(0.14, "cm"), type = "closed")) +
  geom_label(data = nodes, aes(x, y, label = label, fill = fill),
             colour = "white", size = 2.7, label.padding = unit(0.34, "lines"),
             label.r = unit(0.18, "lines"), lineheight = 0.95) +
  scale_fill_identity() +
  scale_x_continuous(limits = c(0.6, 4.6)) +
  scale_y_continuous(limits = c(0.6, 3.4)) +
  theme_void() +
  labs(title = "How the three sources combine")
```

### Join rules

```{r join-rules}
join_rules <- tribble(
  ~Source, ~Grain, ~`Join key`, ~`Rule that has to be stated`,

  "GitHub Copilot activity",
  "PersonId x calendar day",
  "PersonId, date rolled up to week",
  "The daily panel is sparse. A missing person-day is unobserved rather than zero. Roll up to the week by summing observed days, and record how many days were observed.",

  "Feature, model, language breakdowns",
  "PersonId x day x dimension",
  "PersonId, week",
  "These are long tables and do not sum to the person-level totals once a dimension is filtered. Use them for mix rather than for volume.",

  "Person Query",
  "PersonId x week, week starting Sunday",
  "PersonId, week",
  "Align the GitHub week to the Person Query week start before joining, rather than to the ISO week.",

  "People metadata",
  "PeopleHistoricalId",
  "Requires a crosswalk to PersonId",
  "Licensing status lives here and does not join on PersonId directly. Without the crosswalk, unlicensed and inactive cannot be separated."
)

kable(join_rules, caption = "Join rules. The week alignment and the sparse panel rule are the two that most often go wrong.")
```

Row {data-height=380}
-------------------------------------

### Effects placed in the simulation

```{r effect-register}
effect_register <- tribble(
  ~Dimension,      ~Relationship,                                          ~Direction,   ~Strength,
  "Meeting load",  "Meeting hours to total AI activity",                   "Negative",   "Weak",
  "Meeting load",  "Meeting hours to chat share of AI activity",           "Positive",   "Modest",
  "Meeting load",  "Conflicting meeting hours to AI activity",             "None",       "Null",
  "Meeting load",  "Agent adoption to meeting load",                       "None",       "Null",
  "Meeting load",  "Generated load to AI activity",                        "None",       "Null",
  "Focus time",    "Uninterrupted hours to code completions accepted",     "Positive",   "Modest",
  "Focus time",    "Open one-hour blocks to chat share",                   "Negative",   "Weak",
  "Focus time",    "Interrupted hours to acceptance rate",                 "None",       "Null",
  "Ways of working", "AI activity to collaboration hours",                 "Mirrors meeting load", "By construction",
  "Ways of working", "AI activity to after-hours collaboration",           "None",       "Null",
  "Ways of working", "AI activity to collaboration span",                  "None",       "Null",
  "Ways of working", "AI activity to internal and external network size",  "None",       "Null",
  "Confounding",   "Tenure band drives both meeting load and AI appetite", "Both",       "By design"
)

kable(effect_register, caption = "Effects injected by the simulation. Everything else is noise.")
```

Publishing the effect register up front is what makes a synthetic template
useful: a reader can check that the analysis recovers what was put into the
data, and equally that it reports the nulls as nulls. Effects are deliberately
modest and noisy, and several dimensions carry nothing at all. Tenure is the
planted confounder: more senior developers carry more meetings and also adopt
AI more readily, which works against the weak negative effect of meeting load
on AI use and is why the rigour-checks page has to recover it explicitly.

### Synthetic source inventory

```{r inventory}
inv <- tribble(
  ~Table, ~Grain, ~Rows, ~People, ~`Stands in for`,
  "gh_daily", "PersonId x day", nrow(gh_daily), n_distinct(gh_daily$PersonId),
  "PersonGitHubActivityMetrics.csv",
  "model_mix", "PersonId x model", nrow(model_mix) * length(models),
  n_distinct(model_mix$PersonId),
  "GitHubActivityBreakdownByModelFeatureMetrics.csv",
  "language_mix", "PersonId x language", nrow(language_mix) * length(languages),
  n_distinct(language_mix$PersonId),
  "GitHubActivityBreakdownByLanguageFeatureMetrics.csv",
  "pq", "PersonId x week", nrow(pq), n_distinct(pq$PersonId),
  "Weekly Person Query"
) |>
  mutate(Rows = comma(Rows), People = comma(People))

kable(inv, caption = "Synthetic tables and the real exports they stand in for.")
```

The generated source tables are saved as ordinary CSV files under
`examples/utility-r/_data/github` whenever the report is rendered. The
`github-query`, `person-query` and `reference` subfolders expose the source
and join structure directly, without requiring anyone to extract an archive.

Row {data-height=400}
-------------------------------------

### Data gaps to carry into the real version

```{r gaps}
gaps <- tribble(
  ~Gap, ~Consequence, ~`What to request`,

  "No crosswalk between PeopleHistoricalId and PersonId",
  "Licensing status cannot be joined, so an unlicensed developer and an inactive one look identical.",
  "The identifier mapping, or licensing status carried on the activity export itself.",

  "No token or cost field anywhere in the GitHub export",
  "Cost per team, cost per accepted completion, and any efficiency measure are impossible.",
  "The billing extract, joined at team level rather than person level.",

  "Absent person-days are ambiguous",
  "Cannot distinguish inactive from suppressed from ineligible, which biases every per-day rate.",
  "Written confirmation of the export rule, including any minimum-activity suppression.",

  "Short observation window",
  "Usage segments and trailing-window network metrics need more history than a one-month extract carries.",
  "A rolling window of at least six months.",

  "Job family is the only developer marker",
  "Developers in non-engineering job families are missed, and non-developers in engineering ones are included.",
  "A repository-contributor list, or a team roster confirmed by engineering leadership.",

  "No seniority within the engineering track",
  "A staff engineer and a graduate are compared as equals on meeting load and focus time.",
  "The engineering level attribute in the Person Query."
)

kable(gaps, caption = "Known gaps. Each one is a request rather than a caveat.")
```

### Replacing the simulation with real data

```{r replacement}
tribble(
  ~Chunk, ~`Replace with`,
  "simulate-people", "The developer roster from the Person Query, filtered by job family or function.",
  "simulate-person-query", "import_query() on the weekly Person Query export.",
  "simulate-github", "The GitHub Copilot activity export, read as a sparse daily panel.",
  "simulate-breakdowns", "The model, feature and language breakdown exports.",
  "weekly-panel", "Keep as written. The aggregation and join logic transfers unchanged."
) |>
  kable(caption = "What to swap out. Everything downstream of the weekly panel runs unchanged.")
```

Four things to preserve when swapping in real data: keep the funnel (define
the developer population from HR attributes, and report how much of it
appears in the tool); keep the observed-days denominator (every rate in this
template is per observed day); keep the reporting threshold (`MIN_GROUP_N`
counts distinct people rather than rows); and keep the nulls (a report that
shows only the metrics that moved is a search rather than an analysis).

Row {data-height=260}
-------------------------------------

### On individual use

These metrics describe working patterns rather than performance. Meeting
load, focus time and assistant usage vary with role, project phase and the
week of the quarter, and none of them ranks one developer against another.
Report them at team level and above, and route any individual-level use
through the same governance that covers other people data.
