---
title: "Developer Experience and Copilot"
output:
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: scroll
    theme: lumen
    self_contained: true
    mathjax: null
    fig_mobile: false
    resize_reload: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE)
source('github-developer-experience-helpers.R', local=knitr::knit_global())
```

<style>
:root { --primary:#4338ca; --bg:#f4f4f4; --border:#e5e7eb; --muted:#475569; }
/* Blue carries comparisons; teal and purple distinguish composition categories. */
body { font-family:'Segoe UI Variable','Segoe UI',system-ui,sans-serif; color:#374151; background:var(--bg); font-size:14px; }
.navbar { background:#fff; border-bottom:1px solid var(--border); }
.navbar-brand { color:#111827 !important; font-weight:600; font-size:17px; }
.navbar-nav>li>a { color:var(--muted) !important; font-size:14px; padding-left:12px; padding-right:12px; }
.navbar-nav>.active>a { background:#eef2ff !important; color:#3730a3 !important; }
.dropdown-menu>li>a { padding:10px 18px; }
.dashboard-page-wrapper { max-width:1320px; margin:0 auto; padding:8px 16px; }
.chart-wrapper { border:1px solid var(--border); border-radius:12px; box-shadow:none; margin-bottom:20px; background:white; height:auto !important; min-height:0 !important; }
.chart-title { font-size:20px; font-weight:600; padding:20px 24px 0; color:#111827; border:0; line-height:1.3; }
.dashboard-row { height:auto !important; }
.dashboard-page, .dashboard-row, .dashboard-column, .chart-wrapper, .chart-stage, .chart-shim { min-width:0 !important; max-width:100%; }
.chart-stage { height:auto !important; flex:0 0 auto !important; padding:16px 24px 24px !important; overflow:visible !important; }
.chart-shim { position:relative !important; inset:auto !important; }
.image-container { position:static !important; background:none !important; }
.image-container img { opacity:1 !important; }
.chart-stage p { line-height:1.6; }
.chart-stage img { display:block; width:100%; height:auto; max-width:100%; }
.report-chart { margin:0; min-width:0; }
.report-chart h4 { font-size:16px; font-weight:600; color:#111827; margin:0 0 12px; min-height:2.6em; }
.report-chart picture { display:block; }
.chart-grid { display:grid; grid-template-columns:repeat(3,minmax(0,1fr)); gap:24px; margin:20px 0; }
.chart-description,.chart-note { color:var(--muted); font-size:13px; line-height:1.5; }
.chart-description { margin:12px 0; }
.chart-note { margin:12px 0 0; }
.kpis { display:grid; grid-template-columns:repeat(4,minmax(0,1fr)); gap:16px; margin:24px 0; }
.kpi { border:1px solid var(--border); border-top:3px solid #3b82f6; border-radius:8px; padding:16px; background:white; }
.kpi span,.kpi small { display:block; color:var(--muted); font-size:12px; line-height:1.5; }
.kpi>span { min-height:3em; }
.kpi strong { display:block; font-size:36px; font-weight:700; color:#111827; margin:8px 0; font-variant-numeric:tabular-nums; }
.pill { display:inline-block; border:1px solid var(--border); border-radius:4px; padding:2px 8px; color:var(--muted); background:#f9fafb; font-size:12px; }
.leadline,.chart-stage .leadline { font-family:inherit; font-size:18px; color:#111827; line-height:1.5; margin:12px 0 16px; max-width:68ch; }
.context { color:var(--muted); font-size:12px; margin:0 0 12px; line-height:1.6; }
.definitions { background:#f9fafb; border-left:3px solid #cbd5e1; padding:12px 16px; margin:16px 0; line-height:1.6; }
details { margin:16px 0 0; border-top:1px solid var(--border); padding:12px 0 0; }
summary { cursor:pointer; font-weight:600; color:var(--primary); padding:4px 0; }
details[open]>summary { margin-bottom:12px; }
table { width:100%; font-size:13px; font-variant-numeric:tabular-nums; border-collapse:collapse; table-layout:fixed; }
th { background:#f9fafb; color:#111827; font-weight:600; }
th,td { padding:10px 12px !important; border-bottom:1px solid var(--border); vertical-align:top; overflow-wrap:anywhere; }
td.num { text-align:right; }
caption { text-align:left; color:var(--muted); font-size:12px; padding:12px 0; }
.mobile-label { display:none; }
.pillar-grid { display:grid; grid-template-columns:repeat(4,minmax(0,1fr)); gap:16px; margin:24px 0; }
.pillar-grid>a { display:block; border:1px solid var(--border); border-radius:8px; padding:16px; color:inherit; text-decoration:none; line-height:1.5; }
.pillar-grid>a:hover { border-color:#a5b4fc; background:#f8fafc; }
.pillar-grid strong { display:block; color:var(--primary); margin-bottom:8px; }
.pillar-grid small { color:var(--muted); }
.next-step { margin:20px 0 0; padding-top:16px; border-top:1px solid var(--border); }
a { color:var(--primary); }
a:focus-visible,summary:focus-visible,button:focus-visible { outline:3px solid #4f46e5; outline-offset:3px; }
@media(min-width:768px) and (max-width:1199px) {
  .navbar-brand { font-size:15px; }
  .navbar-nav>li>a { font-size:13px; padding-left:8px; padding-right:8px; }
  .chart-grid { grid-template-columns:repeat(2,minmax(0,1fr)); }
}
@media(min-width:768px) and (max-width:979px) {
  .navbar-brand { max-width:230px; white-space:normal; line-height:18px; padding-top:7px; }
  .kpis,.pillar-grid { grid-template-columns:repeat(2,minmax(0,1fr)); }
}
@media(max-width:767px) {
  .navbar-brand { font-size:15px; max-width:calc(100% - 65px); white-space:normal; line-height:18px; height:auto; }
  .dashboard-page-wrapper { padding:0; }
  .chart-wrapper { border-radius:8px; margin-bottom:16px; }
  .chart-title { padding:16px 16px 0; font-size:18px; }
  .chart-stage { padding:12px 16px 16px !important; }
  .kpis { grid-template-columns:repeat(2,minmax(0,1fr)); }
  .kpi { padding:12px; }
  .kpi>span { min-height:4.5em; }
  .kpi strong { font-size:30px; }
  .chart-grid,.pillar-grid { grid-template-columns:1fr; }
  .report-chart h4 { min-height:0; }
  .data-table thead { position:absolute; width:1px; height:1px; overflow:hidden; clip:rect(0,0,0,0); }
  .data-table,.data-table tbody,.data-table tr,.data-table td { display:block; width:100%; }
  .data-table caption { display:block; width:100%; }
  .data-table tr { border-top:1px solid var(--border); padding:8px 0; }
  .data-table td { display:grid; grid-template-columns:minmax(0,1fr) minmax(0,1fr); gap:12px; padding:6px 0 !important; border:0; text-align:left; }
  .mobile-label { display:block; font-weight:600; color:var(--muted); }
  .cell-value { min-width:0; overflow-wrap:anywhere; }
}
</style>

<script>
document.addEventListener('click', function(event) {
  var shortcut = event.target.closest('a[data-report-page]');
  if (shortcut) {
    // flexdashboard activates internal page links; move keyboard focus with them.
    event.preventDefault();
    var page = document.getElementById(shortcut.hash.slice(1));
    var heading = page.querySelector('.chart-title');
    heading.setAttribute('tabindex', '-1');
    heading.focus({preventScroll:true});
  }
  if (window.innerWidth < 768 && event.target.closest('.navbar-nav a[data-toggle="tab"]')) {
    var navbar = document.getElementById('navbar');
    var toggle = document.querySelector('.navbar-toggle');
    // Bootstrap ignores hide during opening; finish it for fast menu selections.
    if (navbar.classList.contains('collapsing') && toggle.getAttribute('aria-expanded') === 'true') {
      window.jQuery(navbar).one('shown.bs.collapse', function() {
        window.jQuery(navbar).collapse('hide');
      });
    }
  }
});
</script>

Overview
=====================================

Row
-------------------------------------

### Where should we look more closely at developer working conditions?

<span class="pill">Synthetic demo · Real export schema</span>
<span class="context">`r period_label` · `r num(nrow(baseline))` developers · `r n_distinct(baseline$Team)` teams · `r BASELINE_WEEKS` complete Person Query weeks</span>

<p class="leadline">The median developer records <strong>`r f1(median(baseline$Collaboration_hours))` collaboration hours per week</strong>; use the four pillars to explore team differences, calendar availability and recorded Copilot use.</p>

```{r landscape-kpis, results='asis'}
cat('<div class="kpis">')
card('Collaboration hours / week', f1(median(baseline$Collaboration_hours)), 'Median of person-level eight-week averages')
card('Available-to-focus hours / week', f1(median(baseline$Available_to_focus_hours)), 'Calendar availability; tenant working-hours rules apply')
card('After-hours collaboration / week', f1(median(baseline$After_hours_collaboration_hours)), 'Median hours outside configured working hours')
card('Developers with recorded GitHub use', num(github_active_n),
  paste('Among', num(github_observed_n), 'developers with complete activity coverage'))
cat('</div>')
```

<p class="chart-note">Synthetic examples, not productivity scores or intervention effects. Microsoft 365 complete-window rates are unavailable without independent collection evidence; missing observation is not non-use.</p>

```{=html}
<nav class="pillar-grid" aria-label="Explore report pillars">
<a href="#collaboration" data-report-page><strong>Collaboration &rarr;</strong><small>Where is coordination load concentrated, and which communication modes contribute?</small></a>
<a href="#focus" data-report-page><strong>Focus &rarr;</strong><small>How much calendar time is available, uninterrupted or fragmented?</small></a>
<a href="#after-hours" data-report-page><strong>After-hours &rarr;</strong><small>How much collaboration falls outside working hours, and how often?</small></a>
<a href="#ai-use-and-query-coverage" data-report-page><strong>AI &rarr;</strong><small>What GitHub use is recorded, and what can each product feed support?</small></a>
</nav>
```

<details><summary>Baseline percentiles and how to read the working week</summary>

```{r landscape-tables, results='asis'}
html_table(work_summary, paste('Population percentiles across', num(nrow(baseline)), 'developers’ eight-week averages |', period_label, '| Units are stated per measure; hours and distinct-people counts are not comparable.'))
```

<div class="definitions">
<strong>Reading the working week.</strong> Collaboration hours include meeting hours, scheduled calls, email and chat activity, so they must not be added to meeting hours. Available-to-focus hours are the hours remaining during working hours after excluding meetings and scheduled Teams calls. Uninterrupted hours and interrupted hours partition available-to-focus hours. The working-hours basis is configurable per tenant through metric rules, so the 40-hour basis in this synthetic file is not universal. Available-to-focus hours are calendar availability; they do not measure coding time or psychological flow.
</div>

The roster contains `r num(nrow(people))` synthetic people. The `IsDeveloper` attribute selects developers, not their product activity. Product feeds cover weekdays in `r product_label`, a shorter window than the Person Query history. <a href="#appendix" data-report-page>See source and observation contracts.</a>

</details>

Collaboration
=====================================

Row
-------------------------------------

### How is collaboration load distributed across teams?

<p class="context">Synthetic demo · `r period_label` · Person-level eight-week averages</p>

<p class="leadline"><strong>`r team_league$Team[1]`</strong> has the highest median collaboration load at <strong>`r f1(team_league$Collaboration_hours[1])` hours/person/week</strong>.</p>

<p class="chart-note">Total collaboration overlaps its components: compare the panels, do not add them. These are working conditions, not delivery-quality measures.</p>

```{r collaboration-intervals, results='asis'}
interval_cards(baseline,
  c('Collaboration_hours','Meeting_hours','Email_hours','Chat_hours',
    'Scheduled_call_hours','Unscheduled_call_hours'))
```

<details><summary>Exact team medians</summary>

```{r collaboration-table, results='asis'}
html_table(team_league |>
  transmute(Rank, Team, `Developers (count)`=Developers,
    `Collaboration hours`=f1(Collaboration_hours),
    `Meeting hours`=f1(Meeting_hours),
    `Email hours`=f1(Email_hours),
    `Chat hours`=f1(Chat_hours),
    `Scheduled call hours`=f1(Scheduled_call_hours),
    `Unscheduled call hours`=f1(Unscheduled_call_hours)),
  paste('Sorted by collaboration hours descending | Hours/person/week |', period_label,
        '| These are descriptive working-condition metrics, not delivery-quality measures.'))
```

</details>

<details><summary>Team and role composition context</summary>

```{r roster-composition, results='asis'}
render_chart(stack_plot(team_context, 'Roles represented in each team',
  'Roster composition provides context for coordination needs.',
  base_caption(unit='share of each team, 60 developers per team')))
```

```{r composition-details, results='asis'}
html_table(composition_table(team_context), 'Role shares use the 60 developers in each team as denominator.')
html_table(composition_counts |> transmute(Attribute, Category, `Developers (count)`=People,
  `Share of developers (%)`=pct(People/nrow(baseline))), 'Role, seniority and tenure are illustrative HR attributes carried by the query exports.')
```

</details>

<details><summary>Meeting characteristics: recurring, conflicting and short notice</summary>
```{r meeting-characteristics, results='asis'}
characteristics <- interval_data(baseline, c('Recurring_meeting_hours','Conflicting_meeting_hours','Meeting_hours_with_six_or_fewer_hours_of_advanced_notice')) |>
  group_by(Metric) |> arrange(desc(p50), .by_group=TRUE) |> ungroup() |>
  transmute(Team=Group, `Meeting characteristic`=Metric, `25th percentile`=f1(p25), Median=f1(p50), `75th percentile`=f1(p75))
html_table(characteristics, paste('Hours/person/week |', period_label, '| 60 developers per team. Characteristics overlap and must not be added.'))
```

Recurring meeting hours cover meetings set to recur. Conflicting hours count only the overlapping portion of calendar meetings. Short-notice meetings were scheduled six hours or less before their start. These are overlapping characteristics, not an additive stack.

</details>

<p class="next-step"><strong>Question to take to the team:</strong> which coordination activities unblock work, and which fragment it? Agree the workflow problem before setting a target.</p>

Row
-------------------------------------

### How broad is each team's collaboration network?

<p class="leadline"><strong>`r network_lead_team`</strong> has the largest median internal network at <strong>`r f1(network_lead_value)` people</strong>, against a population median of <strong>`r f1(median(baseline$Internal_network_size))`</strong>.</p>

<p class="chart-note">Network measures count distinct people, not hours. Viva Insights derives them over a trailing window, so they describe a standing level of connection rather than activity inside a single week. A larger network is not inherently better; it reflects role and coordination requirements.</p>

```{r network-intervals, results='asis'}
interval_cards(baseline, NETWORK_METRICS,
  caption=base_caption(unit='distinct people per person. Network measures use a trailing window and are never summed across weeks'))
```

<details><summary>Population network baseline and exact team percentiles</summary>

```{r network-tables, results='asis'}
html_table(network_summary, paste('Population percentiles across', num(nrow(baseline)),
  'developers | Distinct people |', period_label,
  '| Each developer contributes one eight-week average of a trailing-window measure.'))
html_table(interval_data(baseline, NETWORK_METRICS) |>
  transmute(Team=Group, Metric, `25th percentile`=f1(p25), Median=f1(p50), `75th percentile`=f1(p75)),
  paste('Distinct people |', period_label, '| 60 developers per team.'))
```

**Internal network size** counts distinct internal people collaborated with. **Strong ties** and **diverse ties** describe the depth and spread of those connections. **Network outside organisation** counts connections beyond the person's own organisational unit, and **external network size** counts connections outside the company. These overlap and must not be added.

</details>

Focus
=====================================

Row
-------------------------------------

### How much calendar time is available for focus?

<p class="context">Synthetic demo · `r period_label` · `r num(nrow(baseline))` developers · `r BASELINE_WEEKS` complete Person Query weeks each</p>

<p class="leadline"><strong>`r team_league$Team[which.max(team_league$Uninterrupted_hours)]`</strong> has the most uninterrupted time at <strong>`r f1(max(team_league$Uninterrupted_hours))` hours/person/week</strong>.</p>

<p class="chart-note">Calendar availability does not establish coding time or psychological flow. Working-hour rules vary by tenant; the synthetic 40-hour basis is not universal.</p>

```{r focus-intervals, results='asis'}
interval_cards(baseline, c('Available_to_focus_hours','Uninterrupted_hours','Interrupted_hours'))
```

<details><summary>Focus definitions and exact team percentiles</summary>

<strong>Definitions, paraphrased from Microsoft Learn.</strong> <strong>Available-to-focus hours</strong> are hours remaining during working hours after excluding meetings and scheduled Teams calls for focused work. <strong>Uninterrupted hours</strong> are one-hour-or-longer blocks of uninterrupted time. <strong>Interrupted hours</strong> are available-to-focus time interrupted by email, Teams chat, unscheduled calls or Teams channel activity. <strong>Open 1-hour block</strong> is a calendar count without scheduled meetings during the workday. The working-hours basis is set by tenant metric rules; the 40-hour basis in this synthetic data is not universal.

```{r focus-percentiles, results='asis'}
html_table(interval_data(baseline, c('Available_to_focus_hours','Uninterrupted_hours','Interrupted_hours')) |>
  transmute(Team=Group, Metric, `25th percentile`=f1(p25), Median=f1(p50), `75th percentile`=f1(p75)),
  paste('Hours/person/week |', period_label, '| Each person contributes one eight-week average.'))
```

</details>

<p class="next-step"><strong>Question to take to the team:</strong> what interrupts sustained work? Pair the calendar patterns with feedback on review waits and cognitive load. <a href="#collaboration" data-report-page>Explore meeting characteristics in Collaboration.</a></p>

After-hours
=====================================

Row
-------------------------------------

### How is collaboration outside working hours distributed and repeated?

<p class="context">Synthetic demo · `r period_label` · `r num(nrow(baseline))` developers · `r BASELINE_WEEKS` observed weeks each</p>

<p class="leadline"><strong>`r team_league$Team[which.max(team_league$After_hours_collaboration_hours)]`</strong> has the highest median after-hours collaboration at <strong>`r f1(max(team_league$After_hours_collaboration_hours))` hours/person/week</strong>, compared with <strong>`r f1(median(baseline$After_hours_collaboration_hours))` hours</strong> across the population.</p>

<p class="chart-note">Recorded collaboration outside configured working hours is not total work time or a wellbeing diagnosis.</p>

```{r afterhours-distribution, results='asis'}
render_chart(interval_plot(baseline, 'After_hours_collaboration_hours',
  title='After-hours collaboration distribution by team',
  subtitle='Dot: median. Line: middle 50% of person-level eight-week averages. No health bands are assigned.') +
  facet_null())
```

<details><summary>Definition and interpretation limits</summary>

After-hours collaboration is time in meetings, email, Teams chats, calls and channels with at least one other person outside working hours, with overlapping activity deduplicated. Flexible schedules, on-call duties and time zones may matter, but none is recorded here. Definition paraphrased from Microsoft Learn.

</details>

Row
-------------------------------------

### Persistence over the observed weeks

```{r persistence, results='asis'}
persistence <- baseline |> mutate(Band=case_when(Weeks_above==0 ~ '0 weeks',
  Weeks_above<PERSISTENCE_WEEKS ~ paste('1 to', PERSISTENCE_WEEKS-1, 'weeks'),
  TRUE ~ paste(PERSISTENCE_WEEKS, 'to', BASELINE_WEEKS, 'weeks'))) |>
  count(Band, name='People') |> mutate(Share=People/nrow(baseline))
stopifnot(sum(persistence$People)==nrow(baseline))
persistence <- disclose_partition(persistence)
render_chart(if (!nrow(persistence)) unavailable_plot('Persistence composition withheld') else
ggplot(persistence, aes(x=People, y=factor(Band, levels=rev(c('0 weeks', paste('1 to', PERSISTENCE_WEEKS-1, 'weeks'), paste(PERSISTENCE_WEEKS, 'to', BASELINE_WEEKS, 'weeks')))))) +
  geom_col(fill='#c7d2fe', width=.55) +
  geom_text(aes(label=paste0(num(People), '\n(', pct(Share), ')')), hjust=-.12, size=3.3) +
  scale_x_continuous(expand=expansion(mult=c(0,.4))) + theme_report() +
  chart_labs('How often the illustrative convention was met',
    paste('Convention:', AFTER_HOURS_CONVENTION, 'or more after-hours collaboration hours in a week. No burnout diagnosis is implied.'),
    base_caption(unit=paste('developer counts and share of', num(nrow(baseline)))), x='Developers (count)', y=NULL))
```

<p class="next-step"><strong>Question to take to the team:</strong> do the patterns reflect schedules, on-call duties or difficulty disconnecting? Validate work-hour settings and seek anonymous feedback; do not target individuals from these aggregates.</p>

AI {#ai-use-and-query-coverage}
=====================================

Row
-------------------------------------

### What GitHub Copilot use is recorded?

<p class="context">Synthetic demo · `r period_label` · `r num(nrow(baseline))`-developer roster</p>

<p class="leadline"><strong>`r num(github_active_n)` developers have recorded GitHub use</strong> among <strong>`r num(github_observed_n)` with complete activity observation</strong> in the baseline.</p>

<p class="chart-note">Observed means complete activity-row coverage, not licensing. Microsoft 365 complete-window rates and two-product comparisons remain unavailable without independent completeness evidence.</p>

```{r github-observation, results='asis'}
render_chart(if (!nrow(github_use_counts)) unavailable_plot('GitHub observation partition withheld') else
  ggplot(github_use_counts, aes(x=People, y=factor(GH_use, levels=rev(gh_use_levels)))) +
    geom_col(fill='#3b82f6', width=.55) +
    geom_text(aes(label=paste0(num(People), '\n(', pct(Share), ')')), hjust=-.12, size=3.3) +
    scale_y_discrete(labels=function(x) wrap(x, 18)) +
    scale_x_continuous(expand=expansion(mult=c(0,.4))) + theme_report() +
    chart_labs('Recorded GitHub use and observation',
      'No recorded use requires complete observation. Unresolved observation is not inactivity.',
      paste('Shares of the full', num(nrow(baseline)), 'developer roster. Every published group must meet the privacy floor.'),
      x='Developers (count)', y=NULL))
```

<p class="next-step"><strong>Explore the workflows:</strong> <a href="#github-copilot-breakdowns" data-report-page>feature, model and language breakdowns</a>, or <a href="#working-patterns-and-evaluation" data-report-page>working conditions by recorded-use group</a>. Neither demonstrates productivity impact.</p>

<details><summary>Why joint-product comparisons are unavailable</summary>

Microsoft 365 eligibility uses that week's `Total_Copilot_enabled_days`; static metadata cannot override a zero-enabled week. Eligibility and collection completeness are different requirements. The product calendar, static licence metadata and positive consumption rows do not certify complete collection.

```{r joint-footprint, results='asis'}
joint_stack <- joint_counts_public |> transmute(Group='Developer roster', Category=Display, People, Share)
render_chart(stack_plot(joint_stack, 'Recorded product use and eligibility or observation status',
  'Only and neither require established eligibility and observation. Any sub-10 cell withholds the entire partition.',
  base_caption(unit=paste('share of', num(nrow(baseline)), 'developers'))))
```

</details>

<details><summary>Product denominators and coverage derivation</summary>

GitHub activity includes explicit zero rows for inactive weekdays; absent rows leave activity observation unknown. GitHub credits are sparse billable-day records, so credit coverage is resolved separately and never invalidates an observed activity week. Missing Microsoft 365 values stay `NA` unless independent person-week coverage and positive enabled days justify a zero.

```{r joint-details, results='asis'}
html_table(joint_counts_public |> transmute(`Published tool-use status`=Display, `Developers (count)`=People, `Share of roster (%)`=pct(Share)),
  'The entire partition is withheld if any positive cell is below the privacy floor; totals cannot reveal a withheld remainder.')
eligibility_table <- bind_rows(lapply(c('GH','M365'), function(p) {
  valid_col <- if (p=='GH') 'GH_activity_valid' else 'M365_valid'
  b <- baseline_panel |> group_by(PersonId) |>
    summarise(eligible=all(.data[[paste0(p,'_eligible')]]),
              valid=all(.data[[valid_col]]), .groups='drop') |>
    mutate(Status=case_when(
      is.na(eligible) ~ 'Eligibility or observation unknown',
      !eligible ~ 'Not eligible for the full window',
      !valid ~ 'Eligible; completeness unresolved or incomplete',
      TRUE ~ 'Eligible and valid'))
  b |> count(Status, name='People') |> disclose_partition() |>
    mutate(Product=if(p=='GH') 'GitHub Copilot' else 'Microsoft 365 Copilot', .before=1)
}))
html_table(eligibility_table, 'Each reportable product partition reconciles to the roster. An unsafe partition is withheld in full, including its complementary cells. GitHub validity here is activity-row coverage; credit coverage is resolved separately below.')
html_table(baseline_panel |> group_by(PersonId) |>
  summarise(activity=all(GH_activity_valid), credit=all(GH_credit_valid), .groups='drop') |>
  count(`GitHub activity coverage`=if_else(activity,'Resolved','Unresolved'),
        `GitHub credit coverage`=if_else(credit,'Resolved','Unresolved'), name='People') |>
  disclose_partition() |>
  transmute(`GitHub activity coverage`, `GitHub credit coverage`, `Developers (count)`=People),
  'The credit export is sparse, so a week with no billable usage resolves to a measured zero rather than invalidating the observed activity week. Sparse credit rows never void observed activity.')
html_table(coverage_derivation, 'No standalone coverage export is invented. Weekly eligibility, activity coverage and credit coverage are separate requirements.')
```

</details>

Row
-------------------------------------

### Product-specific frequency, intensity and services

```{r product-usage, results='asis'}
html_table(product_summary |> arrange(desc(`Recorded active (%)`)) |> mutate(across(-Product, as.character)) |>
  pivot_longer(-Product, names_to='Product measure', values_to='Reported value') |>
  pivot_wider(names_from=Product, values_from='Reported value'),
  paste(period_label, '| Each product uses its own eligible and valid population. People with no recorded use remain in the denominator; credit measures require separately resolved credit coverage.'))
```

Microsoft 365 population rates are unavailable here, not zero. Credits, sessions, completion suggestions and chat requests are different units and are not added together.

<details><summary>Observed Microsoft 365 service records and concentration definition</summary>

```{r m365-service-records, results='asis'}
html_table(m365_service_mix |> transmute(Service=ServiceName, `People (count)`=People,
  `Sessions`=num(Sessions), `Credits`=num(Credits), `Credit share (%)`=pct(Share)),
  'Recorded Microsoft 365 service rows over baseline weekdays. These observed sums and shares do not establish complete collection or complete-window population rates.')
```

The concentration row is the share of GitHub credits or Microsoft 365 credits attributable to the highest-volume 10% of active developers for that product. Identities and individual rankings are never displayed.

</details>

GitHub breakdowns {#github-copilot-breakdowns data-navmenu="More"}
=====================================

Row
-------------------------------------

### Which GitHub workflows, models and languages appear?

<p class="context">Synthetic demo · `r period_label` weekdays · Daily GitHub breakdown exports</p>

<p class="leadline">The largest reportable feature is <strong>`r leader_label(feature_mix$Feature, feature_mix$Share)`</strong> at <strong>`r pct(feature_mix$Share[1])`</strong> of supplied feature usage.</p>

<p class="chart-note">Shared allocation and completion-model attribution are synthetic illustrations, not verified real-export semantics. Shares are calculated separately within each supplied breakdown.</p>

```{r github-margins, results='asis'}
margin_plot <- bind_rows(
  feature_mix |> transmute(Margin='Feature', Category=Feature, Share),
  model_mix |> transmute(Margin='Model', Category=Model, Share),
  language_mix |> transmute(Margin='Language', Category=Language, Share)) |>
  group_by(Margin) |> slice_max(Share, n=12, with_ties=FALSE) |>
  arrange(Share, .by_group=TRUE) |>
  mutate(Category_margin=factor(paste(Category, Margin, sep=' | '), levels=paste(Category, Margin, sep=' | '))) |>
  ungroup()
if (!nrow(margin_plot)) {
  render_chart(unavailable_plot('GitHub breakdowns unavailable or withheld'))
} else {
  cat('<div class="chart-grid">')
  for (margin in c('Feature', 'Model', 'Language')) {
    render_chart(ggplot(filter(margin_plot, Margin == margin), aes(x=Share, y=Category_margin)) +
      geom_col(fill='#3b82f6', width=.65 * sum(margin_plot$Margin == margin) / max(table(margin_plot$Margin))) +
      geom_text(aes(label=pct(Share)), hjust=-.1, size=3.1) +
      scale_y_discrete(labels=function(x) wrap(sub(' \\| .*', '', x), 16)) +
      scale_x_continuous(labels=label_percent(), expand=expansion(mult=c(0,.3))) +
      theme_report() + labs(x='Share of supplied usage', y=NULL),
      heading=margin, layout='list')
  }
  cat('</div>')
}
```

<details><summary>What the synthetic allocation does and does not establish</summary>

The generator constructs all four breakdown totals from accepted completions plus user-initiated chats and checks its own internal consistency. This is not an authoritative real-export metric definition; the reader does not reject inputs whose totals differ. Model attribution, including completions, is illustrative. An authoritative metric-definition source is needed before enforcing these equalities on real data. `Share` is derived here, not an export column. All linked margins and cross-tabs are withheld if a privacy gate fails.

</details>

Row
-------------------------------------

### Cross-tabs: model × feature and language × model

```{r github-crosstabs, results='asis'}
top_model_feature <- model_feature_mix |>
  arrange(desc(Count)) |>
  slice_head(n=15) |>
  transmute(Model, Feature, `People (count)`=People, `Usage count`=num(Count), `Share within model (%)`=pct(Share))
top_language_model <- language_model_mix |>
  arrange(desc(Count)) |>
  slice_head(n=15) |>
  transmute(Language, Model, `People (count)`=People, `Usage count`=num(Count), `Share within language (%)`=pct(Share))
html_table(top_model_feature, paste('Largest reportable Model x Feature cells |', period_label, '| Daily model-feature export.'))
html_table(top_language_model, paste('Largest reportable Language x Model cells |', period_label, '| Daily language-model export.'))
```

<details><summary>Language × feature detail</summary>

```{r language-feature-detail, results='asis'}
html_table(language_feature_mix |>
  arrange(desc(Count)) |>
  slice_head(n=20) |>
  transmute(Language, Feature, `People (count)`=People, `Usage count`=num(Count), `Share within language (%)`=pct(Share)),
  paste('Largest reportable Language x Feature cells |', period_label, '| Daily language-feature export.'))
```

</details>

**Manager next step.** Use these breakdowns to ask which workflows each feature supports. Do not treat agent-type features or model choice as maturity, quality, productivity or time-saved measures.

Working patterns {#working-patterns-and-evaluation data-navmenu="More"}
=====================================

Row
-------------------------------------

### How do working conditions differ across recorded tool-use groups?

<p class="context">Synthetic demo · `r period_label` · `r num(github_observed_n)` developers with complete GitHub activity coverage</p>

<p class="leadline">Compare working conditions for <strong>`r num(github_active_n)` developers with recorded GitHub use</strong> and <strong>`r num(github_observed_n - github_active_n)` observed developers with no recorded use</strong>.</p>

<p class="chart-note">Unadjusted, descriptive comparisons only. Use and working conditions share the same window, so temporal direction and intervention effects cannot be established.</p>

```{r group-intervals, results='asis'}
interval_cards(gh_observed, c('Collaboration_hours','Meeting_hours','Available_to_focus_hours','Uninterrupted_hours','After_hours_collaboration_hours'), group='GH_use')
```

<details><summary>Exact reportable group sizes and percentile values</summary>

```{r group-percentiles, results='asis'}
sizes <- gh_observed |> count(GH_use, name='People', .drop=TRUE) |> filter(People >= MIN_GROUP_N)
html_table(interval_data(gh_observed, c('Collaboration_hours','Meeting_hours','Available_to_focus_hours','Uninterrupted_hours','After_hours_collaboration_hours'), 'GH_use') |>
  left_join(sizes, by=c('Group'='GH_use'), relationship='many-to-one') |>
  filter(!is.na(People)) |>
  group_by(Metric) |> arrange(desc(p50), .by_group=TRUE) |> ungroup() |>
  transmute(Group, Metric, `Developers (count)`=People, `25th percentile`=f1(p25), Median=f1(p50), `75th percentile`=f1(p75)),
  paste('Hours/person/week |', period_label, '| Each person contributes one eight-week average. Sub-10 use groups are withheld.'))
```

</details>

<details><summary>Comparison population and limits</summary>

The comparison includes developers valid for the Person Query and GitHub activity throughout the baseline. Unresolved GitHub observation remains in the full roster but is excluded here, not recoded as no use. This is not statistical matching. A two-product comparison remains unavailable because Microsoft 365 completeness is not established. Independent completeness evidence and the privacy floor are prerequisites for extending it.

</details>

Row
-------------------------------------

### Network breadth by recorded GitHub use

<p class="chart-note">Whether recorded GitHub use coincides with broader or narrower collaboration networks. Network measures count distinct people over a trailing window, so they are standing levels rather than weekly activity. Differences are unadjusted and descriptive; neither direction implies an effect of tool use.</p>

```{r group-network, results='asis'}
interval_cards(gh_observed, NETWORK_METRICS, group='GH_use',
  caption=base_caption(nrow(gh_observed),
    unit='distinct people per person. Trailing-window measures are never summed across weeks'))
```

<details><summary>Exact network percentiles by reportable use group</summary>

```{r group-network-table, results='asis'}
html_table(interval_data(gh_observed, NETWORK_METRICS, 'GH_use') |>
  left_join(sizes, by=c('Group'='GH_use'), relationship='many-to-one') |>
  filter(!is.na(People)) |>
  transmute(Group, Metric, `Developers (count)`=People,
    `25th percentile`=f1(p25), Median=f1(p50), `75th percentile`=f1(p75)),
  paste('Distinct people |', period_label, '| Sub-10 use groups are withheld.'))
```

</details>

Row
-------------------------------------

### Role and team context for the comparison

```{r role-group-context, results='asis'}
render_chart(stack_plot(role_context, 'Role composition within the comparison groups',
  'Group sizes and roles can shape the comparison. Any group with a role cell below 10 is omitted from this breakdown.',
  base_caption(sum(role_context$People), unit='share of developers within each displayed use group')))
```

<details><summary>Exact role shares and team composition</summary>

```{r group-context-tables, results='asis'}
html_table(composition_table(role_context), 'A role breakdown is withheld in full for any group containing a cell below 10 people.')
html_table(composition_table(team_joint_context), 'A team breakdown is withheld in full for any group containing a cell below 10 people. Shares use all developers in the displayed group.')
```

</details>

**Interpretation.** Team, role, seniority, task selection and tenure can influence working patterns and tool use. These charts are intentionally limited to agreed measures. No broad metric scan, adjustment model, causal claim or preferred tool-use group is presented.

Row
-------------------------------------

### Weekly trends and valid denominators

```{r product-trends, results='asis'}
render_chart(trend |> select(Week, GH_rate, M365_rate) |>
  pivot_longer(-Week, names_to='Product', values_to='Rate') |>
  mutate(Product=recode(Product, GH_rate='GitHub Copilot', M365_rate='Microsoft 365 Copilot')) |>
  filter(!is.na(Rate)) |>
  ggplot(aes(Week, Rate, colour=Product)) + geom_line(linewidth=.9) +
  scale_colour_manual(values=c('#2563eb','#0f766e')) +
  scale_y_continuous(labels=label_percent(), limits=c(0,1)) +
  scale_x_date(date_breaks='1 month', date_labels='%d %b') + theme_report() +
  chart_labs('Recorded active developers as a share of valid weekly observation',
    'At least one positive activity day that week. Missing coverage is not inactivity; Microsoft 365 rates are unavailable by default without independent completeness evidence.',
    paste(history_label, '| Denominator: product-eligible developers with valid observation in that week. UTC week starts.'),
    x='Week beginning (2026)', y='Recorded active (%)'))
```

<details><summary>Coverage and numerators for every weekly window</summary>

```{r weekly-coverage, results='asis'}
html_table(trend |> transmute(`Week beginning (UTC)`=format(Week,'%Y-%m-%d'),
  `PQ developers (count)`=num(PQ_n), `GH valid (count)`=num(GH_n), `GH active (count)`=num(GH_active),
  `GH active (%)`=pct_or_reason(GH_rate, GH_n), `M365 valid (count)`=num(M365_n),
  `M365 active (count)`=num(M365_active), `M365 active (%)`=pct_or_reason(M365_rate, M365_n),
  `Both valid (count)`=num(Joint_n)),
  'A count of 0 is an observed absence of valid weekly observation, not a withheld value: the GitHub feed opens 13 weeks after the Person Query history, and Microsoft 365 has no valid weeks at all because its completeness is not independently established. A rate over an empty denominator is undefined rather than withheld.')
```

</details>

Row
-------------------------------------

### What evidence would support an intervention?

Select a workflow question, then combine these signals with anonymous developer feedback and delivery outcomes. Survey, review-delay, deployment and service-quality outcomes are not in these exports. <a href="#appendix" data-report-page>See the SPACE, DevEx and DORA evidence guide in Methods.</a>

**Decision supported now.** Select a workflow question and close its measurement gaps. An intervention effect, productivity gain or wellbeing conclusion requires additional evidence. No operational action is justified by synthetic values.

Methods {#appendix data-navmenu="More"}
=====================================

Row
-------------------------------------

### Definitions, sources and reporting contracts

<span class="pill">Synthetic demo · Real export schema</span>

**Scope and units.** Pages use the last eight complete Person Query weeks, `r period_label`, and product weekdays that fall in the final 13-week product window. The full Person Query source window is `r history_label`, with Sunday-start weeks. Dates are ISO dates interpreted using a UTC convention. This is an illustrative date contract, not evidence about developers’ time zones. Available-to-focus hours are based on working hours configured through tenant metric rules; the 40-hour basis in this synthetic file is not universal.

**Aggregation.** Working-condition values are weekly Person Query metrics. Baseline charts first average each person’s eight weekly values and then compute group percentiles. Each developer has equal weight. Product-active rates divide active eligible, observed developers by all eligible, observed developers for that product and window. Daily Microsoft 365 service rows are aggregated before person-day and person-week analysis.

**Source definitions.** Metric descriptions are paraphrased from the [Microsoft Viva Insights metrics reference](https://learn.microsoft.com/en-us/viva/insights/advanced/reference/metrics). The source manifest is `_data/README.md`; the report reads the committed CSVs listed there.

```{r source-contracts, results='asis'}
html_table(source_contracts, 'CSV files read by github-developer-experience-helpers.R. No zip, external download or data-generation step is run by this report.')
```

**Joins and missingness.** Person Query anchors the roster and weekly panel. Consumption metadata is keyed by `PeopleHistoricalId`, which is an opaque key read from the activity-file crosswalk and never reconstructed from `PersonId`; the Person Query is used for HR attributes because it already carries `PersonId`. Daily feeds aggregate to unique person-week keys before left joins. One-to-one and many-to-one join assertions prevent row amplification. `0` is an observed or justified zero. `NA` is unresolved measurement and is never silently replaced outside valid observation. GitHub activity coverage and GitHub credit coverage are separate flags, each gating only the measures it governs.

Row
-------------------------------------

### Privacy, reproducibility and validation

**Privacy floor.** At least 10 distinct people are required for a published group. Percentile visuals use `vivainsights::create_boxplot(..., mingroup = 10, return = "table")`. Linked HR margins and cross-tabs share one release gate: every cell of every published margin must clear the floor, so no published margin can be differenced against another to expose a sub-floor group. Joint-use and eligibility partitions are withheld in full when unsafe. Linked GitHub margins and cross-tabs are withheld together, and their release additionally requires that no person's pattern of cell membership is shared by fewer than 10 people. Service releases check positive contributors and overlapping membership-pattern complements within the developer roster. Product rates check both numerator and complement; unavailable values are labelled, not plotted as zero. Missing category labels are shown as “Unknown / not supplied”. No individual points, identifiers or individual leaderboards appear in the HTML.

**Simulation boundary.** The CSV values are synthetic, but the report no longer invents CSV schemas or writes files. It does not use a standalone eligibility or coverage export, fabricated Microsoft 365 action categories, model or language share files, or an agent-use indicator that is absent from the real GitHub activity export. It parses the real `Agent adoption` field and derives all shares locally.

**Coverage boundary.** There is no default fixture-completeness exception or filename-based trust. `derive_m365_coverage(person_weeks, evidence = NULL)` returns unknown completeness unless a caller explicitly supplies independently obtained, person-week-specific evidence with its source. That is an in-memory analysis input, not a fabricated export column or file. A calendar, static licence status, observed positive rows, or missing rows cannot supply that evidence. A future fixture-only opt-in would need validated provenance tied to the exact files; assumptions must not survive replacement or row deletion.

```{r validation-table, results='asis'}
html_table(validation_summary, 'Automated data-contract checks executed during this render.')
```

<details><summary>Runtime versions and package choice</summary>

```{r versions, results='asis'}
html_table(tibble(Component=c('R','dplyr','tidyr','ggplot2','vivainsights','rmarkdown','flexdashboard'),
  Version=c(as.character(getRversion()), vapply(c('dplyr','tidyr','ggplot2','vivainsights','rmarkdown','flexdashboard'),
    function(p) as.character(packageVersion(p)), character(1)))))
```

The [vivainsights function index](https://microsoft.github.io/vivainsights/llms.txt) supports the person-averaged boxplot summary used for percentile calculations. Custom joint coverage aggregation is necessary because single-product usage segments do not establish comparable observation for two products.

</details>

Row
-------------------------------------

### Research framework and future extensions

| Framework | What this report can contribute | Evidence still needed |
|---|---|---|
| SPACE | Activity, collaboration and calendar-derived working conditions across multiple dimensions | Satisfaction and wellbeing surveys, performance outcomes, direct evidence of efficiency and flow |
| DevEx | Context for feedback loops, cognitive load and flow discussions | Anonymous developer feedback, build/test waits, review waits, task context and perceived ability to focus |
| DORA | A framework for evaluating delivery change, without an outcome claim | Service boundaries, production deployments, change lead times, recovery events, failed changes and deployment rework |

- [The SPACE of Developer Productivity](https://queue.acm.org/detail.cfm?id=3454124), ACM Queue, 2021. Productivity is multidimensional and cannot be reduced to one activity measure.
- [DevEx: What Actually Drives Productivity](https://queue.acm.org/detail.cfm?id=3595878), ACM Queue, 2023. Connect feedback loops, cognitive load and flow with developer feedback and system measures.
- [DORA’s software delivery performance metrics](https://dora.dev/guides/dora-metrics/). Current guidance describes five metrics across throughput and instability, best applied to one application or service at a time.
- [Viva Insights metrics reference](https://learn.microsoft.com/en-us/viva/insights/advanced/reference/metrics). Primary reference for the paraphrased collaboration and focus definitions.
